 \documentclass[12pt,a4paper]{article}

\newcommand{\sharedmacros}{../../SharedMacros}
\input{\sharedmacros/ccategories.macros}
\input{\sharedmacros/ermacros}
\input{\sharedmacros/erdiagram}
\input{\sharedmacros/theorems.macros}
\input{\sharedmacros/general.macros.v2}
\usepackage{mathptmx}  % This changes font to roman
\usepackage{anyfontsize}
\usepackage{mathtools}  % why have we got this?
\usepackage{alltt}    
\usepackage{mnsymbol} %used for rightpitchfork
\usepackage{cmll}
\usepackage{ulem}
\renewcommand{\ttdefault}{txtt}
\usepackage[left=1.5cm, right=4cm, marginparwidth=3cm, top=2cm, bottom=2.0cm]{geometry}
\usepackage{framed}
\usepackage[font=small]{caption}
\setlength{\captionmargin}{2cm}
\theoremstyle{remark}
\newtheorem*{lemma*}{Lemma}

\usepackage{arydshln} % vertical dashed lines between columns of an array

%\input{../macros/generalmacros.tex}
\input{../macros/othermacros.tex}
%\input{../macros/beamerclashmacros.tex}
%ENDCOPY

\begin{document}
\section{Categories with Designated Monos and Epis}
\newcommand{\pause}{}
\input{content.macros.tex}


\subsection{Hom functor}
\begin{lemma}
If $a$ is any object of a category \catcw then the Hom functor 
$Hom(a,-) : \catc \morph \Set$ preserves monomorphims.
\end{lemma}
\begin{proof}
Pretty much the definition of a morphism $j:x \morph y$ being a monomorphism is that 
the function $Hom(a,j)$ is injective for any object $a$.
\end{proof}

\begin{lemma}
If \catcw is a category and $F: \catc \morph \Set$ is a funtor then
if $e:x \morph y$ is an epimorphism that is split by $e_s: y \morph x$
then $F(e)$ is an epimorphism that is split by $F(e_s$.
\end{lemma}
\begin{proof}
Because for any functor $F$, $e_s \circ e = id_y$ in \catcw implies 
$F(e_s) \circ F(e) = id_F(y)$ in \Set.
\end{proof}

\begin{lemma}
If $F: \catcw \morph \Set$ and $G: \catcw \morph \Set$ are functors,
if $j: x \morph y$ is a monomorphism in \catc,
if $F(j)$ and $G(j)$ are injective i.e. are monomorphisms in \Set then
then the coproduct $F+G$ in the functor category $\Set^\catc$ preserves the monomorphism
$j$ i.e. (F+G)(j) is alo injective.
\end{lemma}

\subsection{Definitions}
\begin{definition}
If \scalebox{0.9}{\fgsourcediagram} in a \catMEterm \catcw  and if $\reqtc$ is a set of instances of \catcw
then we say that there is a  \textit{functional dependency} of $g$ on $f$ with respect to $\reqtc$ iff
for every $D \in \reqtc$, $D(g)$ factors through $D(f)$. \\
\medskip
\pause Equivalently there is a family of  functions $H_D:img(D(f)) \morph D(c))_{D \in \reqtc}$
such that in each instance $D \in \reqtc$,
\scalebox{0.9}{\fghfactordiagram{\roomup{0.4cm}\roomdown{0.3cm}D(a)}{img(D(f))}{D(c)}{D(f)}{D(g)}{H_D}} commutes.
\end{definition}

\begin{definition}
If \catcw is a category with designated monomorphisms and epimorphsims 
considered as a data specification with requirement $\reqtc$, 
if $\fundep{H}{f}{g}$  is a functional dependency with respect to $\reqtc$
then the functional dependency is weakly represented in \catcw iff
there  is a diagram\\
\medskip
%\begin{minipage}[t][.45cm][t]{5.5cm} %just to get alignment at top small height content pos t 
\scalebox{0.9}{\representationdiagram} 
%\end{minipage} 
in \catcw such that for each $D \in \reqtc$
%\begin{minipage}[t]{4.5cm}
\scalebox{0.85}{\mappedrepresentationdiagram}
%\end{minipage}
commutes.
\end{definition}

\begin{lemma}
If \catcw is a category with designated monomorphisms and epimorphsims 
considered as a data specification with requirement $\reqtc$ then
\begin{enumerate}[(i)]
\item if 
$
\begin{array}{c p{1cm} c p{1cm} c}
          && \Rnode{b}{b} \\[0.6cm]
\Rnode{a}{a} && \Rnode{x}{x} && \Rnode{y}{y}\\[0.2cm]
\end{array}
\begin{arrows}
\ncarr{a}{b}
\alabel{f}
\ncarr{a}{x}
\blabel{k}
\ncarr{x}{y}
\blabel{m}
\addedgebar
\end{arrows}
$
in \catc, where $m$ is a designated monomorphism, then if there is a functional dependency 
$\fundep{H}{f}{k \circ m}$ in \catcw with respect to  $\reqtc$
then there is a functional dependency $\fundep{H'}{f}{k}$ in \catcw with respect to $\reqtc$. If $H$ is weakly represented in \catcw then so to is $H'$.
\item if 
$
\begin{array}{c p{1cm} c p{1cm} c}            
             && \Rnode{b}{b}                              \\[0.6cm]
\Rnode{a}{a} && \Rnode{x}{x} && \Rnode{y}{y}  \\[0.2cm]
\end{array}
\begin{arrows}
\ncarr{a}{b}
\alabel{f}
\ncarr{a}{x}
\alabel{k}
\ncarr{x}{y}
\alabel{h}
\end{arrows}
$
in \catcw then if there is a functional dependency 
$\fundep{H}{f}{k}$ in \catcw with respect to $\reqtc$
then there is a functional dependency $\fundep{H'}{f}{f \comp h}$
with respect to $\reqtc$. If $H$ is weakly represented in \catcw then so to is $H'$.
\end{enumerate}
\end{lemma}
\begin{proof}
Easy, 
\begin{enumerate}[(i)]
\item 
Because $m$ is a designated monomorphism then for each instance $D$,
the function $D(m)$ is injective and therefore has an inverse $D(m)^{-1}$.
Define $H'_D$ to be $H_D \circ D(m)^{-1}$.  
\item
Define $H'_D$ to be $H_D \circ D(h)$.
\end{enumerate}
In both cases it is evident, from the definition of weakly represented, that if
$H$ is weakly represented in \catcw then so to is $H'$. In the first case the diagram
that constitues the weak representation of $H$ may be argmented by the pair of the
designated monomorphism $m$ and the identify morphism. In the second case the diagram
is argmented by composing $h_n$ with $h$.
\end{proof}

\subsection{Representation Lemmas}

\begin{lemma}
If \catcw is a category with designated monomorphisms and epimorphsims  and if \catcw is
\textit{maximally constrained} to a requirement $\reqtc$ then every functional dependency
 $\fundep{H}{f}{g}$ with respect to $\reqtc$ is weakly represented in \catcw.
\end{lemma}
\begin{proof}
Suppose that we have \fgsourcediagram in \catcw and a functional dependency
$\fundep{H}{f}{g}$. We need show the functional dependency is represented in \catc.

Define functor $F: \catc \morph \Fin$ 
to be the coproduct $Hom_{\catc}(a,-) + Hom_{\catc}(a,-)$
with injections $L$ and $R$, respectively so that
for each object $x$ of \catcw the diagram
\begin{center}
$
\begin{array}{c p{0.5cm} c p{0.5cm} c  }
\Rnode{h1}{Hom_{\catc}(a,x)}  &&\Rnode{Fx}{F(x)}  &&   \Rnode{h2}{Hom_{\catc}(a,x)}       
\end{array} 
$
\ncarr{h1}{Fx}
\alabel{L_x}
\ncarr{h2}{Fx}
\blabel{R_x}
\end{center}
is a coproduct in $\Fin$.
\medskip
Define $G: \catc \morph \Fin$ as quotient $F/\sim$ where 
to ensure $G$ preserves designated monomorphisms  $\sim$ is defined as follows
\begin{align*}
L_x(k_1) \sim_x R_x(k_2) & \mbox{ \parbox{6cm}{iff  $k_1=k_2$ and there is a functional dependency 
$f \morph k_1$ in \catcw with respect to $\reqtc$ and  this functional dependency is represented in \catc.}}\\
L_x(k_1) \sim_x L_x(k_2) & \mbox{ iff $k_1 = k_2$,} \\
R_x(k_1) \sim_x R_x(k_2) & \mbox{ iff $k_1 = k_2$.} \\
\end{align*}

Now extend \catcw to \catcpw so that $f$ has an epi-mono factorisation in \catcpw
and so that there  is a morphism $\qq{h}$ such that
\scalebox{0.9}{\unaryfdrepresentationdiagram{\roomup{0.6cm}\roomdown{0.4cm}a}{b}{c}{f}{g}{im(f)}{f_m}{\qq{h}}$
\begin{arrows}\ncarr{a}{d}\alabel{f_e}\end{arrows}$} 
commutes in \catcp. \\
\begin{newtt}I guess the idea here is to designate $f_e$ as an epimorphism. Equally we could simple define an $f_s$ that splits $f_e$ i.e. such that $f_s \circ f_e = id_im{f}$.
\newt{BUT THEN COULD GET RID $\qq{h}$ because would have $\qq{h}=f_s \circ g$
BUT THERE AGAIN PROVISO THAT $f_s$ IS INFORMATION IRRELEVANT.}.

HANG ON THOUGH. Each D can be extended in multiple ways because there can be lots of different splitting functions. CRIKEY!

Need every morphism to be split by a partial. Here $f$ will be split and the splitting of $f_e$ will be $f_m \circ f_s$
\end{newtt}.

\begin{newtt}
BIG THING HERE: Show that every $D \in \reqtc$ extends uniquely to
a functor $D' \catcp \morph \Set$.
\end{newtt}

We have \simpleunaryfdrepresentationdiagram{a}{Im(f)}{c}{f_e}{g}{\qq{h}}  in \catcp.\\
\medskip
Extend $G$ to $G':\catcp \morph \Fin.$ \\
Because $f_m$ is a designated monomorphism and $f_e \circ f_m = f$ then
there is a representation of functional dependency $f \morph f_e$ in \catc
\newt{Not so $f_e$ isn't a morphism of \catc}.
Hence  we have that $[L_b(f_e)]= [R_b(f_e)]$.
\medskip
\begin{align*}
&\mbox{i.e.}& G'(f_e)([L_a(id_a)])              &= G'(f_e)([R_a(id_a)])       \\
&\therefore & G'(\qq{h}) (G'(f_e)([L_a(id_a)])) &= G'(\qq{h}) (G'(f_e)([R_a(id_a)])) \\
&\therefore & G'(g)([L_a(id_a)])              &= G'(g)([R_a(id_a)]) & &    \\
&\therefore & [L_c(g)] &= [R_c(g)] 
\end{align*}
Therefore by definition of $\sim$ the functional dependency
$\fundep{H}{f}{g}$ is represented in $\catc$. \commentary{must mean that sequence of 
back facing monos and morphisms}
\end{proof}

I want to achieve representations of fds as described in this lemma but I really don't want the assumption that the underlying catagory is maximally constrained to the requirement.

There is a sublemma regarding extending a Fin valued functor that preserves monos to one that preserves monos and epis. Not obvious where this sublemma is used. 

\newt{I now believe that I can show ref inc deps are simply
represented by assuming epis are split. The proof using Hom functor from the preparation paper goes through fine. In turn the following lemma now dfoesn't need assume 
that \catc maximally constrained. Big improvement.}  
\begin{lemma}
\representationLemmaPartTwo
\end{lemma}
\begin{proof}
My guess is that I make the additional assumption that the underlying category is maximally constrained to the requirment in order that I can then use the fact that referential inclusion depedencies are represented to inder the existence of $f_1, ... f_n$.

In daybook 17th June 2022 I have a counter example to all referential inclusion dependencies (RIDs) being represented. This counter example is circumvented if all epis are split.
This suggests to me that I should be looking for all epis being split rather than underlying category being maximally constrained.

I don't know where a proof that RIDs are represented along the lines of the proof in the category case breaks down. The breakdown should be circumvented if epis split. 
Actually I have hand written notes from today that suggests that the proof breaks down because the functor hom(a,-) does not preserve (designated) epis unless all (designated) epis are split. It will be a step forward if I can write this up. Assume designated epis split then we get RIDs represented. You might think that we could then use this proof and the extension to a catgeory in which epis split to get all RIDs represented. The counter example of 11th June 2022 might show otherwise.  
\end{proof}


\begin{example}
Suppose sketch S for a category with designated monos and epis 
has these objects and edges 
$$
\begin{array}{c c p{0.5cm}  c c}
\Rnode{s}{s} &            &&\Rnode{b}{b} &               \\[1.0cm]
             &\Rnode{a}{a}&&             & \Rnode{c}{c}
\end{array}
\begin{arrows}
\ncarr{s}{b}
\alabel{g}
\ncarr{b}{c}
\alabel{m}
\addedgebar
\ncline{->>}{s}{a}
\blabel{e}
\ncarr{a}{c}
\alabel{f}
\end{arrows}
$$ and such that this diagram is defined as a commuting diagram.
Consider the category \catcw with designated monos and epis generated by this sketch and
consider that the requirement $\reqtc$ is the category of all mono and epi preserving functors from \catcw to the category \Fin. Since all epis in \Fin are split it follows
that for any functor $F: \catc \morph \Fin$ that $(F^{-1}(e) \circ F(g)) \circ F(m) = F(f)$.

Therefore there is a referential inclusion dependency
$$
a[f] \subseteq b[m]
$$
but this referential inclusion dependency is not represented in \catc.
\end{example}

Two takeaways from this example
\begin{itemize}
	\item if $S$ is seens as a skecth for a category with designated monomorphisms and 
	designated epimoorphisms with designated splits then the  referential inclusion dependency is represented since $e_s \circ g$ will represent it where 
	$e_s$ is the spli of epimorphism $e$.
	\item alternatively we could investigate the consequences of saying that the 
	referential inclusion dependency is represented by the diagram
$
\begin{array}{c p{0.1cm} c p{0.25cm}  c }
\nudgeup{0.6cm}\Rnode{s}{s} &&            &&\Rnode{b}{b}   \\[0.75cm]
             &&\Rnode{a}{a}&&            
\end{array}
\begin{arrows}
\ncarr{s}{b}
\alabel{g}
\ncline{->>}{s}{a}
\blabel{e}
\end{arrows}
$
We can guess straight away that we are going to have to close the definition transitively and define the representation of inclusion dependencies to be in the general case

$$
\begin{array} {c c c c c c c c c c c}
     & \makebox[1cm][c]{\Rnode{s1}{$s_1$}} && 
       \makebox[1cm][c]{\Rnode{s2}{$s_2$}}  &&&&&& 
       \makebox[1cm][c]{\Rnode{sn}{$s_n$}} & \\[0.75cm]
\Rnode{a}{a} && \Rnode{d1}{d_1} && \Rnode{d2}{d_2} && \hdots 
                      && \Rnode{dn1}{d_{n-1}} && \Rnode{b}{b} 
\end{array}
\begin{arrows}
\ncline{->>}{s1}{a}
\blabel{e_1}[0.4]
\ncline{->>}{s2}{d1}
\blabel{e_2}[0.4]
\ncline{->>}{sn}{dn1}
\blabel{e_n}[0.4]
\ncarr{s1}{d1}
\alabel{g_1}[0.4]
\ncarr{s2}{d2}
\alabel{g_1}[0.4]
\ncarr{sn}{b}
\alabel{g_n}[0.4]
\end{arrows}
$$
Are there notions of weak and strong representation?
\end{itemize}









\end{document}

