\documentclass[10pt,a4paper,fleqn]{article}
\usepackage{amsmath}
		%	\usepackage{mathtools}
		%	\usepackage{alltt}
\setlength{\mathindent}{0pt}
%\usepackage[margin=3cm]{geometry}
%\usepackage{pstricks}
%\usepackage{pst-node}
%\usepackage{pst-tree}
%\usepackage{stmaryrd}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{verbatim}
%\usepackage{enumerate}
%\usepackage{calc}
%\usepackage{xifthen}
%\usepackage{array}
%\input{../SharedMacros/ccategories.macros}
%\input{../SharedMacros/gats.macros}
%\input{../SharedMacros/metagat.macros}
%\input{../SharedMacros/erdiagram}
%\usepackage{tipa}
%\usepackage{textcomp}
%\usepackage{mnsymbol}
%\usepackage{ifsym}
%\usepackage{esvect}
%\usepackage{float}
%\usepackage[section]{placeins}
%\usepackage{graphicx}
%\setcounter{equation}{0}

\newcommand{\gat}[1][U]{
\ensuremath{\mathcal{#1}} }
\newcommand{\isagat}[1][U]{\gat[#1] is a g.a.t.}
\newcommand{\heads}[1][U]{\ensuremath{H_{\gat[#1]}}}

\title{Unification in the presence of sequence variables}
\author{John Cartmell}
\begin{document}
\maketitle
\bibliographystyle{plain}

\section{Examination of cases}

Unify $\langle s_1,...s_n \rangle$ with  $\langle t_1,...t_m \rangle$. \\

\noindent
\textbf{Case: Both $s_1$  and $t_1$ are sequence variables.} \\

\noindent
The union of the following cases. \\

\noindent
Both $s_1$ and $t_1$  replaced by the empty sequence:
\begin{equation}
\label{s1emptyt1empty}
 (s_1 \mapsto \langle \rangle,t_1 \mapsto \langle \rangle)
      .unify(\langle s_2,...s_n \rangle[\langle \rangle|s_1, \langle \rangle|t_1],
             \langle t_2,...t_m \rangle[\langle \rangle|s_1, \langle \rangle|t_1]
						) 
\end{equation}

\noindent
Sequence variable $s_1$ replaced by the empty sequence and $t_1$ replaced by some non-empty 
initial subsequence of the subterms $s_2,...s_n$:
\begin{multline}
\label{s1emptyt1nonempty}
\bigcup_{2 \leq i \leq n}
(s_1 \mapsto \langle \rangle, 
 t_1 \mapsto \langle s_2,...s_i\rangle
)
.unify(\langle s_{i+1},...s_n \rangle
       [\langle \rangle|s_1, \langle s_2,...s_i \rangle|t_1], \\
        \langle t_2,...t_m \rangle[\langle \rangle|s_1, \langle s_2,...s_i \rangle|t_1]
						)
\end{multline}

\noindent
Sequence variable $t_1$ replaced by the empty sequence and $s_1$ replaced by some 
non-empty initial subsequence of the subterms $t_2,...t_m$:
\begin{multline}
\label{s1nonemptyt1empty}
\bigcup_{2 \leq j \leq m}
(s_1 \mapsto \langle t_2,...t_j \rangle, 
 t_1 \mapsto \langle \rangle
)
.unify(\langle s_2,...s_n \rangle[\langle t_2,...t_j \rangle|s_1, \langle \rangle|t_1],\\
             \langle t_{j+1},...t_m \rangle[\langle t_2,...t_j \rangle|s_1, \langle \rangle|t_1]
						) 
\end{multline}

\noindent
Sequence variable $s_1$ replaced by some
non-empty non-singleton initial subsequence of the subterms $t_1,...t_m$ :
\begin{equation}
\label{s1extendst1}
 \bigcup_{2 \leq j \leq m}{(s_1 \mapsto \langle t_1,...t_j \rangle)
      .unify(\langle s_2,...s_n \rangle[\langle t_1,...t_j \rangle|s_1],
             \langle t_{j+1},...t_m \rangle[\langle t_1,...t_j \rangle|s_1]
						) }
\end{equation}

\noindent
Sequence variable $s_1$ replaced by $t_1$: 
\begin{equation}
 \label{s1mapstot1}
 (s_1 \mapsto \langle t_1 \rangle)
      .unify(\langle s_2,...s_n \rangle[\langle t_1 \rangle|s_1],
             \langle t_2,...t_m \rangle[\langle t_1 \rangle|s_1]
						) 
\end{equation}	

\noindent
Sequence variable $t_1$ replaced by some
non-empty non-singleton initial subsequence of the subterms $s_1,...t_n$ :
\begin{equation}
\label{t1extendss1}
 \bigcup_{2 \leq i \leq n}{(t_1 \mapsto \langle s_1,...s_i \rangle)
      .unify(\langle s_{i+1},...s_n \rangle[\langle s_1,...s_i \rangle|t_1],
             \langle t_2,...t_m \rangle[\langle s_1,...s_i \rangle|t_1]
						) }
\end{equation}



\noindent
\textbf{Case: $s_1$ is a sequence variable and $t_1$ is not a sequence variable.} \\

\noindent
The union of the following: \\

\noindent
Sequence variable $s_1$ replaced by empty:
\begin{equation}
 \label{s1empty}
 (s_1 \mapsto \langle \rangle)
      .unify(\langle s_2,...s_n \rangle[\langle \rangle|s_1],
             \langle t_1,...t_m \rangle[\langle \rangle|s_1]
						) 
\end{equation}

\noindent
Sequence variable $s_1$ replaced by singleton $t_1$  as in equation (\ref{s1mapstot1}) above. \\

\noindent
Sequence variable $s_1$ replaced by some
non-empty non-singleton initial subsequence of the subterms $t_1,...t_m$ 
as in equation \ref{s1extendst1} above. \\

\noindent
\textbf{Case: $s_1$ is not a sequence variable amd $t_1$ is a sequence variable.} \\

\noindent 
The union of the following: \\

\noindent
Sequence variable $t_1$ replaced by $s_1$:
\begin{equation}
\label{t1mapstos1}
 (t_1 \mapsto \langle s_1 \rangle)
      .unify(\langle s_2,...s_n \rangle[\langle s_1 \rangle|t_1],
             \langle t_2,...t_m \rangle[\langle s_1 \rangle|t_1]
						)
\end{equation}

\noindent
Sequence variable $t_1$ replaced by empty:
\begin{equation}
 \label{t1empty}
 (t_1 \mapsto \langle \rangle)
      .unify(\langle s_1,...s_n \rangle[\langle \rangle|t_1],
             \langle t_2,...t_m \rangle[\langle \rangle|t_1]
						) 
\end{equation}
 

\noindent
Sequence variable $t_1$ replaced by some
non-empty non-singleton initial subsequence of the subterms $s_1,...s_n$ 
as in equation (\ref{t1extendss1}) above. \\

\subsection{Summary of Coding of the algorithm.}
The algorithm computes the union of the following: 
\begin{quote}
(\ref{s1emptyt1empty}) --- if both $s_1$ and $t_1$ are sequence variables \\
(\ref{s1emptyt1nonempty}) --- if both $s_1$ and $t_1$ are sequence variables \\
(\ref{s1nonemptyt1empty}) --- if both $s_1$ and $t_1$ are sequence variables \\
(\ref{s1extendst1}) --- if $s_1$ is a sequence variable  \\
(\ref{s1mapstot1}) --- if $s_1$ is a sequence variable \\
(\ref{t1extendss1}) --- if $t_1$ is a sequence variable \\
(\ref{s1empty}) --- if $s_1$ is a sequence variable and $t_1$ is not\\
(\ref{t1mapstos1}) --- if $t_1$ is a sequence variable and $s_1$ is not\\
(\ref{t1empty}) --- if $t_1$ is a sequence variable and $s_1$ is not \\


\end{quote}

\bibliography{../SharedBibliography/temp/bibliography}
 
\end{document}
