\documentclass[14pt,a4paper]{scrartcl}

\input{../SharedMacros/theorems.macros}
\input{../SharedMacros/ccategories.macros}
\input{../SharedMacros/gats.macros}
\input{../SharedMacros/metagat.macros}
\input{../SharedMacros/ermacros}
\input{../SharedMacros/erdiagram}
\input{../SharedMacros/indexedsets.macros}

\usepackage{mathptmx}
\usepackage{anyfontsize}
%\usepackage{amsfonts}
\usepackage{wasysym}
\usepackage[margin=3cm]{geometry}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage{url}
\usepackage{hyperref}
\usepackage{environ}
\NewEnviron{shrunkdisplay}{%
\begin{equation*}
\scalebox{0.9}{$\BODY$}
\end{equation*}
}

\renewcommand{\erpictureFolder}[0]{../SharedPictures}

\setcounter{equation}{0}
\bibliographystyle{plain} % was hplain

\title{Literature Review}
\author{John Cartmell\footnote{john.w.cartmell gmail.com}\\ \normalsize{\textit{Ad Otium}}}
\begin{document}
\maketitle

\section{Categories with Attributes} 


\subsection{Cartmell 1978}

In my DPhil thesis (\cite{Cartmell78}) and also as published in \cite{Cartmell86}
I defined the notion of a \textit{contextual category} and proved that contextual
categories provide the\footnote{Note the use of the definite article -- meaning that they provide a semantics which is initial among semantics.} algebraic semantics of \textit{generalised algebraic theories}. \\

\noindent 
Contextual categories are, by design,  algebraically equivalent to the raw syntax of dependent type theories as described very nicely by Kapulkin-Lumsdaine (\cite{kapulkinlumsdaine2012}): \\

\begin{quote}
Rather than constructing an interpretation of the syntax directly, we may work via the intermediary notion of contextual categories, a class of algebraic objects abstracting the key structure carried by the syntax. The plain definition of a contextual category corresponds to the structural core of the syntax; further syntactic rules (logical constructors, etc.) correspond to extra algebraic structure that contextual categories may carry. Essentially, contextual categories provide a completely equivalent alternative to the syntactic presentation of type theory.

Why is this duplication of notions desirable? The trouble with the syntax is that it is mathematically tricky to handle. Any full presentation must account for (among other things) variable binding, capture-free substitution, and the possibility of multiple derivations of a judgement; and so a direct interpretation must deal with all of these, at the same time as tackling the details of the particular model in question. By passing to contextual categories, one deals with these subtleties and bureaucracy once and for all, and obtains a clear framework for subsequently constructing models. Conversely, why not work only with contextual categories, dispensing with syntax entirely?

The trouble on this side is that working with higher-order logical structure in contextual categories quickly becomes unreadable. The reader preferring to take contextual categories as primary may regard the syntax as essentially a notation for working within them: a powerful, flexible, and intuitive notation, but one whose validity requires non-trivial work to establish. (The situation is comparable to that of string diagrams, as used in monoidal and more elaborately structured categories.)
\end{quote}

\noindent
Also in \cite{Cartmell78} I defined the notion of a \textit{category with attributes}(
subsequently the term category with attributes was redefined and repurposed\footnote{Vladimir Voevodsy on the hott group commented 
that \textit {There is an inconsistency in terminology related to categories with attributes. 
What is called today a category with attributes is substantially different from what John Cartmell defined on pp. 172-173 of his thesis. The current concept is what one would get if one took only parts i - iv of Cartmell's definition and removed parts v and vi.}}\footnote{Peter LeFanu Lumsdaine replied that \textit{
I remember being surprised  by this ... I’ve never quite worked out who first repurposed “CwA” to have its current meaning...}}.
I defined these as a stepping stone between contextual categories with additional structure and certain kinds of hyperdoctrine to show how such hyperdoctrines could provide a semantics for
a particular variant of Martin-L\"of type theory.  

For the purposes of this note the crucial result is that contextual categories with the additional structure of existential quantification are equivalent to categories with attributes.
What this means is that \textit{certain} contextual 
categories (those with existential quantification) may be defined indirectly by way of a definition of the equivalent category with attributes. Two such, which have existential quantification, are:
\begin{enumerate}
\item the contextual category \cat{Fam} of sets, indexed families of sets, indexed families of families of sets and so on,
\item the contextual category of categories, indexed families of categories, indexed families of families and so on.
\end{enumerate}


\subsection{Erhart 1988}
In a 1988 paper Erhart \cite{erhard88} writes in the introduction that
\begin{quote}
This paper is based on a common work with P.L. Curien (\cite{curien89}\footnote{I have replaced the original ordinal with mine so that the reader can correlate with my references below}). We built up together some categorical structures for first order type dependency which we learned while finishing this work to be identical to Cartmell's contextual categories.
\end{quote}
\noindent
Well, I am confused by this for two distinct reasons:
\begin{enumerate} [(i)]
\item Though Erhard does show a relationship with my notion of \textit{category with attributes} 
the association between types, contexts and objects seems different to the association that I 
established. However I do not see existentially quantified types appearing in their paper and these are crucial in my use of categories with attributes so maybe I just don't understand the 
type syntax that they are using.
\item In my presentation, categories with attributes are representative of only certain particular generalised algebraic theories -- those with existential quantification.
\end{enumerate}
It is possible that these two discrepancies cancel out but I would like to understand how this is. It would be quite remarkable if they did -- it would mean that  categories with attributes specified for the very purpose of representing contextual categories with existential quantification inadvertently (from my point of view), and with a different mapping, represented all contextual categories (though there is a free contextual category with existential quantification generated by a contextual category I don't believe this to be an equivalence of categories.


\subsection{Curien 1989}
See \cite{curien89}.

\subsection{E.Moggi} 
I don't have Moggi's published paper \cite{moggi91} available to me but in notes having the same  title filed by Edinburgh University (\cite{moggi94}) I find:
\begin{quote}
Contextual categories, class of display maps, D-categories and categories with
attributes provide essentially equivalent accounts of dependent types. Unlike
the approach based on locally cartesian closed categories, they give a general
category-theoretic understanding of dependent types
\end{quote}

Moggi's definition of category with attributes is straightforward (an advantage here over Erhard's definitions) and from thsi definition it is clear that categories with attributes are not categories with attributes as I defined them and used them but are much closer to my contextaul categories
and, I believe,  are intended to be equivalent. However the problem that I have with his definitions
is  that by following them distinct dependent types may be mapped to the same extended contexts -- so we get an interpretation of types that are not equivalent forming contexts that are,  which is nonsense.  I ought to investigate whether his definition may be patched to yield a definition equivalent to my definition of contextual category.

\noindent
Differences are:
\begin{enumerate}
\item the functor G is not a monomorphism
\item the functor G is not surjective onto objects other than 1 and therefore there may be objects in C (which purportedly represent contexts) which do not correspond to extensions of contexts by types. 
\item also need the relation of dependency to be well-founded.
\item not even acyclic - is this true!!
\end{enumerate}

\noindent
If categories with attributes as defined by Moggi are to be the algebraic equivalents of GATs then we require the above conditions to hold. Then we are able to build a round trip
 from categories with 
attributes to contexual categories and back. 
\noindent From a moggi category with attributes define the contextual category as follows:
\begin{enumerate}
\item the underlying category is the base category C of the moggi catgeory with attributes.
\item the distinguised morphisms are those that are of the form $p_a$ for some object $a \in D(X)$.
\end{enumerate}


\subsection{Pitts}
See Pitts \cite{pitts2000}.

\subsection{Streicher}
Thomas Streicher PhD.(can we source this?).

\subsection{Jacobs 1993}
See \cite{jacobs93}.
Bart Jacobs refers back to Curien \cite{curien89}.

\subsection {Hofmann94}
\subsubsection{Hofmann 1994}
This paper refers to Hofmann's PhD definition of category with attributes.
(postcript downloaded from U. of Edinburgh). PhD probably covered in \cite{hofmann94}.
Hofmann's thesis refers to Bart Jacobs paper \cite{jacobs93}.
\noindent
Hofmann defines a \textit{syntactic category with attributes}.
This definition starts out as Moggi but deviates in that the  type substitution square is defined only to commute (i.e. not as a pullback)then comes the interesting twist - 
a sort of terms is introduced for each type and  the equivalent of Vladimir's s operator is defined. The sort of terms is then shown to be isomorphic 
to the sets of sections of the projection maps. Hoffman explains that
the additional sort is introduced just so that an equational definition and references Curien for this approach. I need to get hold of the Curruien paper.
\begin{quote}
The original definition of categories with attributes involves conditional equa­ 
tions, since certain diagrams are required to be pullbacks. This is a disadvantage if 
one wants to construct models internally in some type theory and check the model 
equations using normalisation. We thus describe a notion of model without con­ 
ditional equations inspired by Curien's categorical combinators [21,22,23]. This 
means that in addition to abstract contexts and families we introduce a third sort 
of abstract sections which are in one­one correspondence to sections (right inverses) 
to projections. Their separate introduction allows to quantify over them without 
using conditional equations. Ehrhard [30] and Ritter [92] achieves the same goal 
by introducing arbitrary morphisms between families as a new sort. For our pur­ 
poses this appears overly general and the price to pay is that a formal unit type 
must be introduced and that the equational theory becomes more complicated. 
Since a guideline in our description of the semantics of dependent types is to 
stay as close to the syntax as possible without losing the advantages which abstract 
semantics offers, we have chosen the name syntactic categories with attributes for 
the notion of model to be described here. 
\end{quote}

This definition probably ends up equivalent and definitely 
has same weaknesses from my point of view as the Moggi definition.

\noindent
In a discussion of the definition and one of these features which I consider a weakness Hofmann states that:
\begin{quote}
In particular, 
the map which assigns the corresponding context extension to a family need not 
be injective. This extra generality will be required in almost all the model con­ 
structions we shall study. 
\end{quote}

Hofmann also makes reference to categories with attributes of cartmell and moggi and category with types of pitts.
READ THIS 

\subsubsection{Hoffman 1995}
I have also found another Hofmann paper \cite{hofmann95} in Springer LNotes.

\section{Categories with Families}

The first definition of category with families is found in Dyber \cite{dybjer96}.

\subsection{Hofmann -- Syntax and Sematics of Dependent Types}
Refers to Dybjer \cite{dybjer96} for the definition of categories with families.
``Advantage of contextual categories is that they rule out semantic contexts which do not arise from the empty context by successive applications of comprehension.''

\section{Display Structures}
\subsection{Taylor}
\begin{verbatim}

EXAMPLE 8.3.3 Let C be a category with specified terminal object 1 and specified binary products. 
Then the class D of maps that are specified left projections, p0:DxX® D, forms a display structure. 
The resulting interpretation of the context $[x1:\typeX1,...,xn:\typeXn]$
is the left-associated product $((···(((1x \typeX1)x \typeX2)x \typeX3) x···)x \typeXn)$
which was defined in Remark 4.5.15.

PROOF: Given any map u:G® D, let the specified product projection p0:GxX® G serve as the choice of pullback. The ``dependent'' types in this example are in fact constant. []

PROPOSITION 8.3.4 Let C be a category with a class of displays. If all product projections and pullback diagonals are displays, then every map is (isomorphic to) a composite of displays, and C has all finite limits.

PROOF: Consider the graph of u:X x Y (Remark  5.1.7):

omitted diagram environment

Then the pullback of u against any |:Z® Y must exist. []

REMARK 8.3.5 The diagonal is the equality type,

{y0,y1|y0 = y1} x [y0,y1:Y],
so Proposition 8.3.4 says that the classifying category of a 
generalised algebraic theory has all finite limits 
(and all of its maps are isomorphic to displays) 
iff the theory has all equality type
\end{verbatim}

\section{Categories of Partial Maps}
\subsection{Robinson and Rosolini}
See \cite{robinson1988}.
Well written paper!
In later section of this give an overview of 2-category theory and say that they are concerned with 
PO-enriched categories which they say are usually called locally ordered categories.
They go on to give a definition of partial Cartesian category  adapted from a definition given by Curien and Obtulowicz
(\cite{curien1986}),
1986. They describe that partial maps to a terminal object correspond to domains of partial maps (boolean attributes in MY words) is decribed. INTERESTING!

Also described are p-categories, dominical categories. It is going to take some study to figure which of all these I want.
po-enriched category seems to be a good start. 

For what others called total maps authors decribe as weakly total. Defined in one case in 
terms of bottom morphisms in a pointed category. Need to spend some time sorting out what I need when i gave a cartesian product. Cartesian product enables (weakly) total maps to be defined in terms of composition and bottom morphisms. 
  
\section{C systems, B-systems}
\subsection{Voevodsky -- Subsystems and regular quotients of C-systems}
See \cite{Voevodsky16}.
\subsubsection{Summary}
\begin{enumerate}
\item Suggests contexual categories should be called  "contextual pre-categories". Suggests to use the name C-systems instead.
In a footnote states that the distinction between categories and pre-categories becaomes precise in univalent foundations where not all collections of objects are sets and references Ahrens, Kapulkin and Shulman:'Univalent Foundations and Rezk completion'.
\item Describes C0-systems as categories with certain additional structure. Describes C-systems as C0-systems plus an `s' operator satisfying three identities
((s1), (s2) and (s3) in my description.). Gives my contextual category description using pullbacks as an alternative definition of C-systems and proves the two equivalent (Proposition 2.4).
\item Notes that C-systems can be considered as models of essentially algebraic theories and that therefore all the results of Palmgren and Vickers: 'Partial Horn Logic and Cartesian Categories' can be applied. 
\item Notes that \textit{C0-systems and C-systems can not be described, in general, by generators and relations}(my italics). 
Thinking about this and in light of what he has to say a few paragraphs later then I don't think he means what he has says here.
I think he means to say that an arbitrary set of objects, morphisms and relations cannot be considered to be the generators and relations 
describing a C0-system or a C-system. Contrary to what he actually says every C0-system and every C-system can be described by generators and relations.
For example, he asks, what is the C0-system generated by $X \in Ob$. He goes on to say that no such universal object because we do not know what l(X) is. 
But this ony illustrates what I believe he means to say not what he actually said.
\item An aside: I might be wrong here but in my opinion Vladimir somethimes places an unnecessary definite article into a phrase 
and this makes it more difficult to understand. The phrase seems to mean more than it actually does. 
Examples of this: those uses of these objects that are invariant under *the* equivalences, a suitable category of *the* GATs. 
\item States that \textit{Presentations of C-systems in terms of GATs uses constructions that are substantially non-finitary -- a C-system given by finite sets of generators and relations can rarely be represented by a generalised algebraic theory with finitely many generating objects}. 
I don't belive what is said here is true ... or, more precisely,  I cannot find a way of reading this so as to make it true. I can examine carefully what is actually said and I can look at the remainder of the text for clues but ultimately I don't think it is clear what is meant. By my first reading and the one that I think is most obvious, 
the proposition \textit{a C-system given by finite sets of generators and relations can rarely be represented by a generalised algebraic theory with finitely many generating objects} is false. I would say, to the contrary, that a C-system presented by finite sets of generators and relations can be presented as the 
C-system of (roughly speaking) the equivalence classes of types and terms of a finite generalised algebraic theory. I would say that equality of objects and morphisms in the C-system presented by a given finite set of generators and relations is decidable iff type and term equality of the corresponding generalised algebraic theory is decidable because a decision procedure for one will yield a decision procedure for the other and vice-versa. Whatever \textit{substantially finitary} might mean I don't think either algebra or syntax has the upper hand and is somehow more substantially finitary then the other. 

\begin{enumerate}[(i)] 
\item What does substantially finitary mean as used here? Whatever it means we must suppose that the description `the C-system generated by a single object of length 1 and no other relations' is to be considered \textit{substantially finitary}. 
\item What are the substantially non-finatary constructions which are use in the presentations of C-systems in terms of GATs? 
\item What is meant by \textit{presentations of C-systems in terms of GATs}? Two possible readings of this. In the first reading
I might say that 'the C-system that I present is the C-system of equivalence classes of types and terms of such and such generalised algebraic theory'.
For example, the C-system corresponding to the generalised algebraic theory with a single sort symbol $X$ introduced by the rule `$X$ is a type' and no other sort symbols and no operator symbols. In the second reading I might say that 'the C-system I present is the initial model of such and such an extension of the generalised algebraic theory of contextual categories'. With the latter reading the theory involved is always infinite because the generalised algeebraic theory
of contextual categories has infinite numbers of sorts -- those mentioned by Vladimir as $Ob_n$ and $Mor_n,m$. With the second reading I might say that
`the C-system I present is the initial model of the theory of contextual categories extended by a constant $X$ of type $Ob$ '.
\end{enumerate}
\end{enumerate}

\bibliography{../SharedBibliography/temp/bibliography}

\end{document}