 \documentclass[10pt,a4paper]{article}
%\input{thepreamble.tex}
\newcommand{\sharedmacros}{../../SharedMacros}
\input{\sharedmacros/ccategories.macros}
\input{\sharedmacros/ermacros}
\input{\sharedmacros/erdiagram}
\input{\sharedmacros/theorems.macros}
\input{\sharedmacros/general.macros}
\usepackage{mathptmx}  % This changes font to roman
\usepackage{anyfontsize}
\usepackage{mathtools}  % why have we got this?
\usepackage{alltt}    
\usepackage{mnsymbol} %used for rightpitchfork
\usepackage{cmll}
%\usepackage{yhmath}  % used for better version of \widehat
\usepackage{ulem}
\renewcommand{\ttdefault}{txtt}
\usepackage[left=1.5cm, right=4cm, marginparwidth=3cm, top=2cm, bottom=2.0cm]{geometry}
\usepackage{framed}
\usepackage[font=small]{caption}
\setlength{\captionmargin}{2cm}


\theoremstyle{remark}
\newtheorem*{lemma*}{Lemma}


\renewcommand{\term}[1]{\textit{#1}}  %SIMPLE UNINDEXED VERSION

%ENDCOPY

\begin{document}
\title{Range Categories and the Theory of Data}

% abstract here for ams

\author{John Cartmell}

\date{}

\maketitle

\begin{center}
DRAFTING December 2023
\end{center}

\newcommand{\binarysourcediag}[5]{
$
\begin{array}{c p{0.5cm} c  }
             &&   \Rnode{b}{#2} \\[0.01cm]
\Rnode{a}{#1} &&                \\[0.01cm] 
             &&   \Rnode{c}{#3} 
\end{array} 
\begin{arrows}
\ncarr{a}{b}
\alabel{#4}
\ncarr{a}{c}
\blabel{#5}
\end{arrows}
$  
}
\newcommand{\fgsourcediag}{\binarysourcediag{a}{b}{c}{f}{g}}

\newcommand{\paralleldiag}[4]
{
 $
\rule[-0.3cm]{0pt}{0.9cm} %to add vertical space of diagram -- based on lowering diagram 0.3cm and heght 0.9cm
                            % change thickness from 0pt to 1 pt to debug
\begin{array}{c p{0.5cm} c  }
 \Rnode{a}{#1}     &&   \Rnode{b}{#2}
\end{array} 
\begin{arrows}
\ncarc[nodesep=2pt,arcangle=10,offset=2pt]{->}{a}{b}
\alabel{#3}
\ncarc[nodesep=2pt,arcangle=-10,offset=-2pt]{->}{a}{b}
\blabel{#4}
\end{arrows}
$  
}

\newcommand{\sequentialdiag}[5]
{$
\begin{array}{c p{0.5cm} c p{0.5cm} c }
 \Rnode{a}{#1}     &&   \Rnode{b}{#2} && \Rnode{c}{#3}
\end{array} 
\begin{arrows}
\ncarr{a}{b}
\alabel{#4}
\ncarr{b}{c}
\alabel{#5}
\end{arrows}
$}

\newcommand{\alternatediag}[5]
{$
\begin{array}{c p{0.5cm} c p{0.5cm} c }
 \Rnode{a}{#1}     &&   \Rnode{b}{#2} && \Rnode{c}{#3}
\end{array} 
\begin{arrows}
\ncarr{a}{b}
\alabel{#4}
\ncarr{c}{b}
\blabel{#5}
\end{arrows}
$}

\newcommand{\alternatingchainfg}[5]
{
\begin{array}{c p{0.5cm} c p{0.5cm} c p{0.5cm} c p{0.5cm} c  c p{0.5cm} c p{0.5cm} c}
 \Rnode{a1}{#1_1} &&  \Rnode{b1}{#2_1} && \Rnode{a2}{#1_2} && \Rnode{b2}{#2_2} 
                  && \Rnode{dots}{\hspace{0.3cm}...} 
                  &\Rnode{an}{#1_#5}&&\Rnode{bn}{#2_#5}&&\Rnode{an1}{#1_{#5+1}}
\end{array} 
\begin{arrows}
\ncarr{a1}{b1}
\alabel{#3_1}
\ncarr{a2}{b1}
\blabel{#4_1}
\ncarr{a2}{b2}
\alabel{#3_2}
\ncarr{dots}{b2}
\blabel{#4_2}
\ncarr{an}{bn}
\alabel{#3_#5}
\ncarr{an1}{bn}
\blabel{#4_#5}
\end{arrows}
}

\newcommand{\alternatingchainff}[5]
{
\begin{array}{c p{0.5cm} c p{0.5cm} c p{0.5cm} c p{0.5cm} c  c p{0.5cm} c }
 \Rnode{a1}{#1_1} &&  \Rnode{b1}{#2_1} && \Rnode{a2}{#1_2} && \Rnode{b2}{#2_2} 
                  && \Rnode{dots}{\hspace{0.3cm}...} 
                  &\Rnode{an}{#1_#5}&&\Rnode{bn}{#2_#5}
\end{array} 
\begin{arrows}
\ncarr{a1}{b1}
\alabel{#3_1}
\ncarr{a2}{b1}
\blabel{#4_1}
\ncarr{a2}{b2}
\alabel{#3_2}
\ncarr{dots}{b2}
\blabel{#4_2}
\ncarr{an}{bn}
\alabel{#3_#5}
\end{arrows}
}

\newcommand{\alternatingchaingg}[5]
{
\begin{array}{c p{0.5cm} c p{0.5cm} c p{0.5cm} c p{0.5cm} c  c p{0.5cm} c p{0.5cm} c}
 \Rnode{b0}{#2_0} && \Rnode{a1}{#1_1} &&  \Rnode{b1}{#2_1} && \Rnode{a2}{#1_2} 
                  && \Rnode{dots}{\hspace{0.3cm}...} 
                  &\Rnode{an}{#1_#5}&&\Rnode{bn}{#2_#5}&&\Rnode{an1}{#1_{#5+1}}
\end{array} 
\begin{arrows}
\ncarr{a1}{b0}
\blabel{#4_0}
\ncarr{a1}{b1}
\alabel{#3_1}
\ncarr{a2}{b1}
\blabel{#4_1}
\ncarr{a2}{dots}
\alabel{#3_2}
%\ncarr{dots}{b2}
%\blabel{#4_2}
\ncarr{an}{bn}
\alabel{#3_#5}
\ncarr{an1}{bn}
\blabel{#4_#5}
\end{arrows}
}


\newcommand{\alternatingchaingf}[5]
{
\begin{array}{c p{0.5cm} c p{0.5cm} c p{0.5cm} c p{0.5cm} c c p{0.5cm} c }
\Rnode{b0}{#2_0} && \Rnode{a1}{#1_1} && \Rnode{b1}{#2_1} && \Rnode{a2}{#1_2} 
                  && \Rnode{dots}{\hspace{0.3cm}...} 
                  &\Rnode{an}{#1_#5}&&\Rnode{bn}{#2_#5}
\end{array} 
\begin{arrows}
\ncarr{a1}{b0}
\blabel{#4_0}
\ncarr{a1}{b1}
\alabel{#3_1}
\ncarr{a2}{b1}
\blabel{#4_1}
\ncarr{a2}{dots}
\alabel{#3_2}
%\ncarr{dots}{b2}
%\blabel{#4_2}
\ncarr{an}{bn}
\alabel{#3_#5}
\end{arrows}
}



\iffalse
\newcommand{\fnsourceqnsource}
{
$
\begin{array}{c p{0.25cm} c  p{0.25cm} c }
             &&   \Rnode{b1}{b_1} &&              \\[0.4cm]
\Rnode{a}{a} &&                   && \Rnode{c}{c} \\[0.4cm]
             &&   \Rnode{bn}{b_n} &&              
\end{array} 
\begin{arrows}
\ncarr{a}{b1}
\alabel{f_1}
\ncarr{c}{b1}
\blabel{q_1} 
\ncarr{a}{bn}
\blabel{f_n}
\ncarr{c}{bn}
\alabel{q_n}
\end{arrows}
$   
}
\fi

\section{Restriction Categories}

From Cockett \& Lack, a restriction category is a category \catcw with an additiona operator (the restriction operator) such that
if $f: a \morph b$ in \catcw then
$$\bar{f}: a \morph a$$ 
and satisfying

R.1 For $f:a \morph b$ in \catcw $$\bar{f} \circ f =f$$.

R.2. If \fgsourcediag in \catcw then
$$\bar{g} \circ \bar{f}=\bar{f} \circ \bar{g}.$$

R.3. If \fgsourcediag in \catcw then
$$\overline{\bar{f} \circ g} = \bar{f} \circ \bar{g}$$.

R.4. If \sequentialdiag{a}{b}{c}{f}{g} in \catcw then
$$f \circ \bar{g} = \overline{f \circ g} \circ f$$.

\section{Range Categories}

A \term{range category} is a retriction category with an additional operator as follows
if $f: a \morph b$ in  \catcw then
$$\hat{f}: b \morph b$$
satisfying

RR.1 For $f:a \morph b$ in \catcw $$\bar{\hat{f}} = \hat{f}.$$

RR.2 For $f:a \morph b$ in \catcw $$f \circ \hat{f} = f.$$

RR.3. If \sequentialdiag{a}{b}{c}{f}{g} in \catcw then
$$\widehat{f \circ \bar{g}} = \hat{f} \circ \bar{g}.$$

RR.4. If \sequentialdiag{a}{b}{c}{f}{g} in \catcw then
$$\widehat{(hat({f}) \circ g)} = \widehat{f \circ g}.$$

\begin{lemma}
\label{restrictioncatlemma}
In any restriction category \catc
\begin{enumerate} [(i)]
\item For $f:a \morph b$ in \catcw,
$$\bar{\bar{f}}=\bar{f}$$
\item If \sequentialdiag{a}{b}{c}{k}{j} in \catcw 
and $k$ is total i.e. $\bar{k}=id_a$
and if $k \circ \bar{j}$ is total i.e. $\overline{k \circ \bar{j}}=id_a$
then
$$k \circ \bar{j} = k.$$
\item If \sequentialdiag{a}{b}{c}{k}{j} in \catcw 
and $k$ is total i.e. $\bar{k}=id_a$ then
$$\overline{k \circ j} = \overline{k \circ \bar{j}}.$$
\end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate} [(i)]
\item Use R3 with $g$ being $id_b$.
\item 
\begin{align*}
k \circ \bar{j} &= k \circ \bar{\bar{j}}              && \mbox{from (i),}\\
                &= \overline{k \circ \bar{j}} \circ k && \mbox{by R.4,}   \\
                &= k                                  && \mbox{from assumption $k \circ \bar{j}$ is total.}
\end{align*}
\item 
\begin{align*}
\overline{k \circ j} &= \overline{k \circ j} \circ \bar{k}    && \mbox{from assumption that $k$ total,}\\
                     &= \overline{\overline{k \circ j} \circ k} && \mbox{by R.3,} \\
                     &= \overline{k \circ \bar{j}}              && \mbox{by R.4.}
\end{align*}
\end{enumerate}
\end{proof}

Let the category \FinPar be the range catgory of finite sets and partial functions.

\subsection{Split assumption}
\begin{definition}
In a range category \catcw then a morphism $f:a \morph b$ is split iff
there exists $f_s:b \morph a$ 
such that
\begin{equation*} 
f_s \circ f = \hat{f} = \bar{f_s}
\end{equation*}
and
\begin{equation*} 
f \circ f_s = \hat{f_s} = \bar{f}
\end{equation*}
\end{definition}

\begin{lemma}
\label{jsbarj}
In a range category \catcw if $f_s$ is a splitting of a morphism $f$ 
then $f_s \circ \bar{f} = f_s.$
\end{lemma}
\begin{proof}
\begin{align*}
f_s \circ \bar{f} 
      &= f_s \circ (f \circ f_s) 
             && \mbox{because $\bar{f}=f \circ f_s$, by definition of splitting, } \\
      &= \bar{f_s} \circ f_s     
             && \mbox{because $f \circ f_s = \bar{f_s}$, from same definition,}    \\
             && 
      &= f_s   &&\mbox{by R.1, as required.}
\end{align*}
\end{proof}

\begin{lemma}
If $f:a \morph a$ in a range category \catcw then for any morphism 
$f: a \morph b$
\begin{enumerate}[(i)] 
\item $\hat{\hat{f}}=\hat{f}$,
\item $\hat{f} \comp \hat{f} = \hat{f}$,
\item $\hat{\bar{f}}=\bar{f}$,
\item $\bar{\hat{f}}=\hat{f}$,
\item $\bar{f}$ is a splitting of itself,
\item $\hat{f}$ is a splitting of itself.
\end{enumerate}
\end{lemma}
\begin{proof}
Straightforward.
\end{proof}

\begin{definition}
A range category with coherent splittings is a range category \catcw along with for every
morphism $f:a \morph b$ a splitting $f_s:b \morph a$ such that if
\sequentialdiag{a}{b}{c}{f}{g} in \catcw then $(f \circ g)_s=g_s \circ f_s$.
\end{definition}

\begin{lemma}
\label{splittingsequencepair}
If \catcw is a range category with coherent splittings then if
\sequentialdiag{a}{b}{c}{f}{g} in \catcw
 then 
 \begin{enumerate}[(i)]
 \item $\overline{f \circ g}=f \circ \bar{g} \circ f_s$,
 \item $\widehat{f \circ g}=g_s\circ \hat{f} \circ g$
\end{enumerate}
 \end{lemma}
\begin{proof}
\end{proof}

\begin{lemma}
\label{splittingalternatepair}
If \catcw is a range category with coherent splittings then if
\alternatediag{a}{b}{c}{f}{g} in \catcw
 then 
 \begin{enumerate}[(i)]
 \item $\overline{f \circ g_s}=f \circ \hat{g} \circ f_s$,
 \item $\widehat{f \circ g_s} = g\circ \hat{f} \circ g_s$
\end{enumerate}
 \end{lemma}
\begin{proof}
Use lemma \ref{splittingsequencepair}.
\end{proof}

\newcommand{\range}[1]{\hat{}(#1)}
\newcommand{\psplit}[1]{#1^{-1}}
\begin{lemma}
\label{splittingalternatesequences}
If \catcw is a range category with coherent splittings then 
\begin{enumerate}[(i)]
\item if $\alternatingchainfg{a}{b}{f}{g}{n}$ in \catcw then
\begin{align*}
\overline{f_1 \circ \psplit{g_1} \circ f_2 \circ \psplit{g_2} 
... \circ f_n \circ \psplit{g_n} } 
= f_1 \circ \psplit{g_1}  ... f_n  \circ \psplit{g_n}   
\circ g_n \circ \psplit{f_n} ... \circ g_1 \circ \psplit{f_1}\\
\range{f_1 \circ \psplit{g_1} \circ f_2 \circ \psplit{g_2} 
... \circ f_n \circ \psplit{g_n} } 
= g_n \circ \psplit{f_n} ... g_1 \circ \psplit{f_1} \circ f_1 \circ \psplit{g_1}
... \circ f_n \circ \psplit{g_n}
\end{align*}
\item if $\alternatingchainff{a}{b}{f}{g}{n}$ in \catcw then 
\begin{align*}
\overline{f_1 \circ \psplit{g_1} \circ f_2 \circ \psplit{g_2} 
... \circ f_n  } 
= f_1 \circ \psplit{g_1}  ... f_n  \circ \psplit{f_n} ... \circ g_1 \circ \psplit{f_1}\\
\range{f_1 \circ \psplit{g_1} \circ f_2 \circ \psplit{g_2} 
... \circ f_n  } 
= \psplit{f_n} ... g_1 \circ \psplit{f_1} \circ f_1 
... \circ f_n \circ \psplit{g_1}
\end{align*}
\item if $\alternatingchaingg{a}{b}{f}{g}{n}$ in \catcw then 
$$etc.$$
\item if $\alternatingchaingf{a}{b}{f}{g}{n}$ in \catcw then
$$etc.$$
\end{enumerate}
\end{lemma}
\begin{proof}
Use lemma \ref{splittingalternatepair}.
\end{proof}

\newpage
\section{Construction of a Category with Coherent Splittings from a Range Category}
\begin{definition}
If \catcw is a range category then specify a category \catcaltw as follows.

The objects of \catcaltw are exactly the objects of \catc.

The morphisms of \catcaltw are certain alternating sequences of morphisms of \catcw.

Each sequence is either co- or contra-headed and either co- or contr-tailed as follows.
\begin{enumerate}[(i)]
\item If $f: a \morph b$ in \catcw then the singleton sequence with element $f$ is a morphism $f: a \morph b$ in \catc. This sequence is defined to be co-headed and co-tailed.
\item If $g: b \morph a$ in \catcw and $g$ is not an identity morphism and neither is 
$g=\bar{g}$ nor $g=\hat{g}$ 
 then the singleton sequence with element $g$ is a morphism $g: a \morph b$ in \catc. This sequence is defined to be contra-headed and contra-tailed.
\item If $f: a \morph b$ in \catc, if  $S:b \morph c$ is a contra-header sequence such that $S:b \morph c$ in \catcaltw then the sequence $(f,S): a \morph c$ is a morphism 
of \catcaltw  provided that
       \begin{enumerate}
       \item $f$ is not the identity morphism 
       \item $f$ is not $\hat{g}$ where $g$ is the head of $S$.
       \end{enumerate}
This sequence is defined to be co-headed with head $f$.
\item If $g: b \morph a$ in \catc, if  $S:b \morph c$ is a co-headed sequence such that $S:b \morph c$ in \catcaltw then the sequence $(g,S): a \morph c$ is a morphism 
of \catcaltw provided that
       \begin{enumerate}
       \item $g$ is not an identity morphism and neither is $g=\bar{g}$ nor $g=\hat{g}$
       \item $f$ is not $\bar{g}$ where $f$ is the head of $S$.
       \end{enumerate} 
This sequence is defined to be contra-headed with head $g$.
\end{enumerate}

The identity morphisms in \catcaltw are just the identity morphisms of \catcw considered as singleton co-headed sequences.

Composition in \catcaltw is defined as follows
\begin{enumerate}
\item If $S,f$ is a co-tailed sequence $S,f:a \morph b$ in \catcaltw then
       \begin{enumerate}
       \item if $f',S'$ is a co-headed sequence $f',S':b \morph c$ in \catcaltw
           then 
       \begin{equation*}
              (S,f) \circ (f',S') \mbox{ is defined as } 
          \begin{cases*}
              S \circ S' & if $f \circ f'$ is an identity morphism     \\
              S \circ S' & if S' headed by g and $f \circ f'=\hat{g}$  \\
              S \circ S' & if S has tail g and $f \circ f'=\bar{g}$    \\ 
              (S,f \circ f',S') & otherwise
          \end{cases*}
       \end{equation*}
       \item if $g,S'$ is a contra-headed sequence $g,S':b \morph c$ in \catcaltw
              \begin{equation*}
               (S,f) \circ (g,S') \mbox{ is defined as } 
                 \begin{cases*}
                     (S,S')         & if $f=g$             \\
                     S \circ (g,S') & if $f = \hat{g}$     \\ 
                     (S,f,g,S')     & otherwise
                 \end{cases*}
              \end{equation*}
       \end{enumerate}

\end{enumerate}



\end{definition}

\section{Construction of Restriction Functors}

\subsection{$HomP$ is a Restriction Functor}
If \catcw is a restriction functor then
let $HomP_\catc(a,-): \catc \morph \SetP$ be the  functor
defined as follows. $HomP(a,x)=Hom(a,x)$.
If $j: x \morph y$ then partial function $HomP(a,j): HomP(a,x) \morph HomP(a,y)$
is defined as follows:
\begin{align*}
HomP(a,j)(k) = k \circ j &\mbox{ if $k \circ \bar{j} = k$} \\
                         & \mbox{ is undefined otherwise}
\end{align*}
\begin{lemma}
If \catcw is a restriction catgory and $a$ is an object of \catcw 
then the functor $HomP_\catc(a,-)$ preserves the restriction operator i.e.
for all $j:x \morph y$ in \catc
$$HomP_\catc(a,\bar{j})=\overline{HomP_\catc(a,j)}$$
\end{lemma}
\begin{proof}
Both $HomP_\catc(a,\bar{j})$ and $\overline{HomP_\catc(a,j)}$ are less than id function on the set $Hom(a,x)$. 
The former because by definition of $HomP$ 
if $HomP_\catc(a,\bar{j})(k)$ is defined then $HomP_\catc(a,\bar{j})(k)=k$.
The latter by definition of $\bar{f}$, for any function $f$.

It suffices to show that $HomP_\catc(a,\bar{j})$ is defined iff and only if
$HomP_\catc(a,j)$ is defined which we can do as follows.
 Let $k \in Hom(a,x)$. $HomP_\catc(a,\bar{j})(k)$
is defined iff $k \circ \bar{\bar{j}}$ i.e. $k \circ \bar{j}$ is defined
iff $HomP_\catc(a,j)(k)$ is defined. 
\end{proof}

\subsection{$HomT$ Functor is a Restriction Functor}
If \catcw is a restriction functor then
let $HomT_\catc(a,-): \catc \morph \SetP$ be the sub functor of $Hom_\catc(a,-)$
where $HomT_\catc(a,x)$ is the set of total morphisms i.e.
$$HomT(a,x)=\setsuchthat{k \in Hom(a,x)}{\bar{k}=id_a}.$$
and where when $j:x \morph x' in \catc$ then $HomT(a,j)(k)=k \circ j$ provided the rhs is total or is undefined otherwise.  
 
\begin{lemma}
If \catcw is a restriction catgory and $a$ is an object of \catcw 
then the functor $HomT_\catc(a,-)$ preserves the restriction operator i.e.
for all $j:x \morph y$ in \catc
$$HomT_\catc(a,\bar{j})=\overline{HomT_\catc(a,j)}$$
\end{lemma}
\begin{proof}
Both $HomT_\catc(a,\bar{j})$ and $\overline{HomT_\catc(a,j)}$ are less than id function on the set $Hom(a,x)$. 
The former because of lemma \ref{restrictioncatlemma} (ii).
The latter by definition.

Therefore it suffices to show that each are defined on exactly the same morphisms.

LHS is defined on exactly those $k:a \morph x$ such that $k \circ \bar{j}$ is total.
and LHS less than or equal id becayuse
WE CAN SHOW THAT if $k \circ \bar{j}$ is total then $k \circ \bar{j}=k$

RHS is defined on exactly those $k:a \morph x$ such that $k \circ j$ is total
and is less than or equal id. Therefore required to show that
$k \circ j$ is total iff $k \circ \bar{j}$ is total. and this is the case 
because by 
lemma \ref{restrictioncatlemma} (iii), $\overline{k \circ j}=\overline{k \circ \bar{j}}$ 
\end{proof}
\subsection{Coproducts of Set valued Range Functors}
\subsection{Quotients of Set valued Restriction Functors}
Suppose \catcw is a restriction category and that $F: \catc \morph \SetP$ is a restriction functor into the restriction category of sets and partial functions. Suppose that for every object $x$ of \catcw there is an equivalence relationship $\sim_x$ defined on the set $F(x)$. Suppose that this equivalence relationship has the following property(s):
For every morphism $j:x \morph y$ in \catcw, for all elements $k1,k2 \in F(x)$,
\begin{enumerate}
\item $F(j)(k_1)$ is defined iff $F(j)(k_2)$ is defined, 
\item $k_1 \sim_x k_2 \mbox{ and } F(j)(k_1) \mbox{ is defined } \implies F(j)(k_1) \sim_y F(j)(k_2).$
\end{enumerate}
Define functor $F_\sim: \catcw \morph \SetP$ by defining
\begin{align*}
&x             &&\mapsto F(x)/\sim_x 
                & \mbox{i.e. the set of equivalence classes of $F(x)$ wrt equivalence relation $\sim$}\\
&j: x \morph y &&\mapsto F(j)/\sim
\end{align*} 
where $F(j)/\sim$ is defined by
\begin{align*}
[k] \mapsto [F(j)(k)] && \mbox{provided $F(j)(k)$ is defined, is undefined otherwise.}
\end{align*}
which is well-defined because of the assumption made above.

It is easy to see that $F_\sim$ is a functor (i.e. respects composition $\circ$ and identity morphisms.)

\begin{lemma}
The functor $F_\sim$ is a restriction functor.
\end{lemma}
\begin{proof}
We need to show that  $F_\sim$ respects the restriction operator
i.e. we need to show that for any morphism $j:x \morph y$ in \catcw, $F_\sim(\bar{j})= \overline{F_\sim(j)}$.

The LHS, $F_\sim(\bar{j})$, is defined to  be the partial function
$$ [k] \mapsto [F(\bar{j})(k)] \mbox{ providing $F(\bar{j})(k)$ is defined, is undefind otherwise,}$$
in otherwords, since  $F$ assumed to be a restriction functor it can be defined as mapping
$$ [k] \mapsto [\overline{F(j)}(k)] \mbox{providing $\overline{F(j)}(k)$ is defined, is undefind otherwise,}$$
and this in turn, because of the definition of restriction in \SetP, means that it maps
$$ [k] \mapsto [k] \mbox{providing $F(j)(k)$ is defined, is undefined otherwise.}$$

Meanwhile the RHS, $\overline{F_\sim(j)}$  is defined as the partial function that 
maps $[k]$  to $[k]$  provided  $F_\sim(j)(k)$ is defined, and to be undefined otherwise.

Because,
from the definition of $F_\sim$, $F_\sim(j)([k])$ is defined iff $F(j)(k)$ is defined then
we have shown that LHS and RHS are identical functions, as required.
\end{proof}

\section{Construction of Range Functors}

\subsection{Split assumption and $HomT$ as a Range Functor}

\begin{lemma}
If \catcw is a range category in which every morphism is split then
for every object $a$ of \catcw the functor $HomT_\catcw: \catcw \morph \SetP$ preserves the range operator.
\end{lemma}
\begin{proof}
If $j: x \morph y$ in \catcw then we are required to prove that
$$\widehat{HomT_\catcw(a,j)}=HomT_\catcw(a,\hat{j})$$
and for this it suffices to show that the range of $HomT_\catcw(a,j)$ i.e. this set
$$\setsuchthat{ k \circ j}{k: a \morph x, k \circ j \mbox{total}}$$ 
is identical to the range of $HomT_\catcw(a,\hat{j})$ i.e. this set
$$\setsuchthat{k' \circ \hat{j}}{k':a \morph y,  k' \circ \hat{j} \mbox{ total}}.$$  

The first set is included in the second because any total morphism of the form $k \circ j$,  is of the form $k' \circ \hat{j}$,  because we can take $k'$ to be $k \circ j$ 
because then $k' \circ \hat{j} = (k \circ j) \circ \hat{j}= k \circ(j \circ \hat{j})= k \circ j$. 

Let $j_s : y \morph x$ be the splitting of $j$ i.e. be such that $j_s \circ j = \hat{j}$.

The second set is included in the first because if a morphism is of the form $k' \circ \hat{j}$ then it of the form $k \circ j$ because $k$ can be taken to be $k' \circ j_s$ for then we have $k' \circ \hat{j} = k' \circ j_s \circ j = k \circ j$. 
 \end{proof}

 \subsection{Split assumption and $HomP$ as a Range Functor}

\begin{lemma}
If \catcw is a range category in which every morphism is split then
for every object $a$ of \catcw the functor $HomP_\catcw: \catcw \morph \SetP$ preserves the range operator.
\end{lemma}
\begin{proof}
If $j: x \morph y$ in \catcw then we are required to prove that
$$\widehat{HomP_\catcw(a,j)}=HomP_\catcw(a,\hat{j})$$
and for this it suffices to show that the range of $HomP_\catcw(a,j)$ 
i.e. this set
$$\setsuchthat{ k \circ j}{k: a \morph x \mbox{ and } k \circ \bar{j} = k }$$ 
is identical to the range of $HomP_\catcw(a,\hat{j})$ 
i.e. this set
$$\setsuchthat{k' \circ \hat{j}}{k':a \morph y \mbox{ and } k' \circ \bar{\hat{j}} = k' }.$$   
i.e. this set
$$\setsuchthat{k' \circ \hat{j}}{k':a \morph y \mbox{ and } k' \circ \hat{j} = k' }.$$

The first set is included in the second because for any morphism of the form $k \circ j$, 
where $k \circ \bar{j}=k$, 
is of the form $k' \circ \hat{j}$, for $k' \circ \hat{j} = k'$, 
because we can take $k'$ to be $k \circ j$ 
because then 
$k' \circ \hat{j} = (k \circ j) \circ \hat{j}= k \circ(j \circ \hat{j})= k \circ j = k'$,
as required. 

Let $j_s : y \morph x$ be the splitting of $j$ i.e. be such that $j_s \circ j = \hat{j}$.

The second set is included in the first because if a morphism is of the form $k' \circ \hat{j}$, where $k' \circ \hat{j} = k'$, then it of the form $k \circ j$, where $k \circ \bar{j}=k$, because $k$ can be taken to be $k' \circ j_s$ for then we have 
$k' \circ \hat{j} = k' \circ j_s \circ j = k \circ j$
and we have $k \circ \bar{j}= k' \circ j_s \circ \bar{j} =  k' \circ j_s = k$ using lemma
\ref{jsbarj}, as required. 
 \end{proof}

\begin{corollary}
If \catcw is a range category in which every morphism is split then
for every object $a$ of \catcw the functor $HomT_\catcw(a,-): \catcw \morph \SetP$ is a range functor.
\end{corollary}

\subsection{Coproducts of Set valued Range Functors}
\newpage
\subsection{Quotients of Set valued Range Functors}
\begin{lemma}
The functor $F_\sim$ is a range functor.
\end{lemma}
\begin{proof}
We need to show that  $F_\sim$ respects the range operator
i.e. we need to show that
 for any morphism $j:x \morph y$ in \catcw, $F_\sim(\hat{j})= \widehat{F_\sim(j)}$.

The LHS, $F_\sim(\hat{j})$, is defined to  be 
the partial function which maps $[k']$, where $k' \in F(y)$, as follows
$$ [k'] \mapsto [F(\hat{j})(k')] \mbox{ providing $F(\hat{j})(k')$ is defined, and is undefined otherwise.}$$
In otherwords, since  $F$ assumed to be a range functor it can be defined as mapping
$$ [k'] \mapsto [\widehat{F(j)}(k')] \mbox{ providing $\widehat{F(j)}(k')$ is defined and being undefined otherwise,}$$
and this in turn, because of the definition of ranges in \SetP, means that it maps
$$ [k'] \mapsto [k'] \mbox{providing there exists $k \in F(x)$ such that $k'=F(j)(k)$ and being undefined otherwise.}$$

Meanwhile the RHS, $\widehat{F_\sim(j)}$  is defined as the partial function that 
maps $[k']$  to $[k']$  provided  there exists $k \in F(x)$ such that
$F_\sim(j)([k])=[k']$ to be undefined otherwise.

Because,
from the definition of $F_\sim$, $F_\sim(j)([k])= [F(j)(k)]$ and therefore 
we have shown that LHS and RHS are identical functions, as required.
\end{proof}
\newpage

\section{MToD Definitions}

\begin{definition}
If $\catc$ is a category and $\reqtc$ is a set of instances and if \fgsourcediag
in $\catc$ then there is a  \term{functional dependency} of $g$ on $f$ with respect to $\reqtc$ iff
there is a family of functions $H_D)_{D \in \reqtc}$ such that 
in each instance $D$, $H_D$ is a unique function $H_D: D(b) \morph D(c)$, such that $D(f) \circ H_D = D(g)$ and $\overline{H_D}=\widehat{D(f)}$.
If there is such a functional dependency then we say that $\fundep{H}{f}{g}$ in $\catc$ with respect to $\reqtc$.
\end{definition}

\begin{lemma}
\label{fdrangesublemma}
If $\fundep{H}{f}{g}$ is a functional dependency then
$$\widehat{H_D}=\widehat{D(g)}.$$
\end{lemma}
\begin{proof}
\begin{align*}
\widehat{D(g)} &= \widehat{D(f) \circ H_D} \\
               &= \widehat{hat({D{f})} \circ H_D} &&\mbox{by RR.4}\\               
               &= \widehat{\overline{H_D} \circ H_D} &&\mbox{by defn of functional dependency}\\
               &= \widehat{ H_D} &&\mbox{by R.1}\\
\end{align*}
\end{proof}

\begin{definition}
If $\catc$ is a range category and $\reqtc$ is a set of instances, if
\fgsourcediag
in $\catc$ 
and if there is a functional dependency $\fundep{H}{f}{g}$ then say that 
the functional dependency $H$ is \term{represented} in $\catc$ 
iff there exists a morphism $h:b \morph c$ in $\catc$ such that for each instance $D \in \reqtc$, $D(h)=H_D$. 
\end{definition}

\begin{oldtt}
Question: Do we wish to add to the above definition that $\bar{h}=\hat{f}$????
Maybe we don't need to because if $f$ didn't have this property then
there would be a morphism, $\hat{f}\circ h$, that would have the property.
\end{oldtt}

\input{fd_representation_lemma}

\bibliographystyle{alpha} 
\bibliography{../SharedBibliography/temp/bibliography}
\end{document}
