
\documentclass[10pt,a4paper]{scrartcl}
%\usepackage[margin=3cm]{geometry} %see below where I make room for margin notes

\usepackage{url}
\usepackage{hyperref}
\usepackage{environ}
\NewEnviron{shrunkdisplay}{%
\begin{equation*}
\scalebox{0.9}{$\BODY$}
\end{equation*}
}

\input{../SharedMacros/theorems.macros}
\input{../SharedMacros/ccategories.macros}
\input{../SharedMacros/gats.macros}
\input{../SharedMacros/metagat.macros}
\input{../SharedMacros/ermacros}
\input{../SharedMacros/erdiagram}
\input{../SharedMacros/indexedsets.macros}
\usepackage{ulem} % for \sout for use by \oldt macro
\input{../SharedMacros/general.macros}

\renewcommand{\erpictureFolder}[0]{../SharedPictures}
% support margin notes
\usepackage[left=1.5cm, right=4cm, marginparwidth=3cm, top=2cm, bottom=2.0cm]{geometry}
\setcounter{equation}{0}
\bibliographystyle{plain} % was hplain

\title{ Theories Part}
\author{John Cartmell\footnote{john.w.cartmell gmail.com}\\ \normalsize{\textit{Ad Otium}}}
\subtitle{Status: \normalsize{WORK IN PROGRESS}}
\begin{document}
\maketitle

\section{New November 2020 Material}

\newcommand{\sEquiv}[2]{#1 \equiv #2}
\newcommand{\wEquivOne}[2]{#1\ we_1\ #2}
\newcommand{\wEquivTwo}[2]{#1\ we_2\ #2}


Write $\sEquiv{\encxDelta{n}}{\encyOmega{n}}$ to say that $\encxDelta{n}$ is a context and $\encyOmega{n}$ is a context and $\encxDelta{n}$ is strongly equivalent to $\encyOmega{n}$.

Write $\wEquivOne{\encxDelta{n}}{\encyOmega{n}}$ to say that $\encxDelta{n}$ is weakly equivalent in sense (1) to $\encyOmega{n}$ i.e. as defined in August 2018.

Write $\wEquivTwo{\encxDelta{n}}{\encyOmega{n}}$ to say that $\encxDelta{n}$ is weakly equivalent in sense (2) to $\encyOmega{n}$.

\begin{definition}
If $\Delta=\encxDelta{n}$ and $\Omega = \encyOmega{n}$ are contexts then $\wEquivTwo{\Delta}{\Omega}$ iff there exists contexts $\Delta' = \encxDeltap{n}$ and $\Omega' = \encyOmegap{n}$ such that
$\sEquiv{\Delta}{\Delta'}$ and $\sEquiv{\Omega}{\Omega}$ and $\wEquivOne{\Delta'}{\Omega'}$.
\end{definition}

\begin{lemma}
\llabel{seRestriction}
if $\encxDelta{n}$ 
and $\encyOmega{n}$ are contexts 
and $\sEquiv{\encxDelta{n}}{\encyOmega{n}}$ 
then \foreachi,  $\sEquiv{\encxDelta{i}}{\encyOmega{i}}$
\end{lemma}
\begin{proof}
\end{proof}


\begin{lemma}
The relation $\wEquivTwo{}{}$ is 
symmetric i.e. if $\Delta=\encxDelta{n}$ 
and $\Omega = \encyOmega{n}$ are contexts 
and $\wEquivTwo{\Delta}{\Omega}$ 
then $\wEquivTwo{\Omega}{\Delta}$.
\end{lemma}
\begin{proof}
Assume $\wEquivTwo{\Delta}{\Omega}$. Therefore there are contexts $\Delta'=\encxDeltap{n}$ and $\Omega' = \encyOmegap{n}$  such that $\sEquiv{\Delta}{\Delta'}$ and $\sEquiv{\Omega}{\Omega'}$ and there is a permutation $\sigma$ such that  $\encyOmegapsigma{n}$ is a context and 
\begin{equation}
\sEquiv{\encxDelta{n}}{\encyOmegapsigma{n}}
\end{equation}

First note that $\wEquivOne{\encyOmegap{n}}{\encyOmegapsigma{n}}$ for $\sigma$ being a  permutation has an inverse permutation $\sigma^{-1}$ and  $\encyOmegapsigma{n}[\sigma^{-1}]=\encyOmegap{n}$.

Now we have $\sEquiv{\encyOmega{n}}{\encyOmegap{n}}$ and $\wEquivOne{\encyOmegap{n}}{\encyOmegapsigma{n}}$ and $\sEquiv{\encyOmegapsigma{n}}{\encxDelta{n}}$.

Therefore $\wEquivTwo{\encyOmega{n}}{\encxDelta{n}}$ as required.
\end{proof}

\begin{lemma}
The relation $\wEquivTwo{}{}$ is 
transitive i.e. if $\Delta=\encxDelta{n}$ 
and $\Omega = \encyOmega{n}$ and $\Gamma$ are contexts 
and $\wEquivTwo{\Delta}{\Omega}$ 
and $\wEquivTwo{\Omega}{\Gamma}$ then 
$\wEquivTwo{\Delta}{\Gamma}$.
\end{lemma}
\begin{proof}
show that the relationship $\wEquivOne$ is transitive. 
\end{proof}

\section{Restricting the class of Generalised Algebraic Theores}

Rationale for restricting to the class of theories in day notes 12 August 2012. \commentary{CHECK this rationale and document here.}
\begin{definition}
A generalised algebraic theory $\gat$ is said to be \term{cancellative}\commentary{Introduced as a restricted theory in day notes of Mon 13 Aug 2018}
iff for all derived rules 
$$
\frac{\con{x}{\Delta}{n}}{\Delta=\Delta'} \mbox{ of $\gat$,}
$$
such that variable $x_n$  does not appear in either
$\Delta$ or $\Delta'$,
$$
\frac{\con{x}{\Delta}{n-1}}{\Delta=\Delta'} \mbox{ is a derived rule of $\gat$.}
$$
\end{definition}
\section{Definitions}
\subsection{Strongly Equivalent Contexts}
\begin{definition}
Contexts $\tuple{\con{x}{\Delta}{n}}$ 
and $\tuple{\con{y}{\Omega}{m}}$ will be said to be \textit{strongly equivalent} iff $n=m$ and for each $i$, $1 \leq i \leq n$ the rule 
$$
\frac{\con{x}{\Delta}{i-1}}{\Delta_i=\Omega_i[x_1|y_1,...,x_{i-1}|y_{i-1}]}
$$
\intheory.
\end{definition}

This definition is that of equivalence from (\cite{Cartmell78})  and, as described there, the  equivalence classes of contexts for a generalised algebraic theory \gat\  are the objects of the contextual category $\conu$ corresponding to \gat.

\begin{oldtt}
\subsection{Aside: Middling Equivalent Contexts}

A context $\tuple{\con{x}{\Delta}{n}}$ 
is  \textit{middling equivalent}  to a $\tuple{\con{y}{\Omega}{m}}$ iff $n=m$ and for each $i$, $1 \leq i \leq n$ the rule 
$$
\frac{\con{x}{\Delta}{n}}{\Delta_i=\Omega_i[x_1|y_1,...,x_{i-1}|y_{i-1}]}
$$
is a derived rule of \gat. \\

This definition doesn't achieve anything -- it doesn't result in an equivalence relation because it is not symmetric for 
\newt{non-cancellative}\oldt{improper} theories.\commentary{same as strong equivalence for cancellative theories}
\end{oldtt}
\subsection{Weakly Equivalent Contexts}
\begin{definition}
A context $\tuple{\con{x}{\Delta}{n}}$ 
is \textit{weakly equivalent} to a context $\tuple{\con{y}{\Omega}{m}}$  iff $n=m$ and there exists a permutation $\sigma$ of the set $\{1,2,..n\}$
such that the sequence 
$\tuple{\permcon{y}{\Omega}{m}{\sigma}}$
is a context and 
is strongly equivalent to $\tuple{\con{x}{\Delta}{n}}$.
\end{definition}
\begin{lemma}
In a cancellative generalised algebraic theory, weak equivalence is an
equivalence relation.
\end{lemma}
\begin{proof}
\tbd\daynote{Mon 12 Aug 2018 for attempted proof.}

\begin{framed}
A good counter example to this lemma was given in day notes Monday 13th Aug 2018 after a supposed proof from the same day.
In fact proof of symmetry is incorrect because the permuted context is not shown to be a context and the counter example shows that
it may fail to be a context.
\end{framed}
\end{proof}

\begin{framed}
This proof breaks down and all heck breaks loose trying to modify the definition of weak equivalence to patch it up.
\end{framed}

\begin{framed}
It occurs to me today (11th November 2020) that one way of patching up the definition is be saying (we(2))
that $\Delta$ w.e(2) $\Omega$ iff there exists $\Delta'$ and $\Omega'$ such that $\Delta$ strongly equivalent to $\Delta'$
and $\Omega$ strongly equivalent to $\Omega'$ and $\Delta'$ we(1) $\Omega'$. Then can reduce each context to a reduced form in which 
no extraneous variables in type expressions.
\end{framed}
\noindent 
If $\tuple{\con{x}{\Delta}{n}}$ 
and $\tuple{\con{y}{\Omega}{m}}$ are weakly equivalent 
and if $\sigma$ is a permutation of the set $\{1,2,..n\}$
such that the sequence 
$\tuple{\permcon{y}{\Omega}{m}{\sigma}}$
is a context and 
is strongly equivalent to $\tuple{\con{x}{\Delta}{n}}$
then define $\sigma$ to be \textit{cannonical}\commentary{Is this one of the definitions of monotonic in the day notes?} iff
for $i_1$,$i_2$,$1 \leq i_1  < i_2 \leq n$, if 
$$
\frac{\con{x}{\Delta}{n}}{\Delta_{i_1}=\Delta_{i_2}}
$$
then $\sigma(i_1) < \sigma(i_2)$.\\

With this weaker definition of equivalence \newt{we would like to}  show that the 
equivalence classes of contexts are the objects of a dependency category
$\depu$ corresponding to a theory \gat. \daynote{ August 2018-- 9th October 2018 struggles to modify this definition to achieve 
this}\\
\begin{framed}
\begin{enumerate}
\item 11th Aug 2018 -- say a permutation is significant if it doesn't permute types that are equivalent
\item this becomes a definition of monotonicity I think -- but is this to do with showing weak equivalence 
is an equivalence relation or is it to do with defining equivalence of realisations -- need to get to bottom of this.
\end{enumerate}
\end{framed}
\noindent
Note that if there is an admittable  permutation between two contexts then it is the unique such permutation. 

\subsection{Weakly Equivalent Realisations}
If $\tuple{r_1,...r_m}: \tuple{\con{x}{\Delta}{n}} \morph \tuple{\con{y}{\Omega}{m}}$ and
$\tuple{r'_1,...r'_m}: \tuple{\con{x}{\Delta'}{n}} \morph \tuple{\con{y'}{\Omega'}{m}}$ are realisations 
and if $\tuple{\con{x}{\Delta}{n}}$ and 
$\tuple{\con{x}{\Delta'}{n}}$ are weakly equivalent and 
$\tuple{\con{y}{\Omega}{m}}$ and $\tuple{\con{y'}{\Omega'}{m}}$ are weakly equivalent then
define $\tuple{r_1,...r_m}$ and $\tuple{r'_1,...r'_m}$ to be weakly equivalent 
iff 
$\tuple{r_1,...r_m}$
and
$\tuple{r'_{\pi(1)},...r'_{\pi{m}}}[x_1|x'_{\sigma(1)},...,x_n|x'_{\sigma(n)}]$
are strongly equivalent
where
$\sigma$ is unique permutation such that 
and $\pi$ is unique permutation such that \highlight{missing text}.


\section{Equivalent Contexts -- Example}

In a g.a.t. \gat which includes the following type definitions:
\begin{displaymath}
\begin{array}{l}
\tstyle \isT{A_1} \\
\tstyle \isT{A_2} \\
\ofT{x}{A_1}\tstyle \isT{B_1(x)} \\
\ofT{x}{A_2}\tstyle \isT{B_2(x)} \\
\end{array}
\end{displaymath}

The context 
\begin{equation}
\ofT{x_1}{A_1},\ofT{x_2}{A_2},\ofT{y_1}{B_1(x_1)},\ofT{y_2}{B_2(x_2)}
\end{equation}
can be rearranged in five different ways to give, all in all, six different ways of expressing
what is much the same context:
\begin{equation}
\ofT{x_1}{A_1}, \ofT{x_2}{A_2}, \ofT{y_1}{B_1(x_1)}, \ofT{y_2}{B_2(x_2)}
\end{equation}
\begin{equation}
\ofT{x_1}{A_1}, \ofT{x_2}{A_2}, \ofT{y_2}{B_2(x_2)}, \ofT{y_1}{B_1(x_1)}
\end{equation}
\begin{equation}
\ofT{x_1}{A_1}, \ofT{y_1}{B_1(x_1)}, \ofT{x_2}{A_2}, \ofT{y_2}{B_2(x_2)}
\end{equation}
\begin{equation}
\ofT{x_2}{A_2}, \ofT{x_1}{A_1}, \ofT{y_1}{B_1(x_1)}, \ofT{y_2}{B_2(x_2)}
\end{equation}
\begin{equation}
\ofT{x_2}{A_2}, \ofT{x_1}{A_1}, \ofT{y_2}{B_2(x_2)}, \ofT{y_1}{B_1(x_1)}
\end{equation}
\begin{equation}
\ofT{x_2}{A_2}, \ofT{y_2}{B_2(x_2)}, \ofT{x_1}{A_1}, \ofT{y_1}{B_1(x_1)}
\end{equation}


Each one of these contexts and each subcontext is represented by an object
in the contextual category  \cat{C(\gat)} that corresponds to the theory. This is
shown in this diagram:

%\scalebox{0.9}{
\begin{shrunkdisplay}
%\scalebox{0.9}{
\begin{array}{cccccc}
\Rnode{c1}{\crossx
              {(\crossx
							     {(\crossx{A_1}{A_2}{1})}
									 {B_1}
									 {A_1}
								) 
								\hspace{-0.2cm}
							}
							{\hspace{-0.2cm}(\crossx{A_1}{B_2}{1})}
							{\crossx{A_1}{A_2}{1}} 
				  } 
&
\Rnode {c2}{\crossx
              {(\crossx{A_1}{B_2}{1})}
							{B_1}
							{A_1}
           }
&
\Rnode{c3}{
           \crossx{B_1}{B_2}{1}
					}
&
\Rnode{c4}{
           \crossx{B_2}{B_1}{1}
					}
&
\Rnode{c5}{\crossx
           {(\crossx{A_2}{B_1}{1})}
					 {B_2}
					 {A_2}
					}
&
\Rnode{c6}{\crossx
              {(\crossx
							     {(\crossx{A_2}{A_1}{1})}
									 {B_2}
									 {A_2}
								) 
								\hspace{-0.2cm}
							}
							{\hspace{-0.2cm}(\crossx{A_2}{B_1}{1})}
							{\crossx{A_2}{A_1}{1}} 
				  } \\ [1.2cm]
\Rnode{sc1}{\crossx
							     {(\crossx{A_1}{A_2}{1})}
									 {B_1}
									 {A_1}				
				   } 
&
\Rnode {sc2}{
             \crossx{A_1}{B_2}{1}
           }
&
\Rnode{sc3}{
           \crossx{B_1}{A_2}{1}
					}   
&
\Rnode{sc4}{
           \crossx{B_2}{A_1}{1}
					}   
&
\Rnode{sc5}{
           \crossx{A_2}{B_1}{1}
					}   
&
\Rnode{sc6}{
           \crossx
					         {(\crossx{A_2}{A_1}{1})}
					         {B_2}
									 {A_2}
					}  \\ [1.2cm]
\multicolumn{2}{c}{
         \ \ \ \ \ \ \ \Rnode{A1xA2}{\crossx{A_1}{A_2}{1}}   
} 
& \Rnode{B1}{B_1} 
& \Rnode{B2}{B_2}
& \multicolumn{2}{c}{\Rnode{A2xA1}{\crossx{A_2}{A_1}{1}}}
\\ [1.2cm] 
  \multicolumn{3}{c}{ \hspace{3cm}\Rnode{A1}{A_1}}
& \multicolumn{3}{c}{             \Rnode{A2}{A_2}} \hspace{3cm} \\ [1.2cm]
\multicolumn{6}{c}{
         \Rnode{abs}{1}
					        }
\end{array}
\end{shrunkdisplay}
%}
 
\ncsar{c1}{sc1}
\ncsar{c2}{sc2}
\ncsar{c3}{sc3}
\ncsar{c4}{sc4}
\ncsar{c5}{sc5}
\ncsar{c6}{sc6}
\ncsar{sc1}{A1xA2}
\ncsar{sc2}{A1xA2}
\ncsar{sc3}{B1}
\ncsar{sc4}{B2}
\ncsar{sc5}{A2xA1}
\ncsar{sc6}{A2xA1}
\ncsar{A1xA2}{A1}
\ncsar{B1}{A1}
\ncsar{A1}{abs}
\ncsar{B2}{A2}
\ncsar{A2xA1}{A2}
\ncsar{A2}{abs}

Each of the six contexts are equivalents. This is reflected by the facts that there are cannonical 
isomorphisms in the contextual catgory between the six objects: \\
\begin{displaymath}
\def\arraystretch{2.2}
\begin{array} {l}
      \crossx
          {(\crossx
			       {(\crossx{A_1}{A_2}{1})}
						 {B_1}
						 {A_1}
			     ) 
			     \hspace{-0.2cm}
					}
					{\hspace{-0.2cm}(\crossx{A_1}{B_2}{1})}
					{\crossx{A_1}{A_2}{1}}
\\
    \crossx
       {(\crossx{A_1}{B_2}{1})}
							{B_1}
							{A_1}
\\

     \crossx{B_1}{B_2}{1}
\\
     \crossx{B_2}{B_1}{1}
\\
     \crossx
           {(\crossx{A_2}{B_1}{1})}
					 {B_2}
					 {A_2}
\\
     \crossx
              {(\crossx
							     {(\crossx{A_2}{A_1}{1})}
									 {B_2}
									 {A_2}
								) 
								\hspace{-0.2cm}
							}
							{\hspace{-0.2cm}(\crossx{A_2}{B_1}{1})}
							{\crossx{A_2}{A_1}{1}} 

\end{array}
\end{displaymath}
\section{Cancellative Property}
\noindent 
\begin{definition}
A contextual category \ccat is said to be cancellative iff for all objects $X$, $A$, $B$ and $B'$ such that
$X < A$, $X < B$ and $X < B'$ in \ccat, 
if $\crossx{A}{B}{X} = \crossx{A}{B'}{X}$ then $B=B'$. \footnote{What about requiring q's to b equal?}.
\end{definition} 
\textbf{Assume from now on that we always mean contextual categories to be cancellative.}

\section{Decompositions and Projections of Objects}
For any given object $Y$ of a contextual category there may or may not be objects $A$, $B$ and $X$ so that
$Y = \crossx{A}{B}{X}$. If there are we say the object is decomposable. If an object $Y$ is decomposable into objects that
are themselves decomposable then there may be a number of decompositions implied by coherence conditions of contextual
categories but since we we assume the cancellative property upto these equivalent decompositions given by coherence
the decomposition is unique. 
\noindent
For example if $X < A$, $X < B$, $X < C$ in a \ccat then from coherence of pullbacks in \ccat we have:
$\crossx{A}{(\crossx{B}{C}{X})}{X} = \crossx{(\crossx{A}{B}{X})}{C}{X}$ as shown here:
 
\begin{displaymath}
\begin{array}{cp{1.0cm}cp{1.0cm}c}
\Rnode{J1}{}\Rnode{D1} {\crossx{(\crossx{A}{B}{X})}{C}{X}}\Rnode{J2}{} \ \ \ \ \   &&  &&  \\ 
= && && \\
\Rnode{D2} {\crossx{A}{(\crossx{B}{C}{X})}{X}}    &&  &&                        \\ [1.3cm]
\Rnode{AB}{\crossx{A}{B}{X}}&& \Rnode{BC}{\crossx{B}{C}{X}} &&                      \\[1.3cm]
\Rnode{A}{A}&& \Rnode{B}{B} && \ \ \ \ \ \ \ \ \ \ \ \ \ \Rnode{C}{C}                                        \\[1.3cm]
             && \Rnode{X}{X} &&                                                     
\end{array}
\end{displaymath}

\ncaarr[50]{q(\pc{\crossx{A}{B}{X}}{X},C)}{J2}{C}
%\ncsar[-30]{J1}{AB}
\ncsar{D2}{AB}
\ncsar{AB}{A}
\ncsar{BC}{B}
\ncsar{A}{X}
\ncsar{B}{X} 
\ncsar{C}{X}
\ncaarr{q(\pc{A}{X},B)}{AB}{B}
\ncaarr{q(\pc{B}{X},C)}{BC}{C}
\ncaarr{q(\pc{A}{X},\crossx{B}{C}{X})}{D2}{BC}

\noindent Say that an object is primitive if it cannot be decomposed.
\noindent Anyway we look at it we can decompose an object into an n-tuple of simple objects. 
\noindent Any object $Y$ of a contextual category \ccat is either primitive or else there is an $n > 1$ and a tuple of morphisms $f_i: Y \morph A_i$, $1 \leq i \leq n$, such that for each $i$, $1 \leq i \leq n$,  $A_i$ is primitive, 
and the diagram
\begin{center}
\psset{arrowsize=5pt,arrowinset=0.7}
\psset{nodesep=7pt}
\setlength{\arraycolsep}{0.3cm}
\begin{displaymath}
\begin{array}{ccp{0.2cm}p{0.075cm}c}
&&\Rnode{top}{$Y$}& & \\[1.5cm]
\Rnode{A}{A_1}&\Rnode{B2}{A_2}&& ... & \Rnode{Bm}{A_n}\\
\end{array}
\end{displaymath}
\psset{nodesepA=4pt,nodesepB=7pt}
\ncline{-}{top}{A}
\tlput{\footnotesize $f_1$}
\ncput[nrot=:R,npos=1]{\pstriangle(0,0)(.2,.2)}
\ncline{-}{top}{B2}
\trput{\footnotesize $f_2$}
\ncput[nrot=:R,npos=1]{\pstriangle(0,0)(.2,.2)}
\ncline{-}{top}{Bm}
\trput{\footnotesize $f_n$}
\ncput[nrot=:R,npos=1]{\pstriangle(0,0)(.2,.2)}
\end{center}
 is a limit cone to its own characteristic diagram.

\noindent The object $\crossx{A}{B}{X}$ has projections $\pc{\crossx{A}{B}{X}}{A}: \crossx{A}{B}{X} \smorph A$ and
$q(\pc{A}{X},B): \crossx{A}{B}{X} \morph B$. Given that  $A$ and $B$ are maybe themselves decomposable then we  can define a tree of morphisms which we shall call the projection tree. Whereas the projection tree is not unique the leaves of the tree are and the composed morphisms to the leaves of the tree are too. The composed morphisms to the leaves of the tree we
will call the \textit{basis} of the object $Y$. 

\newpage
\noindent If $X < Y$ in a contextual category \ccat then we define the projection tree of $Y$ beyond $X$, which we denote
$\proj{Y}{X}$, recursively as follows:
\begin{enumerate}
\item if $Y$ is primitive then 
$$\proj{Y}{X} = id_Y$$

\vspace{2cm}

\item
if $Y$ is $\crossx{B}{C_m}{A}$ for some $B$, $C_1,...C_m$ and $A$ where  $A \base C_1 \base ... \base C_m$ and $A < B$ then
\vspace{1.0cm}
\begin{displaymath}
\begin{array}{cp{0.5cm}cp{0.2cm}c   p{0.2cm}c c  p{0.2cm}l   p{0.2cm}c   c  p{0.1cm}c }
&& \Rnode{A}{A} &&    &&&     \Rnode{B}{B} && \Rnode{AC1}{\crossx{B}{C_1}{A}} 
&& \Rnode{AC2}{\crossx{B}{C_2}{A}}     &\Rnode{BCe}{\hdots}    && \Rnode{BCm}{\crossx{B}{C_m}{A}}           \\ [0.6cm]
               &&                && \Rnode{C1}{C_1}   && \Rnode{C2}{C_2} &\Rnode{Ce}{\hdots}   && \Rnode{Cm}{Cm}     &&  &&     
\end{array}
\end{displaymath}
\ncsar{C1}{A}
\ncsar{C2}{C1}
\ncsar{Cm}{Ce}
\ncsar{AC1}{B}
\ncsar{AC2}{AC1}
\ncsar{BCm}{BCe}
\ncaarr[15]{q(\pc{B}{A},C_m)}{BCm}{Cm}
\nccdar{B}{A}
\ncbcdar[-30]{\pc{\crossx{B}{C_m}{A}}{B}}{BCm}{B}
\vspace{0.5cm}

\begin{enumerate}
\item 
if $X < B$
$$
\proj{\crossx{B}{C_m}{A}}{X} = \{ \tuple{\pc{\crossx{B}{C_m}{A}}{B},\proj{B}{X}} ;\tuple{q(\pc{B}{A},C_m),\proj{C_m}{A}} \}
$$
\item 
if $X = B$
$$
\proj{\crossx{B}{C_m}{A}}{B} = \{  ;\tuple{q(\pc{B}{A},C_m),\proj{C_m}{A}} \}
$$


\item if $X$ is $\crossx{B}{C_i}{X}$ 
$$
\proj{\crossx{B}{C_m}{A}}{\crossx{B}{C_i}{A}} = \{\tuple{q(\pc{B}{A},C_m),\proj{C_m}{C_i}}\}
$$

\end{enumerate}
\end{enumerate}

\newpage

\section{Sketch}
\begin{enumerate}
\item{
from a contextual category $\cat{C}$ we may construct a dependency category $dep(\cat{C})$ as a quotient category
}
\item{
The functor $q:\cat{C} \morph dep(\cat{C})$ is surjective and faithful.
Therefore it preserves limits.
}
\end{enumerate}




\section{Equivalence Classes of Contexts}
\subsection{Subcontexts}
By a \textit{subcontext} of a context $\context{x}{\Delta}{n}$ we mean any context
of the form
$\subcontext{x}{{\Delta}}{i}{k}$ for some sequence 
$1 \leq i_1 \leq i_2 \leq ... i_k \leq n$. \\

\noindent
For example:
\begin {itemize}
\item the empty context  is a subcontext of every 
context,
\item any initial subsequence of a context is a subcontext of that context.
\end{itemize}

\noindent
In addition to the above, a context such as $\ofT{x_1}{A_1},\ofT{x_2}{A2},\ofT{y}{B(x_1)}$ has 
$\ofT{x_2}{A2}$ as a subcontext. This illustrates that each non-initial subcontext is indicative of a type in the context ($A_2$ in this case)
which does not depend on one of the earlier variables.
 



\bibliography{../SharedBibliography/temp/bibliography}

\end{document}
