
\begin{frame}{Definition -- \datacat}
\medskip
I shall use the shorthand \textit{\datacat} to mean a triple $\tuple{\catc,M,v}$ where 
\begin{itemize}
\item \catcw is a \rangeplus category with specified finite restriction products,
\item $M$ is a set of designated monomorphisms of \catcw closed under composition
and such that each $m \in M$   has a partial inverse $m^{-1}$,
\item a distinguished object $v$, such that every morphism $f: v \morph x$ in \catcw 
factors through $m^{-1}$, for some monomorphism $m$.
\end{itemize}

\medskip
Note that it follows from this definition that a sketch for a \textit{\datacat} has no need for edges with domain $v$.
\end{frame}

\begin{frame}{Data Specification and Data Specification Instance}
In this presentation, 
\begin{itemize}
\item by \textit{data specification} I shall mean a sketch for a \datacatw such that the designated object $v$ has no outgoing edges -- 
 neither edges  \underline{$v \morph v$} 
nor edges \underline{$v \morph non\localjchyphen v$}.\\
\item If $S$ is a sketch for \datacatw denote by $\catc(S)$ the \datacatw generated from $S$. \\
\pause \item Define an \textit{instance} of a data specification $S$ to be a 
range functor $F: \catc(S) \morph \Par$ 
that preserves the specified restriction products
and maps the object $v$ to the set $V$. \\
\medskip
Note that such an $F$ will preserve designated monomorphisms and their inverses.
\end{itemize} 
I will muddle up data specifications and sketches in these slides. \\ 
I will speak of $\catc(S)$ as the theory category. \\
\medskip
Next I want to give some examples to show how all this works in practice.
The very next next example is of the ... relational model of data.
\end{frame}

\ifNotesnAll
\begin{frame}{Example}
Now I want to give some examples to show how all this works in practice and I want to start with the dear old relational model of data. 
\end{frame}
\fi
\input{../slides/Shlaer&Lang/relationalDataSlide}  % this slide defines inclusion dependencies
\input{../slides/Shlaer&Lang/relationalDataSpecificationSketchSlide}

\begin{frame}{Classifying Data Specifications}
\begin{center}
\pstree{\TR{\framebox[3cm][c]{data specification}}}{%
\TR{\framebox[2cm][c]{logical}}
\pstree{\TR{\framebox[2cm][c]{physical}}}{\TR{\framebox[2.5cm][c]{relational}} \TR{\framebox[2.5cm][c]{non-relational}}}
}
\end{center}
\medskip
\begin{itemize}
\pause \item in \textit{relational} sketches all edges are 
of the \underline{$non\localjchyphen v \morph v$} type and each such represents a column of a table/relation,
\pause \item other \textit{physical sketches} (\textit{non-relational}) in addition 
to the \underline{$non\localjchyphen v \morph v$} type edges have edges of the \underline{$non\localjchyphen v \morph non\localjchyphen v$} type and these represent structural containment,
\item non-relational physical data specifications are also said to be \textit{hierarchical}.
\end{itemize}
\end{frame}

\begin{frame}{Characterisation of Relational Data Specifications}
\begin{definition}
A data specification is \textit{relational} iff
\begin{itemize}
	\item all edges are of the \underline{$non\localjchyphen v \morph v$} type,
\item every non-v-node is the domain of at least one v-valued mono-source
i.e. for every non-v-node $a$, for some $n \geq 1$, there exists a source
\begin{displaymath}
\parallelsource{a}{v}{m}{n}
\end{displaymath}
which is designated as a mono-source i.e. for which $\tuple{m_1,...m_n}$ is a designated monomorphism.
\end{itemize}
\end{definition}
\end{frame}

\begin{frame}{Construction -- Transform a Relational Sketch to a Logical Sketch}
\begin{lemma}
For any classic relational data specification 
there is an equivalent data specification
(i.e. one with the same theory category) which is logical.
\end{lemma}

\begin{proof}
In outline: We construct a series of equivalent sketches by eliminating each 
inclusion dependency in turn. When all eliminated the resulting sketch is the required logical sketch. Eliminate the inclusion dependency 
$a[f_1,...f_n] \subseteq b[m_1,...m_n]$
as follows:
\begin{itemize}
\item remove the inclusion dependency,
\item replace by an edge $f: a \morph b$, 
\item remove those $f_i$ that are edges and 
rewrite any occurrence of such $f_i$ in the remaining inclusion dependencies by $f \circ  m_i$, 
\item for those $f_i$ that are not edges add a path equivalence (i.e. a commuting diagram)
$f \circ m_i = f_i$.
\end{itemize}
\end{proof}
An example follows.
\end{frame}

\input{../slides/Shlaer&Lang/relational2logicalSketchTransformationSlide}

\input{../slides/Shlaer&Lang/logicalDataSpecificationSketchSlide}

\begin{frame}{Characterisation of  Logical Data Specification}
\begin{definition}
A data specification $S$ is \textit{logical} iff
\begin{itemize}
	\item there does not exist an edge $e$ of the sketch $S$ for which there is a decomposition in the theory category $\catc(S)$ i.e. such that 
	for some morphisms $f_1$ and $f_2$  distinct from $e$,
	$e = f_1 \circ f_2$. 
\end{itemize}
\end{definition}
\end{frame}

\newcommand{\transone}{\parbox{2cm}{\textit{Chen's 1976\\transform \\(automated)}}}
\newcommand{\transtwo}{\parbox{2cm}{\textit{normalise \\ a la Codd \\(manual)}}}
\begin{frame}{Best Practice -- Structured Systems Analysis and Design}
Current Best Practice
\begin{center}
\begin{tabular}{ p{1.25cm} p{2cm} p{1.25cm} p{2cm}  p{1.25cm}}
\parbox{3cm}{logical \\ data \\ specification}&  $\xRightarrow{\transone}$ & \parbox{3cm}{relational \\ data \\specification} & $ \xRightarrow{\transtwo} $ & \parbox{3.5cm}{relational \\data \\specification}
\end{tabular}
\end{center}
\end{frame}

\begin{frame}{ Chen's Transformation 1976 \onslide<3->{\textcolor{red}{made diagram aware}}}
\begin{construction}
\onslide<3>{\hspace{6.4cm}\textcolor{red}{with the same theory category}\\}
From a logical data specification construct a relational data specification
\onslide<3>{\raisebox{-0.1cm}{\Large\textcolor{red}{$\leftthreetimes$}}}. 
\end{construction}
\textbf{Chen's 1976 Method}
Replace $f:a \morph b$ in the sketch by edges $\fn$
where $\wanton{m}$ is a v-valued mono-source with domain $b$ 
and add inclusion dependency 
$a[f_1,...f_n] \subseteq b[m_1,...m_n].$

\pause
\textbf{Problem with this method}
\begin{itemize}
	\item Doesn't take account of commutative diagrams,
	\item therefore resulting relational specification
	\begin{itemize}
		\item doesn't have equivalent theory category,
		\item often is not be in normal form.  
	\end{itemize}
	\item This weakness negatively impacts how data specifications are written and maintained.
\end{itemize}
\pause
\textbf{Mission}
\begin{itemize}
	\item Theoretically justify an improved algorithm, i.e. one that takes account of commutative diagrams, and thereby change how data specifications are managed and databases are programmed.
\end{itemize}
\end{frame}

%\newcommand{\newtrans}{\parbox{2cm}{\textit{diagram \\ aware \\ transformation \\ (automated)}}

\input{../slides/revisedBestPractice}
\input{../slides/Shlaer&Lang/nestedRelationalDataSlide}
\input{../slides/Shlaer&Lang/hierarchicalDataSpecificationSketchSlide}

\begin{frame}{Characterisation of Physical Data Specification}
\begin{definition}
A data specification is \textit{physical} iff
\begin{itemize}
\item every non-v-node is the domain of at most one edge of the \underline{$non\localjchyphen v \morph non\localjchyphen v$} type.
\end{itemize}
\end{definition}
In a physical data specification every node and every edge has physical significance
in the database or message structure.
\begin{itemize}
\item Nodes other than v in a physical data specification represent entity types (ER-notation) 
or  tables (relational) or structs (IDL) or similar.  
\item Edges of the \underline{$non\localjchyphen v \morph non\localjchyphen v$} type represent those relationships in the data that are physically represented  by \textit{structural containment}.
\item Remaining edges (i.e. those of the \underline{$non\localjchyphen v \morph v$} type) represent
attributes (ER) or columns of tables (relational) or scalar fields within structs (IDL) or such like.  
\end{itemize}
\end{frame}

\input{../slides/Shlaer&Lang/contextualSketchSlide}

\begin{frame}{Example -- LCMSMS Data}
\scalebox{0.2}{
\input{\erpictureFolder/chromatogramanalysisrecord}
}
\pause
\vspace{-5cm}
\begin{block}{}
This example has 
\begin{itemize}
  \item 33  relationships implemented by structural containment, 
  \item 26  relationships implemented by relational referencing (inclusion depedencies),
  \pause \item 16 non-trivial commutative diagrams,
  \pause \item 6 pullback diagrams.
\end{itemize}
Generated into code in XML, ECMA Javascript and Python. 
\end{block}
\end{frame}

\ifNotesnAll
\begin{frame}Example LCMSMS Notes
\begin{itemize}
  \item This is an example of a data specification written by myself on a project when I worked for a science company. This was a project to build an application for managing and interpreting data from mass spectrometers using liquid chromatography. The application was written in javascript and python and used XML and IDL at various positions in the architecture.
\end{itemize}
\end{frame}
\fi

