
\usepackage{mathptmx}
\usepackage{amsfonts}
\usepackage{wasysym}
\usepackage{url}
\usepackage{hyperref}

\newcommand{\sharedmacros}{../../../SharedMacros}
%\input{\sharedmacros/general.macros}    %causes problems when used with bamer
\input{\sharedmacros/ccategories.macros}
\input{\sharedmacros/gats.macros}
\input{\sharedmacros/metagat.macros}
\input{\sharedmacros/ermacros}
\input{\sharedmacros/erdiagram}    %beamer aware version
\input{\sharedmacros/general.macros.v2} %try this instead
\input{\sharedmacros/diagram.macros}
\input{../macros/beamermacros}
%\input{../macros/generalmacros} %get rid of this
\input{../../macros/othermacros}
\input{../macros/sketchmacros}
\input{../macros/directedgraphsmacros}




\renewcommand{\erpictureFolder}[0]{../../../SharedPictures}
\setcounter{equation}{0}

\makeatletter
\newcommand{\xRightarrow}[2][]{\ext@arrow 0359\Rightarrowfill@{#1}{#2}}
\makeatother


\title[John Cartmell]{Universals and Particulars and the Nature of Data}


\author{John Cartmell}
\institute{\\}
\date{Mar 3, 2025}
\bibliographystyle{plainnat}
\usepackage{framed}
\usepackage{bibentry}
\usepackage{colortbl}
\usepackage{ulem}   % for \dashuline{dashing} for seconday key
\usepackage{soul,xcolor} % also used for striking out text 

\usepackage{listings}
\lstset{%
  escapeinside={(*}{*)},%
}
\usepackage{arydshln} % vertical dashed lines between columns of an array
\usepackage{pst-arrow} %for \bigarrow


%Redefine the \Fin macro to be category of sets
%\renewcommand{\Fin}{\Set}



\newcommand{\fgsourcediag}{$\binarysourcediag{a}{b}{c}{f}{g}$}
\newcommand{\rangeplus}{RR.5 range } % has trailing space
\newcommand{\datacat}{$\gamma$-structured category}
\newcommand{\datacatw}{\datacat\ }
\mathchardef\localjchyphen="2D
\newtheorem{construction}[theorem]{Construction}

\newcommand{\IfSforGammaCwithRCwords}{
If $S$ is a sketch for \datacatw \catcw considered as a data specification with requirement $\reqtc$\ }
\newcommand{\IfSforGammaCwithRCwordsvariant}{
If $S$ is a sketch for \datacatw  \catcw and if $S$ is considered as a data specification with requirement $\reqtc$\ }

\begin{document}

\setstcolor{red}
\begin{frame}
\titlepage
\nobibliography{../../SharedBibliography/temp/bibliography}
\end{frame} 


\begin{frame}{Terminology}
\begin{itemize}
\item \textit{modelling} for me is \textit{theorizing}
\item I speak of \textit{instances of theories}  rather then \textit{models of theories}
\item I speak of \textit{data specifications} except when I forget and I call them \textit{data models}
\item the act of constructing data specifications is \textit{data modelling}
\item a \textit{model of data} is a meta-theory (a meta-model) describing what constitutes a data specification. Most significantly there are
\begin{itemize}
\item \textit{relational} and 
\item \textit{nested relational} 
models of data
\end{itemize}
\item the \textit{mathematical theory of data} is a meta-theory of data that supports technology  independent reasoning about data specifications in all their forms.
\end{itemize}
\end{frame}

\begin{frame}{Why?}
\begin{itemize}[<+->]
  \item There are gross inefficiencies in the methodologies and working practices used in  a key activity in s/w systems development and maintenance namely in  the creation and maintenance of specifications of the data stored in databases and 
  represented in messages variously intra-communicated between components of systems and inter-communicated between systems. 
  \item These inefficiences have been established and endorsed by a theory which is grossly inadequate.
  \item A new theory is required to expose and remedy the shortcomings.
  \item The challenge is to positively impact best practice.
\end{itemize}
\end{frame}

\begin{frame}{Mathematical Theory of Data}
\begin{itemize}
\item is a meta-theory,
\item it covers \textit{principles} and \textit{criteria} for goodness of data specifications,
\item it reveals the significance of commutative diagrams and therefore category theory.
\item The slogan on the tin is \textit{Good Data Modelling is Good Theorising}. 
\end{itemize}
\end{frame}

\input{../slides/normalformliterature.slides}
\input{../slides/generalintroductory.slides}
\input{../slides/specificintroductory.slides}
\input{../slides/universals.slides}

\end{document}