\section{Example --- Chen Manufacturing Company}
\label{ChenManufacturingCompany}
In figure \ref{ChenManufacturingExample} I have reproduced an example entity relationship diagram  given in Chen's 1976 paper. This example documents principle types of entity in the organisation of an imaginary manufacturing company.  I wanted to redraw
this diagram in the Barker-Ellis style used in this book. To do this I needed to make some educated guesses about the meaning of some of Chen's relationships since use of the Barker-Ellis style  required me to descriptively label the relationships. Having made these guesses I drew the diagram shown
 in figure \ref{chenManufacturingCo..diagram}. 

\begin{erboxedFigure} {H}{ChenManufacturingExample}{
In effect this is the first ever entity relationship diagram. It is an entity relationship diagram in the Chen style from his seminal paper of 1976 and it is there described as being an analysis of information in a manufacturing firm. In this diagram
diamonds represent what we describe here as relationships and boxes represent types of entities.  
For completeness we should mention that in Chen's 1976 terminology the boxes were said to be entity sets and the diamonds were said to be relationship sets but by 1983 in a paper published in that year Chen is instead using the terms `entity type' and `relationship type'. While sympathising with this terminology we think it correct to to use the term `relationship' rather than 'relationship type' because this brings us closer to the terminology of formal logic.
}
\begin{center}
\scalebox{0.80}{\input{\handCraftedImagesFolder/ChenStyleManufacturing.tex}}
\end{center}
\end{erboxedFigure}

\begin{erboxedFigure} {H}{chenManufacturingCo..diagram}{
Chen's 1976 entity relationship diagram (shown in figure \ref{ChenManufacturingExample}) 
redrawn as a structured entity model using the Barker-Ellis notation. 
I have to guess some of the meaning of Chen's relationships in order to label them meaning fully in
 this diagram. 
 The rail at the top of the diagram represents the whole of the model which in this case is the manufacturing company in question.
}
\begin{center}
\scalebox{0.95}{\input{\ImagesFolder/chenManufacturingCo..diagram.tex}}
\end{center}
\end{erboxedFigure}
\newpage
\subsection*{Ternary Relationships}
I should draw attention to the fact that Chen's example as shown in figure \ref{ChenManufacturingExample} contains an 
example of a ternary relationship between a project, a supplier and a part which is shown  like this:

\begin{equation}
\raisebox{-1.0cm}{\scalebox{0.85}{\input{\handCraftedImagesFolder/ChenStyleSupplierProjectPart.tex}}}
\end{equation}

In the Barker-Ellis style there are no ternary relationships nor are any needed since ternary relationships can be represented as a combination of binary relationships and entity types. 
The fact that all relationships are binary and are represented by lines only rather than diamonds  is of practical importance in that more details can be fitted onto diagrams.
\mynote 
Those constructs which Chen represents as diamonds 
names as relationships sets when he first introicues them 
in 1976 and then describes as relationship types in a revised terminology  in his 1983 paper
(when also entity sets have become entity types).
His 1983 paper says equates these relationship types (using the '83 terminology) are suggested by
the occurrence of verbs in English in statements like
\begin{center}
example here
\end{center}

The general principle should be that these same constructs should be 
represented by entity types in Barker-Ellis style diagrams which a characterised by having
all their outgoing relationships being  annotated with bars.\commentary{I am struggling with this explanation because I haven't done idenitfying features yet.} 
The general principle : replace the diamond by a box and draw \textit{n} many-one relationships 
leaving the box one for each of Chen's connectors, draw bars across the relationships and this will indicate that the entity type is representing n-ary relation a relationship type in Chen's terminology.
For some authors such an entity type is said to be an intersection entity (type)
and for some it is said to be an associative entity (type). Mathematically it represents an n-ary relation but I hesitate to use that term because of the overload of terms relation and relationship. This overload gets worse earher than better when we get to the section on data. 

\mynote 
Translating Chen's \verb'SUPP-PROJ-PART' construct directly into Barker-Ellis notation we get a frag ment like this
\begin{center}
\scalebox{0.95}{\input{\ImagesFolder/ChenSupplierProjectPartBarkerStyle.tex}}
\end{center}
but don't ask me what it means exactly. 
When I translating it in figure \ref{chenManufacturingCo..diagram} I made a guess as to what Chen exactly intended. 
\commentary{Don't forget I can come back to this and explain further later in the section on identifying features.}
Having made my guess I have represented  by two types \textit{project supply option}
and type \textit{part used on project} . 
I hope that you agree that this Barker-Ellis style diagram leaves less to the imagination than Chen's original diagram did --- his meanings for his relationships in this area are not clear.

\mynote
Now add mathematical characterisation of intersention entitites and indicate that this is
the meaning of the bars and is part of a more general notation that we come to shortly.

\mynote Add that mathamtical a set of jointly injective functions with common domain is equivalent to a subset of a cartesian product type. This then brings us back to the definition in mathematics of a relation as a subset of a cartesian prouct of sets. \commentary{debug this language.}

\newpage
\subsection{Background Wikipedia entry for `Part number'}
Has this para headed ``the part design versus instantiations of it''
As a part number is an identifier of a part design (independent of its instantiations), a serial number is a unique identifier of a particular instantiation of that part design. In other words, a part number identifies any particular (physical) part as being made to that one unique design; a serial number, when used, identifies a particular (physical) part (one physical instance), as differentiated from the next unit that was stamped, machined, or extruded right after it. This distinction is not always clear, as natural language blurs it by typically referring to both part designs, and particular instantiations of those designs, by the same word, "part(s)". Thus if you buy a muffler of P/N 12345 today, and another muffler of P/N 12345 next Tuesday, you have bought "two copies of the same part", or "two parts", depending on the sense implied.

