\section{Communicating Entity Instances}
\label{CommunicatingEntityInstances}

\begin{newtt}
\mynote
Sometimes in a situation being modelled, entities of a type are named and then in the model
we can represent name as an attribute associated with entities of the type.  
Such name attributes are archetypal examples of identifying attributes --- we name things so that we can identify them in context. In such a situation the name of an entity, i.e. the value of the name attribute,  can be used in context to definitively identify the entity.
Quoting this value, i.e. using the name, is one way of referencing an entity of the type from within the context of another entity. This is illustrated in  an example given earlier in which we presented this statement
\begin{equation}
\label{JupiterIoShort}
\mbox{\textit{Jupiter is orbited by Io.}}
\end{equation}
as a representation of an instance of the binary relationship
\begin{gather}
\label{planetMoonSecondInstance}
\raisebox{-1cm}{
\input{\ImagesFolder/planetMoon.tex}
}
\end{gather}

In the context of statement (\ref{JupiterIoShort}), Jupiter is the name of, and therefore  a reference to, a planet and  Io is the name of, and therefore a reference to, a moon i.e. a natural satellite.  
We understand  (\ref{JupiterIoShort}), therefore, to be shorthand for the 
more pernickerty 
\begin{equation}
\mbox{\textit{The planet with name `Jupiter' is orbited by the moon with name `Io'.}}
\end{equation}

\mynote
We are very much stating the obvious of course but there is a pattern here and it is repeated over and over when instances of relationships are communicated  and the same pattern occurs more formally  and uniformly when instances of relationships are communicated and/or stored in data.

\mynote
Note that within the context of the solar system the name Io  doesn't uniquely identify this moon of Jupiter amongst all named solar system objects since it is also the name of an asteroid that orbits within the asteroid belt.
The names of moons are only unique as names of moons. 
Of course in other contexts, both Jupiter and Io are the names of mythological figures. 

\mynote
 Not just name attributes but identifying attributes more generally provide us with ways of 
 referencing entities and thereby communicating instances of relationships. 
 This is often the most pragmatic reason for requiring types to have associated identifying feature. 

 \mynote
Focusing on just this aspect of the solar system we  draw the model shown in figure
\ref{planetMoonModel}
containing the relationship () and showing name attributes of both moons and planets and
underlining these to showing that each of them is unique and therefore identifying. 

\begin{erboxedFigure} {H}{planetMoonModel}
{
caption blah blah blah
 }
\input{\ImagesFolder/planetMoonModel.tex}
\end{erboxedFigure}

\mynote
Whereas the names of moons and planets are unique in the context of the solar system then the names of characters in Shakespeare's plays are not. Shakespeare's characters are named uniquely only within the context of individual plays. To document this on an entity relationship diagram we underline the name attribute and put a bar through the relationship that contributes to the identification like this \barkerEllisJ, for example, as we illustrate twice over in figure \ref{playwrightPlayCharacterModel}.


\begin{erboxedFigure} {H}{playwrightPlayCharacterModel}
{
caption blah blah blah
 }
\input{\ImagesFolder/playwrightPlayCharacterModel.tex}
\end{erboxedFigure}

In this diagram we have shown characters being identified by name and by play 
and plays being identified by title and by playwright. Playwrights we have indicated are identified by their names.\commentary{cascaded identifier}

\mynote What this diagram implies is that characters from plays need be identified by three referentials - the name of the character, the title of the play and the name of the playwright.
Example.
\mynote Naively this implies that if we wish to communicate a instance of relationship between two different plays then we will be required to communicate six referentials
 -- three to identify one party in the relationship and three to identify the other party. This is not the case, and so not what we instinctively do, when we have \textit{a priori} knowledge of the relationship being communicated is limited in scope. If we know the relationship is confined to individual playwrights or is internal to individual play or script then clearly we do not have to communicate the playwright and play title twice over. The number of referentials may in such a case collapse to four: name of playwright, title of play and name of character.
Shlair-Lang in a very much underrated paper coined the term \textit{collapsed referentials} for this phenomena in the context of data representation of models represented in entity relationship notation.

\mynote 
So we have, in the context of the plays of William Shakespeare,
\begin{equation}
\label{AntonioLovesSebastian}
\mbox{\textit{The character Antonio in Twelfth Night loves Sebastian.}}
\end{equation} 

You might think that a pedantic elaboration of (\label{AntonioLovesSebastian}) would be:
\begin{equation}
\text{\parbox{9cm}{\textit{The character named Antonio in Shakespeare's play Twelfth Night loves the character named Sebastian in Shakespeare play Twelfth Night.}}}
\end{equation} 
but there is a reason why the shorter
\begin{equation}
\text{\parbox{9cm}{\textit{The character named Antonio in Shakespeare's play Twelfth Night loves the character named Sebastian.}}}
\end{equation} 
is equally unambiguous. This is because the relationship `loves' in this context is a relationship which is internal to the plot of a single individual play. 
It is an intra-play relationship rather than an inter-play relationship.
Reading from left to right we establish that we are making a statement about 
Antonio from Twelfth Night  and so must look to establish Sebastian as a character from the same play i.e. also from Twelfth Night. This is an example of a relationship which is not global in scope (it is local to the play) and therefore describing instances of this fictional loves relationship
doesn't require as many referential attributes as would otherwise be the case if it was global in scope. 

The relationship is conveyed by a total of  4 referential attributes
(Shakespeare, Twelfth Night, Antonio and Sebastian) not the six that might have been required
if the relationship was global in scope so that characters from one play might be described as being in love with a character from a different play.

On the other hand contrast with
\begin{equation}
\mbox{\textit{Romeo from Shakespeare's Romeo and Juliet resembles Pyramus from Ovid's 
Metamorphoses.}}
\end{equation}

in this there is a relationship instance conveyed by six referentials. This is because the resemblance relationship to is not an internal relationship. It is between characters and a character from one play may be perceived as resembling a character from a different play.

In the first example we say that there are collapsed referentials. 

None of these referentials collapse. The models of the two relationships
look the same but behind the scenes they are different. They have different scopes and we come back to talking about these later.

Antonio and Twelfth Night refer
Sebastian and Twelfth Night refer.

Twelfth Night doubles up.

Now two plays are referenced 
 \end{newtt}

\mynote From this section most interest and significance  in how relationship instances are/can be communicated but this question cannot be separated from how entitites are communicted and how entity identitites are communicated.

\mynote I will tend talk abouty how they are commuincated in data but everything that is siad applies to how they are communicated generally including nay especially in natural language. There is something deeply logical and linguistic about this but it is not ususally consided as a part of formal logic.

\mynote Rationale: What is universally true need not be communicated. Only that which is particular and the relation of the particular with the universal need be communicated. \commentary{Need to find a home for this -- I would be sad to lose it.}

\mynote An entity model systematically describes all that can be known of an entity
in terms of core functional relationships with universals and core functional relationships with particulars i.e. in terms of core attributes and core outgoing directional relationships.

\mynote But of all this which  may be known of an entity all we can communicate of an entity is its functional relationships with universals. Other relationships must be communicated indirectly via derivative (i.e. non-core) such relationships with universals. 

\mynote For this reason a systematic way of identifying and referencing each particular type  of entity in an entity model become significant if the model is to be used as any kind of data or message specification. There is some subtlety to identification and referencing as we will explain.

\subsection*{Communicating Relationships}
\mynote To communicate the value of a relationship is to communicate the identity of the relationship and  to identify (i.e. to communicated the identity of) the related entity. 

\mynote The related entity is a particular we are required therefore to have ways of identifying particulars and communicating these identifications.  
Only universals can be communicated so each type of entity is required 
to have one or more identifying attributes. 

\mynote Every entity may be identified by the values of related universals i.e. values that can be attributed to it. These attributions  may be core or derivative.

\mynote Since these relationships with other particulars themself need communication then ultimately every entity can be identified by a set of attributes each one of which may be core or may be derivative.  

\mynote
These attributes of a particular are called identifying attributes and subsequent use of these as attributes in a communication scheme in order to identify an entity as a related entity are known as referential attributes. 

\mynote For a long time now identifying attributes have been called key attributes and referential attributes have been called foreign key attributes. 

\mynote If you think that each outgoing directional relationship requires a distinct set of referential attributes to support its communication then you would be wrong. 
There is a possibility of `collapsed referentials' (a phrase coined in Shlaer and Long) whereby a single referential attribute may support communication of two or more distinct relationships. If you haven't understood this collapsed referentials concept then you have not  fully and completely understood the nature of data. 

\mynote In the presence of collapsed of referentials the number of bits of information required to communicate all the relationships of an entity is less that the sum of the bits required for their individual communication. 

\mynote the resolved set of attributes

\mynote conceptual core versus data core

\begin{oldtt}
\mynote To communicate the value of an attribute is to communicate its identity as an attribute and the universal that is its value. Both the identity of an attribute and its value are universals and communication of universals is a given i.e. does not require further explanation. 
\end{oldtt}

\begin{noteforfuture}
I imagine describing regular communication schemes... and why there can be a choice of schemes for relationships because of the choice of identifying features of target entity. 
\end{noteforfuture}
\begin{noteforfuture}
I imagine describing the communication of outgoing directional relationships of an entity by way of referentials and subsequently explaining that referentials can collapse which yields the term collapsed referentials.
\end{noteforfuture}