\documentclass[10pt,a4paper]{article}

%\usepackage{geometry}
\usepackage{pstricks}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{stmaryrd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{enumerate}
%\usepackage[shortlabels]{enumitem}
\usepackage{calc}
\usepackage{float} % for H option for tables and figures
\usepackage{changepage} % used for adjustwidth
\usepackage{framed}
\usepackage{wasysym}
\usepackage{ulem}   %added because \sout was not striking out 
                    % supports use uline and uwave also dashuline dotuline
\usepackage{epsf}
\usepackage{caption} % then I can use itemize in caption to figure

\newcommand{\SharedMacros}{../SharedMacros}
\newcommand{\SharedText}{../SharedText}

%\usepackage{enumitem} % to be able to [resume] an enumeration but clashes with [(i)] !!
\input{\SharedMacros/ermacros}
\input{\SharedMacros/ccategories.macros}
\input{\SharedMacros/gats.macros}
\input{\SharedMacros/metagat.macros}
\input{\SharedMacros/erdiagram}
\input{\SharedMacros/theorems.macros}
\input{\SharedMacros/general.macros.v2}
\input{\SharedMacros/alternate.nonbeamer.macros}
\input{\SharedMacros/indexedsets.macros.tex}
\input{\SharedMacros/barkerEllis.macros.tex}

\usepackage{hyperref}
\setcounter{equation}{0}

\newcommand{\ImagesFolder}{images}
\renewcommand{\erpictureFolder}[0]{images}
\newcommand{\handCraftedImagesFolder}{handCraftedImages}
\newcommand{\photosFolder}{photos}

\input{localMacros}


\title{Introduction to Entity Modelling}
\author{John Cartmell}
\begin{document}
\iffalse
\maketitle
\label{tableofcontents}
\tableofcontents
\label{contents}
\addcontentsline{toc}{section}{\nameref{tableofcontents}}

\subsection{How it is with Referentials}
\fi
Consider that an instance of the supervision relationship sketched out earlier in figure xxx might be comunicated like this
\begin{equation}
\label{xxx}
\mbox{\textit{\rdash{Maths} student \rdash{Brown} 
\ \uwave{is supervised by}\ \ \rdot{Philosophy} professor \rdot{Reece}}}
\end{equation}
and suppose that we already knew that the Maths department has a student named Smith and that Philosophy, likewise, has a professor called Reece. This message structure is appropraite and has no redundant parts if we assume a model like this

\begin{equation}
\scalebox{0.9}{\input{\ImagesFolder/shlaerLang-DeptStudentProfessor..collapsedReferential..diagram.tex}}
\end{equation}
in which the relationship d1 between a student and their department is modelled as an identifying fetaure of \textit{student} and likewise the relationship d2 is shown as an idenitfying feature of \textit{department}.

Notice that in the message () 
we see  that a student from one department is being supervised by a professor from 
a different  department. That means that this path from student to department mediated by supervising professor
\begin{equation*}
\scalebox{0.9}{\input{\ImagesFolder/shlaerLang-StudentProfessorDept..path..diagram.tex}}
\end{equation*}
is not subordinate to the trivial path consisting of the relationship\commentary{improve these diagrams sometime later after text improved and so validated}
\begin{equation*}
\scalebox{0.9}{\input{\ImagesFolder/shlaerLang-StudentDept..path..diagram.tex}}
\end{equation*}
Now we should consider some different scenariaos. Consider, instead of (),  
this message
\begin{equation}
\label{xxx}
\mbox{\textit{\rdash{Maths} student \rdash{Brown} 
\ \uwave{is supervised by}\ professor \rdot{Reece}}}
\end{equation}
and others of the same form.
There are two  scenarios in which it would be uniformly appropriate to convey r3 instances in this way.

1. if professors are uniquly identified by name but students are identified by name within the cvontextr of the departments they are within. 

2. if professors as well as students are unqiuly identified by name but only within the context of their departments AND there is a rule that says students must be supervised by professors from their own departments. This we should recognise as the case of a collapsed referential. We illustarte this case in figure xxx.


Finally, consider when messages of this form
\begin{equation}
\label{zzz}
\mbox{\textit{Student \rdash{Brown} 
\uwave{is supervised by} professor \rdot{Green}}}
\end{equation}
 might be appopriate. There are two cases. First case, obviously, when
students and professors can be idenitfied unquely by name. The pattern of relationships is this and no relationships are included amongst idenitfying features.

But there is also another case when (5) is appropriate even if the names of professors 
are only unique within the context of their departments.
If we already know that maths has student Brown,  and supposing that Maths and maybe 
Philosophy also has a professor Reece. Then the message structure
(zzz) is absolutely appropriate and of the correct form  if we know that students are supervised by professors in the departments they are within i.e. if diagram (xxx) is a dominance diagram (scope diagram???). 

Here the referential Maths is absent from tyhe message. Hence use of the term absent referential. Diagram xxx is an example of a general pattern which implies messages with absent referentials. A more general pattern is like this:

or like this.

. It doesn't help to try and give the most general possible paettern right now but you see the idea. There are two equivelnt paths that are comparable. One is a path of identifyingrelationships and the other is not or at least not in its entirety. 


Tabular data.

\setlength{\tabcolsep}{3pt} % Default value: 6pt
\footnotesize
\begin{tabular}{|l | l|l| l| l| l|}
Touchstone&Shakespeare&As You Like It&April-May 1975&Oxford Playhouse&Bob Hoskins
\end{tabular}



 \iffalse
 \input{Overview.tex}
 \input{Perspective.tex}
 \input{EntityModels.tex}
 \input{Relationships.tex}
 \input{ERDiagrams}
 \input{Paths.tex}
 \input{Attributes.tex}
 \input{IdentifyingFeatures.tex}
 \input{ReferencingEntities.tex}
 %\fi
 \input{CommunicatingRelationships.tex}
 %\iffalse
 \input{Scope.tex}
 \input{CommunicatingConjunctions}
 \input{CoreversusDerivative.tex}
 \input{DataModelling.tex}
 \input{TypeInheritance} 
 \input{StructuredEntityModelling.tex}
 \input{TheAbsolute}Can I talk about programming languages? 
terminology subtype and supertype Barker?
non-standard part of the SSADM method.
Use in car example from SSADM
\input{DistinctionBetweenCompositionAndReference}
\input{BarkerAirlineExample}
\input{Notes.tex}
\fi

\end{document}
