
\section{Relationships}
\label{Relationships}
%\iffalse
\mynote
A defining feature of entity modelling is the description
 (the modelling) of the relationships that may or may not  exist between things according as to the types of things that they are. 
\mynote 
We need to say something about this term `relationship' as it is used in
entity modelling.
As a starting point consider these two general definitions that I found online. 

\begin{erquote}
a relationship between two things is the way in which they are connected
\end{erquote}
and 
\begin{erquote}
a relationship is a connection, association, or involvement between two things.
\end{erquote}

At first reading these two dictionary entries seem to be in accord and indeed 
they both describe what would more specifically be called \textit{binary relationships} involving as they do exactly two things in relationship with each other.
A more careful reading reveals a significant difference between the two because \textit{friendship} as an abstraction satisfies the first definition but not the second whereas a particular friendship such as that between \textit{Don Quixote} and \textit{Sancho Panza} satisfies the second  but not the first. 
To distinguish the two meanings some authors use the term \textit{relationship type} for the first sense and reserve the term \textit{relationship} itself for the second. Here we will stick with the \commentary{re-jiggle}single word relationship, which, for us, can therefore have two distinct meanings\footnote{interpretations?}. If we need to emphasise the second of the two we may speak of instances of relationships rather than relationships. When speaking in this way we might say that there is a relationship called `friendship' between people and that there is an instance of this relationship between \textit{Don Quixote} and \textit{Sancho Panza}.

\mynote
Just for the record, in another dictionary we find 
\begin{erquote}
In logic and mathematics relationship is another name for relation.
\end{erquote}
and in yet another there is a definition of mathematical relation as a subset of the cartesian product of two sets 
(evidently, therefore, a definition of a binary relation)\footnote{Chambers definition here}\footnote{Need to say more here because this latter definition is a relation in a model of logic not in logic itself}. 

\mynote
Friendship,  marriage, adjacency, ownership and succession are all examples of binary relationships. There are binary relationships too  between parent and child, between teacher and student and between a planet and the star that it orbits.   
The following are also binary relationships:
being employed by, in the sense that a person might be employed by a particular employee,
managing, in the sense that an employee might manage a project. 
There are binary relationships too between a sales order and the customer making the order and between a debit card and  the bank that issues the card. As a final example, a banking transaction has binary relationships both with the account debited and with the account credited. 

\mynote The most important feature of relationships as the term is used  in this book is that
whether or not there exists a relationship instance between any two particular entities 
should be a matter of fact --- it should be a  black or white matter with no middle ground. 
For example between any particular Jack and any particular Jill we must take it that there either is or is not a friendship between them. There can be no equivocation or matter of degree --- either the relationship is or is not. This is a point of difference, by the way, between our terminology and that of Chen --- Chen's relationships may have attributes (properties) assigned to them, our relationships may not which is why they are, mathematically, binary relations. We will see that such (defintive, black-or-white) binary relationships are all we need for modelling what is and that neither is there a need for many-many relationships. 

\mynote 
Every binary relationship can be described from two different points of view. 
For the planet-star relationship the two points of view are
\begin{itemize}
\item every planet orbits a star,
\item every star may be orbited by planets.
\end{itemize}

In these descriptions the language is slightly ambiguous and questions arise.
Does every planet orbit exactly one star or may it orbit multiple stars? 
Are there stars without any planets orbiting them?

In an entity model ambiguities such as these are removed. 
The planet-star relationship would be expressed in more detail as
\begin{itemize}
\item every planet must be orbiting exactly one star,\footnote{You may disagree with this if your cosmology is more nuanced than mine.}
\item every star may be orbited by one or more planets.
\end{itemize}

Note that these two prescriptions describe the very same relationship from opposite ends. 
They are complementary points of view, a fact which  can be expressed like this:
\begin{itemize}
\item a planet is orbiting a star if and only if the star is orbited by the planet.
\end{itemize}
\subsection{Chen's Notation}
\mynote
Chen introduces diagrammatic representations of entities and relationships in his 
1976 paper. There is a great deal to find in this paper and it has been very influential and therefore is historically important but his diagram style many people find less practical than other notations which were subsequently published many of which were then summarised by Chen himself in his 1983 paper. 
Of these, our preferred notation is the Barker-Ellis notation and it is a variation on this that is used in this book. 
Because it is frequently seen, at least academically, it is appropriate
to describe the Chen notation for relationships and then to contrast with the Barker-Ellis style.
\mynote
In his introduction to entity relationship diagramming, Chen gives an example of an entity relationship diagram describing the types of entities relevant to the organisation of an imaginary manufacturing company.
In it he gives the example of a relationship between an employee and the department in which they work. In his diamond notation Chen draws the relationship like this:
\begin{center}
\scalebox{0.9}{\input{\handCraftedImagesFolder/ChenStyleDepartmentEmployee.tex}}
\end{center}

\noindent In this diagram, the numeric $1$ that labels the left hand line in this diagram indicates that every employee is in exactly one department and the corresponding annotation N on the right side indicates the fact that there are many employees within each department. \\

\noindent This relationship can therefore be described from  two complementary points of view as
\begin{align}
\label{employingText}&\mbox{every department may be employing one or more employees,} \\
\label{employedByText}&\mbox{every employee must be employed by exactly one department.}
\end{align}


\subsection{Barker-Ellis Notation}
\noindent In the notation used in this book, which is based on the Barker-Ellis notation,  this diamond and its two connectors are replaced by a single line. The fact that there
are multiple employees within a department is represented by drawing a crows foot at the employee end of the line. The relationship is drawn like so:
\definecolor{lightyellow}{cmyk}{0,0,0,0}

\begin{gather}
\label{employsRelationshipDiagram}
\raisebox{-1cm}{\input{\ImagesFolder/departmentEmployee.tex}}
\end{gather}

\noindent In this form the relationship can be read directionally
 --- left to right it
reads as  (\ref{employingText}), above,  
right to left it reads as (\ref{employedByText}). 

\noindent
We can think of there being two directional relationships here inseparably
connected (meta-related, in fact) as the poles of a magnet.
The right to left reading expresses a many-one directional relationship. It is significant that
in other contexts this would be described  as a functional relationship. We will return to this viewpoint shortly.


\subsection{Recap -- Relationships Looked at Directionally}

\noindent
To be clear, every binary relationship can be thought of as consisting of a pair of directional relationships each one of which is one view of it. \commentary{subsection rejig required}
\mynote Strictly speaking, directional relationships are not, as their name suggests, types of relationships, but are ways of looking at or considering binary relationships. Each directional relationship is a binary relationship viewed at or considered from one end or the other. 
Conversely, each binary relationship can be considered as a directional relationship in two different ways and this is the way that  binary relationships are represented in the entity modelling meta-model given later.

\mynote In the language of mathematics, a directional relationship which is many-one is said to be a functional relationship or, more simply, a function.
In this way,  the directed relationship 
shown in (\ref{employsRelationshipDiagram}), above, might be considered to be a function 
from the set of all employees to the set of all departments and might be
notationally represented by an arrow like this:
\begin{equation}
\label{employedByFunction}
employed\,by : employee \longrightarrow department
\end{equation}
\begin{noteforfuture}
give planet and star and maybe other examples here.
\end{noteforfuture}

\subsection{Mathematical Inverse}
\mynote Mathematically speaking, every function has an inverse which generally is many-valued and which therefore is represented as a separate function whose target is a power set i.e. a set of subsets. 
Accordingly the inverse to the function \textit{employed by} represented in 
(\ref{employedByFunction}) will in one mathematical dialect be represented by a function
\begin{equation}
\label{employingFunction}
employing : department \longrightarrow \mathcal{P}(employee).\footnote{Where $\mathcal{P}$ represents power set so that $\mathcal{P}(x)$ is the set of subsets of set $x$.}
\end{equation}

\noindent Let me recap this illustration. The function signatures (\ref{employedByFunction}) and (\ref{employingFunction}) express the same 
information as the words do in (\ref{employedByText}) and (\ref{employingText}).
Diagram (\ref{employsRelationshipDiagram}) encapsulates the same information and also expresses  the fact that (\ref{employingText}) and (\ref{employedByText}) are complementary 
i.e. that an employee works for a particular department if an only if that particular department employs the employee.
Finally, this complementarity may be expressed in the mathematical language of
sets and functions by asserting  function (\ref{employingFunction}) to be the inverse to function (\ref{employedByFunction}).

\subsection{Mathematical Terminology } \commentary{I think this and the styling subsection that follows needs to move up before the subsection describing many-many relationships.}
\commentary {this will cause two mathematical sections to be consequetive.}
\commentary{will need to rename this section. Maybe ``Mathematical Properties of Functions''}
\mynote
I am going to  describe how, in the Barker-Ellis style, 
the representations of functional relationships can be styled 
to give additional information about them but first 
 it is appropriate to bring to mind some further mathematical notation.

\mynote Functional relationships in mathematics that are total are usually depicted by the arrow symbol $\longrightarrow$ so the assertion that $f:A \longrightarrow B$ in mathematics generally means that $f$ is a functional relationship from $A$ to $B$ which is total i.e. defined everywhere.\footnote{The more abstract use of $\longrightarrow$ in category theory is a notable exception and is entirely germane as we will discuss later.} 
For a functional relationship that is not known to be total 
then a partial arrow symbol $\longrightharpoonup$ is used instead so, in the case of such a relationship, it might be asserted that $f:A \longrightharpoonup B$.
Finally, the notation $f: A \longtwoheadrightarrow B$ can be used to assert that $f$ is a total functional relationship that is surjective.\footnote{A function f from A to B is said to be surjective iff TBD.}

\noindent All of these notations have equivalents in the Barker-Ellis style, as we will now see.

\subsection{Barker-Ellis Styling of Relationships}
\mynote In the  Barker-Ellis style, in the most general case, functional relationships of A to B
are shown like this:
\begin{center}
A\,\barkerEllisA\,B
\end{center}
\noindent for example as here:
\begin{equation}
\label{shlaerMellorDeptStudentProfessor0A}
\raisebox{-0.45cm}{\input{\ImagesFolder/shlaerMellorDeptStudentProfessor0A.tex}}
\end{equation}
\commentary{add to this?}
If however a functional relationship is known to be total then the left half line is drawn as a solid line instead of a dashed line like so \barkerEllisC
% \begin{center}
% A\,\barkerEllisC\,B
% \end{center}
for example as here: 
\begin{equation}
\label{shlaerMellorDeptStudentProfessor0C}
\raisebox{-0.8cm}{\input{\ImagesFolder/shlaerMellorDeptStudentProfessor0C.tex}}
\end{equation}
In either case if the map is known to be surjective then the right hand half line is solidified and looks as this \barkerEllisB or \barkerEllisD
% \begin{center}
% A\,\barkerEllisB\,B
% \end{center}
or this
% \begin{center}
% A\,\barkerEllisD\,B
% \end{center}
as for example in the relationship between a professor and the department they are a part of:
\begin{equation}
\label{shlaerMellorDeptStudentProfessor0B}
\raisebox{-0.45cm}{\input{\ImagesFolder/shlaerMellorDeptStudentProfessor0B.tex}}
\end{equation}
\footnote{In these examples we have assumed (and then documented) that departments \textbf{must} have professors and \textbf{may} have students.}

If in any of these cases it is known in addition  that the map is injective i.e. that two distinct instances of the A type cannot be related to the vary same instance of the B type then the crowsfoot is omitted and so the relationship is depicted by a line that
may be dashed or half dashed as the case might be, the possibilities being:
A\,\barkerEllisE\,B, 
A\,\barkerEllisF\,B, 
A\,\barkerEllisG\,B or 

A\,\barkerEllisH\,B.

\commentary{maybe omit these four depictions}
Note that in this final case the relationship in question establishes a 1-1 correspondence between types A and B.

\begin{noteforfuture}
have an example of some sort here with a variety of cardinalities or refer forward to the manufacturing example?
\end{noteforfuture}

\subsection{Diagramming Relationships}
The next thing to point out is that multiple relationships can be depicted on a single diagram and the relationship lines can be routed about as convenient for the drawing. This is a characterising feature of entity relationship diagramming.
For example we can depict all three relationship 
(\ref{shlaerMellorDeptStudentProfessor0A}), 
(\ref{shlaerMellorDeptStudentProfessor0C}) and 
(\ref{shlaerMellorDeptStudentProfessor0B}) on a single diagram in figure \ref{shlaerLangDiagram}.
\begin{erboxedFigure}{H}
{shlaerLangDiagram}
{This is a representation in Barker-Ellis style of a diagram  previously drawn 
in the Shlaer-Melor style by Neil Lang 
 in a  little known paper written  in conjuntion with Sally Shlaer. In a later section we will explain 
 a significant feature of this diagram.}
\input{\ImagesFolder/shlaerMellorDeptStudentProfessor0..diagram.tex}
\end{erboxedFigure}

 In this book there will be many of these smaller examples as shown in figure \ref{shlaerLangDiagram} and we can imagine  these to be fragments of larger diagrams  representing substantial entity models such as 
 might represent entire educational institutions or operational businesses. 
In a later section on structured entity modelling we will describe additional meanings that we ascribe to relationships dependent on their orientations on the diagrams. These are then the ascribed meanings in the remainder of this book. 

\subsection{Many-Many Binary Relationships}
In Chen's paper the relationship between employees and projects (whose meaning is that
an employee is assigned to work on a project) is depicted like this.
\begin{gather}
\label{projectWorkerChenStyle}
\raisebox{-0.9cm}{\scalebox{0.9}{
\input{\handCraftedImagesFolder/ChenStyleProjectWorker.tex}
}}
\end{gather}
On this diagram there is an annotation $N$ which documents that an employee may work on many projects (we are to understand that $N >= 0$). There is also an annotation $M$ to document that
a project has many employees assigned to it, $M$ in fact, for some $M >= 0$. \\
\noindent Such a relationship is called a many-many relationship.  \\
\noindent In Barker-Ellis notation which is the basis for the notation used in this book the diamond and its two connectors are replaced by a single line. The fact that the relationship is many-many is represented in the diagram by having a crows foot at each end of the line like so:

\begin{gather}
\label{employeeProjectWorker}
\raisebox{-1cm}{
\input{\ImagesFolder/employeeProjectWorker.tex}
}
\end{gather}

\noindent As before, the relationship is labelled at both ends. Each text label gives the name of the relationship from the point of view of the type of entity at that end of the line and can be used to phrase the relationship:
\begin{align}
\label{assignedToText}&\mbox{every employee may be assigned to one or more projects,} \\
\label{assignedText}&\mbox{every project may be assigned one or more employees.}
\end{align}
\newpage
\subsection{Elimination of Many-Many Relationships}
\noindent In fact, many-many relationships are infrequently used in entity models and never used at all when modelling for the purpose of data specification.
Consider that we can make a thing out of the fact of the assignment of an employee to a project. This means that we can entify the many-many relationship and remodel the relationship (\ref{employeeProjectWorker}) like this:
\begin{gather}
\label{employeeProjectWorkerMediated}
\raisebox{-0.3cm}{\scalebox{0.95}{\input{\ImagesFolder/employeeProjectWorkerMediated.tex}}}
\end{gather}
\noindent
In (\ref{employeeProjectWorkerMediated}) the  many-many relationship depicted in (\ref{employeeProjectWorker}) is now mediated by the conjunction of  type
\textit{project assignment} and two functional relationships. 
Because we can `entify' many-many relationships in this way and mediate them by entity types and functional relationships they can be eliminated from entity modelling. 
When we do this only functional relationships (and their inverses) remain. Significantly, only functional relationships have direct representations in data, 
however this data is represented\footnote{This is true whether the data is held relationally or non-relationally in databases, or represented in XML, or in messages such as described by IDL 
or in other program structures}, and for this reason it is usual to eliminate many-many relationships from entity models in favour of functional relationships. After elimination, since the remaining relationships are functional, in one direction or the other, so it is that in most entity models, types are related solely by functional relationships and their inverses. 
In this form, \textit{entity modelling models the world in functional relationships}, a fact that 
deserves to be articulated but which usually isn't.

\mynote
If we compare (\ref{employeeProjectWorkerMediated}) 
with the representation \textit{a la} Chen in (\ref{employeeProjectWorker})
then we see that the diamond in (\ref{employeeProjectWorker}), which is annotated to express that it represents a many-many relationship.
is in the Barker-Ellis notation in (\ref{employeeProjectWorkerMediated})  replaced by an entity type and two functional relationships. 
However in this form in (\ref{employeeProjectWorkerMediated}) we are missing
something as yet compared to the representation in (\ref{employeeProjectWorker}).
This is rectified by drawing a bar across each of the two functional
relationships like this:
\begin{gather}
\label{employeeProjectWorkerMediatedIBarred}
\raisebox{-0.3cm}{\scalebox{0.95}{\input{\ImagesFolder/employeeProjectWorkerMediatedIBarred.tex}}}
\end{gather}

These bars signify, in Barker-Ellis,  that the type \textit{project assignment} represents a many-many relationship.
The precise meaning of this is that
\begin{center}
\begin{parbox}{9cm}{
	\textit{
	for each employee and for each project there is at most one assignment
	of the employee to the project.
	}
}
\end{parbox}
\end{center}
This is the implication of the Barker-Ellis notation once bars are drawn across the relationships. 

\mynote For the sake of completeness, I should mention that in situations
such as (\ref{employeeProjectWorkerMediatedIBarred}) above, the mediating entity or indeed its type is often referred to as an \textit{intersection entity (type)}. 
Other authors refer to it as an \textit{associative entity type}. Later we come back to a slightly
more general use of bars across relationships.

\mynote In summary diamonds and their connectors in the Chen notation are replaced 
in the Barker-Ellis notation
by entity types and barred functional relationships. 
\mynote
Finally, note that barred functional relationships play a wider role in Barker-Ellis than we have discussed here and we shall describe this wider role in a later section.

\subsection{Ternary Relationships}
I should also draw attention to the fact that Chen's example which we show in figure \ref{ChenManufacturingExample} contains an 
example of a ternary relationship between a project, a supplier and a part which is shown  like this:

\begin{equation}
\raisebox{-1.0cm}{\scalebox{0.85}{\input{\handCraftedImagesFolder/ChenStyleSupplierProjectPart.tex}}}
\end{equation}

In the Barker-Ellis style there are no ternary relationships nor are any needed since ternary relationships can be represented as a combination of barred binary relationships and entity types. 

\mynote Chen's ternary relationships can be represented instead in the Barker-Ellis style by following the prescription given earlier\textit{ diamonds and their connectors in the Chen notation are replaced 
in the Barker-Ellis notation
by entity types and barred functional relationships.}\commentary{smooth out this text}

\mynote 
Translating Chen's \verb'SUPP-PROJ-PART' construct directly into Barker-Ellis notation we get a fragment like this\commentary{why such an odd non-symetric layout?}
\begin{center}
\scalebox{0.95}{\input{\ImagesFolder/ChenSupplierProjectPartBarkerStyle.tex}}
\end{center}


\subsection{Mathematical Terminology - Jointly Injective}
\label{JointlyInjective}
\mynote
We can give a mathematical characterisation of the meaning of the barred relationships in the above examples.
We need to use the following definition.
Suppose that $A$, and $B_1,...B_n$ are sets and suppose that \foreachi, $f_i$ is a function, $f_i:A_i \morph B_i$ 
then the collection of functions $f_1,...f_n$ is said to be \textit{jointly injective}
iff for each $x,x' \in A$, if \foreachi, $f_i(x)=f_i(x')$ then $x=x'$.

Recall that our many-one relationships can be thought of as functions, though these might be partial.
The meaning of placing bars across a number of outgoing many-one relationships is to signify that thought of as a collection of functions, this collection is jointly injective. In Barker-Ellis barred relationships are always total functions i.e. the bars are drawn across solid lines not dashed lines.
\mynote
It follows from the definition, by the way, that if the jointly injective functions $\fn$, be combined
to map the set $A$ into the cartesian product $B_1 \times B_2 ... \times B_n$
then the mapping is injective. This completes the picture corresponding as it does to $A$ being isomorphic to an $n$-ary relation,
Chen-diamond like, connected to $B_1,...B_n$.

\subsection{Chen's Relationship Types and English}
\mynote
Chen in his 1983 paper equates his relationship types (diamonds) are suggested by
the occurrence of verbs in English in statements like\commentary{reword required}
\begin{center}
example here
\end{center}

The general principle should be that these same constructs should be 
represented by entity types in Barker-Ellis style diagrams which a characterised by having
all their outgoing relationships being  annotated with bars.
For some authors such an entity type is said to be an intersection entity (type)
and for some it is said to be an associative entity (type). Mathematically it represents an n-ary relation but I hesitate to use that term because of the overload of terms relation and relationship. 




\subsection{Barker-Ellis Phraseology} 
\mynote 
Consider the binary relationship between a moon and the planet it is orbiting. 
I will represent it in the  Barker-Ellis style in this way
\begin{gather}
\label{planetMoon}
\raisebox{-0.65cm}{
\input{\ImagesFolder/planetMoon.tex}
}
\end{gather}
In this notation, there isn't a single name that is given to
 the binary relationship but instead there are two relative terms `orbiting' and `orbited by' which describe each of the parties in the relationship from the point of view of the other. This dual naming of binary relationships is the way that relationships are named in an entity model and the diagram ties these names together as we have discussed.

\mynote
In the Barker-Ellis notation the idea 
is to choose these relationship labels
carefully so that a certain phraseology, i.e. a particular templated form of expression,
can be used to produce textual descriptions of the relationship which are both meaningful and grammatical.
We have already seen an example of this in the rendering of
relationship (\ref{employsRelationshipDiagram}) by texts 
(\ref{employingText}) and (\ref{employedByText}), above. 
Another example is the phrasing of relationship (\ref{employsRelationshipDiagram})
by texts (\ref{assignedToText}) and (\ref{assignedText}). \commentary{Have any other examples been given hithertofore?} 

Now from (\ref{planetMoon}) applying the Barker-Ellis  phraseology and reading the 
relationship left-to-right we read the relationship as telling us something about every planet, because the relationship line is dashed representing optionality then we phrase that every planet \textit{may be} 
in a relationship,
because there is a crowsfoot on the opposite end we phrase that the relationship is to 
\textit{one or more} moons.  In this way, from the standard phraseology, we obtain a textual description:
\begin{align}
&\mbox{every planet may be orbited by one or more moons} 
\end{align}
This is grammatical only because I 
because I chose the label `orbited by' with care and with foreknowledge of the  phraseology.
 
Applying the same rules but now reading  the relationship from right-to-left and interpreting the solidity of the line as 
 meaning that every moon \textit{must be} in relationship with a planet we get:
\begin{align}
&\mbox{every moon must be orbiting  one and only one planet} 
\end{align}
Without foreknowledge of the phraseology I might instead have chosen the word `orbits' for the name of my directional relationship but if I had so chosen then applying the phraseology would have yielded an ungrammatical sentence.\footnote{I would have got "Every moon must be orbits one and only one planet" which is ungrammatical.}
%\fi
\subsection{Phrasing Relationship Instances}
\mynote
When relationships are named with the 
Barker-Ellis phraseology in mind  then a second phraseology can be used to describe instances of the relationships. 
As with the first phraseology, this second, instance level, phraseology,
 can be used in either direction
 --- in our example it can be used from the point of view
of the planet or from the point of view of the moon.
In the right-to-left direction the planet is the subject and the relational term is
`orbited by' and, accordingly, the relationship of planet Jupiter with one of its moons can be phrased like this
\begin{itemize}
\renewcommand\labelitemi{--}
\item Jupiter is orbited by Io
\end{itemize}

This same relationship instance can be phrased in the other direction, right-to-left wrt to diagram (\ref{planetMoon}) and in this phrasing the moon Io is the subject and the relational term is `orbiting' so we get
\begin{itemize}
\renewcommand\labelitemi{--}
\item Io is orbiting  Jupiter.
\end{itemize}

If in  many such facts as these
\begin{itemize}
\renewcommand\labelitemi{--} 
\item Saturn is orbited by Titan, 
\item Mars is orbited by Phobos, 
\item Jupiter is orbited by Europa,
\end{itemize}
and others like them are to be conveyed verbally then it is likely we will contrcat the detail so that we say something like
\textit{Saturn is orbited by Titan; Mars by Phobos; Jupiter by Europa}
If instead  these facts are to be conveyed in a document then
it is likely they will be conveyed in tabular form something like this
\begin{center}
\begin{tabular}{|l | l|}
\hline
planet & orbited by \\
\hline\hline
Jupiter  & Io \\
\hline
Saturn   & Titan    \\
\hline
Mars     & Phobos    \\
\hline
Jupiter  & Europa         \\
\hline
%Jupiter  & Io         \\
%\hline%
%Saturn   & Rhea    \\
%\hline
\vdots    & \vdots  \\
\hline
\end{tabular}
\end{center}
In the first row, 
the column headings refer  back to the relationship (\ref{planetMoon})
and subsequent rows, in the context of the column headings, communicate relationship instances. 
What we are looking at here of course is a data table and an important aspect of this book is the synergy between  entity models and data structure.

