\documentclass[10pt,a4paper]{article}

%\usepackage{geometry}
\usepackage{pstricks}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{stmaryrd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{enumerate}
\usepackage{calc}
\usepackage{float} % for H option for tables and figures
\usepackage{changepage} % used for adjustwidth
\usepackage{framed}
\usepackage{wasysym}
\usepackage{ulem}   %added because \sout was not striking out 
                    % supports use uline and uwave also dashuline dotuline
\usepackage{epsf}
\newcommand{\SharedMacros}{../../SharedMacros}
\newcommand{\SharedText}{../../SharedText}

%\usepackage{enumitem} % to be able to [resume] an enumeration but clashes with [(i)] !!
\input{\SharedMacros/ermacros}
\input{\SharedMacros/ccategories.macros}
\input{\SharedMacros/gats.macros}
\input{\SharedMacros/metagat.macros}
\input{\SharedMacros/erdiagram}
\input{\SharedMacros/theorems.macros}
\input{\SharedMacros/general.macros.v2}
\input{\SharedMacros/alternate.nonbeamer.macros}
\input{\SharedMacros/indexedsets.macros.tex}
\input{\SharedMacros/barkerEllis.macros.tex}

\usepackage{hyperref}
\setcounter{equation}{0}

\newcommand{\ImagesFolder}{../images}
\renewcommand{\erpictureFolder}[0]{../images}
\newcommand{\handCraftedImagesFolder}{../handCraftedImages}
\newcommand{\photosFolder}{../photos}

\input{../localMacros}

\title{patterns of identifying relationships}
\author{John Cartmell}
\begin{document}
\maketitle



\newcommand{\thumbnailscale}{0.4}

\subsection{Understanding Patterns of Identifying Relationships}


To make patterns of identifying relationships easier to recognise we will draw diagrams specifically to represent identification of a certain type of entity in which the only relationships are idenitfying relationships and so that these relationshiops are drawn left to right. The above charcater playright diagram shown in this way will look like this:
\begin{equation*}
\mbox{diagram with character --- play --- playwright left to right}
\end{equation*}

There are recognisable patterns, then, of idenitfying relationships that emanate from an entity type on a diagram.
Sometimes as here there is a single path of identifying relationships 
\begin{equation*}
\scalebox{\thumbnailscale}{
\input{\handCraftedImagesFolder/thumbnailPathIdentifying}}
\end{equation*}

sometimes there are branching paths
\begin{equation*}
\scalebox{\thumbnailscale}{
\input{\handCraftedImagesFolder/thumbnailBranchingPathsIdentifying}}
\end{equation*}

and sometimes the branching paths come back together again in the sense that there is a type that is common to both paths as in this pattern

\begin{equation*}
\scalebox{\thumbnailscale}{
\input{\handCraftedImagesFolder/thumbnailComparablePathsIdentifying}}
\end{equation*}

In this case there are two (there might be more) paths of identifying reltionships which are comparable. This is an important case and it is most significant wherther or notr these comparanble paths are equivalent as we shall see. We will see examples of all these differing patterns in the next section in which we explore their significance.




and the identification of production entititees will be highlighted in this way like this.


\subsection{Making the pattern of identifying features easier to see}
To give ourselves more efficient use of space on the page we then move on and draw these two diagrams in a slightly different way like this

and this.

What I have done is to introduce a representation for the type of all universals on the right of the diagram and to represent attributes as relationships with this type. I have broken with the tradition of having a complete distinction bbweteen attributes and relationships at this point. Not only does this give a more striking depiction of the way that identifying features combine it helps very much in the next sections on referencing and data representation of entitites.

\begin{tabular}{| p{5.5cm} | l |}
\hline
single identifying relationship &
\raisebox{-0.35cm}{\scalebox{\thumbnailscale}{
\input{\handCraftedImagesFolder/thumbnailSingleIdentifying}}
} \\
\hline
path of identifying relationships &
\raisebox{-0.35cm}{\scalebox{\thumbnailscale}{
\input{\handCraftedImagesFolder/thumbnailPathIdentifying}}
} \\
\hline
branching paths of identifying relationships &
\raisebox{-0.85cm}{\scalebox{\thumbnailscale}{
\input{\handCraftedImagesFolder/thumbnailBranchingPathsIdentifying}}
} \\
\hline
comparable paths &
\raisebox{-0.65cm}{\scalebox{\thumbnailscale}{
\input{\handCraftedImagesFolder/thumbnailComparablePathsIdentifying}}
} \\
\hline
\end{tabular}

blah blah blah blah\\



\end{document}
