
\subsection{Intro}
I am wrong about property awR.4 holding in the category of tight zigzags. Because this property doesn't hold I cannot prove that pHom preserves restrictions. 

Comparing with Cockett's work I need a new definition that applies specifically to the catgewory of sticklebacks ---this definition isn't general purpose --- it doesn't apply to all categories with support.

$pHom_C(a,-)$ is a restriction functor if \catcw is a restriction category but not generally on a catgeory with support.     

Rather than update the preceding pages as at today 9th March 2026 I will 
sketch out  in this section the definition of the $pHom_\catc(a,-)$ replacement functor --- dare I call it $rHom_\catc(a,-)$? 

I will sketch out the properties I need of it additional to it being a
restriction functor. 

Of course the functors I really want are the compositions
of the embedding from \catcw into $\Stk$ compsoed with
such $rHom_\catc(a,-)$.
\subsection{Definition rHom}
For every object $a$ of range category \catcw let
\[
rHom_\catc(a,-): \Stk \morph \Par
\]
be the  functor
defined as follows. 

$rHom(a,x)=Hom(a,x)$.


If $G: x \morph y$ in $\Stk$ is the stickleback
$\zigzagTuple{g}{m}$ then the partial function
$rHom(a,G): rHom(a,x) \morph rHom(a,y)$
is the partial function defined at a zigzag 
$F=\zigzagTuple{f}{n}\in rHom(a,x)$ by
\begin{align*}
rHom(a,G)(F) &= F \circ G  &&\mbox{ if $\overline{f_n \circ g_0} = \overline{f_n}$,} \\
             &=\bot        && \mbox{otherwise.}
\end{align*}

\begin{aside}
When I wrote the above definition I had not yet proved that $rHom$ so defined was functorial. Now I believe I have proved this and to prove it I developed some good looking lemmas regarding range categories that I refer to in the daybook (11 April 2026) as lemma A and lemma B. These lemmas look like adjoint relationships. Lemma A, at least, lifts to the category of sticklebacks (at which point in day book I refer to as lemma AS) and this lifted lemma enables me to show that $rHom$ is functorial. I think therefore that there is a characterisation of the category of sticklebacks (of a range catgeory) as a category with support and cosupport and with some additional
property (expressed by the lifted lemma) which implies that $rHom$ is a functor and a restriction functor at that.
\end{aside}

\begin{aside}
I need uinderstand why previous definitions of partial hom functor fail and why this one works. Otherwise I might be kidding myself.
\end{aside}

\begin{aside}
 I don't \textbf{need} this $rHom$ functor (I can proceed more closely to Cockett) but I think it is probably for the greater good so I do need to work through this. It is part of a bigger picture.
\end{aside} 

\subsection{Simplify and follow Cockett more closely}
If \catcw is a range category, define a functor $K_a: \catc \morph \Par$ by defining for each object $x$ of \catc, $K_a(x)=Hom_{\Stk}(a,x)$ and by defining $K_a(j)$, for $j:y_1 \morph y_2$ in \catc, to be the partial function  defined at a zigzag 
$F=\zigzagTuple{f}{n}\in Hom_{\Stk}(a,x)$ by
\begin{align*}
K_a(j)(F) &= F :: \tuple{j}  &&\mbox{ if $\overline{f_n \circ j} = \overline{f_n}$,} \\
             &=\bot        && \mbox{otherwise.}
\end{align*}

Note that the condition $\overline{f_n \circ j} = \overline{f_n}$ ensures
that $F :: \tuple{j}$ is a tightg zigzag and therefore that
$F :: \tuple{j}= F \circ \tuple{j}$.

$K_a$ is a range functor by arguments made by Cockett.

The RR.5 condition is required to ensure that every tight
zigzag has a unique shortening. 




\subsection{Requirements for $K_a$}

From memory require:
\begin{enumerate}[(i)]
\item If $f: a \morph x$ in \catcw then 
\[
K_a(f)(\tuple{\bar{f}})=\tuple{f}
\]
\item If $f: a \morph x$ and $g: a \morph x$ in \catcw then 
\[
K_a(f)(\tuple{\bar{f}\circ\bar{g}})=\tuple{\bar{g} \circ f}
\]
and, therefore,
\[
K_a(g)(\tuple{\bar{f}\circ\bar{g}})=\tuple{\bar{f} \circ g}
\]
\end{enumerate}

This is enough to get me the first representation lemma 2A.

\begin{aside}
I stated out with $S_a$ as the stickelback functor $K_a$.
But ... in the proof I get a clash of letter with $S$ being used as 
a functor $S_a$ which I want to extend to a functor $S'_a$ and also
being used as a sketch $S$ which is extended to a sketch $S'$.

Arguably $S$ for sketch takes precedence over Crockett's $S$ for 
stickleback functor.
What letters to choose. Use $G$ (graph) for sketch??. Use $E$, entity model, or $ER$, or $M$ model, or $P$ for presentation.

What letter to use for the stickleback functor $S_a$. Could be $K_a$. 
K for stic\textbf{K}leback. Could be $D_a$ because type $a$ is populated with $\textbf{D}ata$. Except at the moment I have $D \in \reqt$.
\end{aside}
