Denote by \Par the range category of sets and partial functions
 and let
\FinPar be the range category of finite sets and partial functions. 

\subsection{The partial Hom functor $HomP$}

\subsubsection{Definition of HomP}
If \catcw is a category with support then \commentary{Tidy this definition please.}
let $HomP_\catc(a,-): \catc \morph \Par$ be the  functor
defined as follows. $HomP(a,x)=Hom(a,x)$.
If $j: x \morph y$ then partial function $HomP(a,j): HomP(a,x) \morph HomP(a,y)$
is defined\footnote{
In earlier notes I defined HomP as follows.
\begin{align*}
HomP(a,j)(k) &= k \circ j &&\mbox{ if $k \circ \bar{j} = k$,} \\
             &\bot        && \mbox{otherwise.}
\end{align*}
In a restriction category this is definition is equivalent to the one given but not in an arbitrary category with support.
} as follows:

\begin{align*}
HomP(a,j)(k) &= k \circ j  &&\mbox{ if $\overline{k \circ j} = \overline{k}$,} \\
             &=\bot        && \mbox{otherwise.}
\end{align*}
\begin{notebox}
CHECK that  I used this fact today, 28 Jan 2026, in order to prove
that a shortening of a tight zigzag is tight. Handwritten notes.
\end{notebox}

It follows from lemma  \lref{supportLemma} (\lref{RHomFunctorSubLemma}) that HomP, so defined, is a functor.
\subsubsection{$HomP$ and Supports}
In a restriction category $HomP(a,-)$ preserves the restriction operator
and thus is a restriction functor. 
In a category with support it isn't the case that the support operator is preserved unless
the additional awR.4 property holds:
\begin{lemma}
\llabel{originalHomPSupportLemma}
If \catcw is a category with \newt{awR.4} support 
and $a$ is an object of \catcw then
$HomP_\catc(a,-)$ preserves the support operator
i.e.  
for all morphisms $j: x \morph y$ in \catc,
\begin{equation}
HomP_\catc(a,\bar{j})=\overline{HomP_\catc(a,j)}.
\end{equation}
\end{lemma}
\begin{proof}
Suppose such  $j: x \morph y$ in \catc.
In \Par we have the following partial functions
\[
\paralleldiag{HomP_C(a,x)}{HomP_C(a,x)}{HomP_C(a,\bar{j})}{\overline{HomP_C(a,j)}}
\]
We need show that for all $k \in HomP_C(a,x)$,
\[
HomP_\catc(a,\bar{j})(k)=\overline{HomP_\catc(a,j)}(k).
\]

To this end,
first consider that  the left hand function, $HomP_\catc(a,\bar{j})(k)$, is defined by
\begin{align*}
HomP(a,\bar{j})(k) &= k \circ \bar{j}  &&\mbox{ if $\overline{k \circ \bar{j}} = \overline{k}$,} \\
             &=\bot        && \mbox{otherwise,}
\end{align*}
which  simplifies using wR.4 to
\begin{align*}
HomP(a,\bar{j})(k) &= k \circ \bar{j} &&\mbox{ if $\overline{k \circ j} = \overline{k}$,} \\
             &=\bot        && \mbox{otherwise.}
\end{align*}
and so using \newt{awR.4} to
\begin{align*}
HomP(a,\bar{j})(k) 
            &= k \circ \bar{j} &&\mbox{if $k \circ \bar{j} = k$,} \\
            &=\bot        && \mbox{otherwise,}
\end{align*}
i.e. to
\begin{align*}
HomP(a,\bar{j})(k) 
            &= k          &&\mbox{if $k \circ \bar{j} = k$,} \\
            &=\bot        && \mbox{otherwise.}
\end{align*}

On the otherhand from the definition of $HomP_\catc(a,j)(k)$ it follows that
\begin{align*}
\overline{HomP(a,j)}(k) 
            &= k         &&\mbox{if $\overline{k \circ j} = \overline{k}$,} \\
            &=\bot        && \mbox{otherwise,}
\end{align*}
which simplfies, using \newt{awR.4}, to defining
\begin{align*}
\overline{HomP(a,j)}(k) 
            &= k  &&\mbox{if $k \circ \bar{j} = k$,} \\
            &=\bot             && \mbox{otherwise.}
\end{align*}
We see the two functions, $HomP_\catc(a,\bar{j})(k)$ and $\overline{HomP_\catc(a,j)}(k)$, are identical, as required.
\end{proof}

From this lemma and from lemma \ref{restrictioncatlemma}(\ref{CHECKLemma}) it follows that:
\begin{corollary}
\llabel{HomPRestrictionCorallary}
If \catcw is a restriction category  and $a$ is an object of \catcw 
then the functor $HomP_\catc(a,-)$ preserves the restriction operator i.e.
for all $j:x \morph y$ in \catc
$$HomP_\catc(a,\bar{j})=\overline{HomP_\catc(a,j)}$$
\end{corollary}

\begin{lemma}
\llabel{homPfaithfullemma}
If \catcw is a category with support
and $a$ is an object of \catc, 
if for some object $x$ of \catcw and
$j_1, j_2: a \morph x$ in \catcw 
then if 
\begin{equation}
\label{HomPIdentity}
HomP(a,j_1)=HomP(a,j_2)
\end{equation}
then $j_1=j_2$
\end{lemma}

\begin{proof} 
\commentary{This proof goes around the houses a bit. I haven't seen how to tighten it.}
First we show that 
\begin{equation}
\label{faithstepone}
j_1 = \bar{j_1} \circ j_2
\end{equation}
which we do as follows
\begin{align*}
j_1 &= HomP(a,j_1)(\bar{j_1}) &&\mbox{by definition of $HomP$, } \\
   &= HomP(a,j_2)(\bar{j_1}) &&\mbox{by assumption (\ref{HomPIdentity}), } \\
   &= \bar{j_1} \circ j_2    
                     &&\mbox{\parbox[t]{6cm}{by definition 
                                        of $HomP$ and because
                                        $HomP(a,j_2)(\bar{j_1})$ 
                                        cannot be undefined since
                                         $HomP(a,j_1)(\bar{j_1})$
                                         is not undefined . }..} 
  \end{align*}
By symmetry we also have
\begin{equation}
\label{faithsteptwo}
j_2 = \bar{j_2} \circ j_1.
\end{equation}
Next we show that
\begin{equation}
\label{faithstepthree}
\bar{j_2} \circ j_1 = \bar{j_1} \circ j_2
\end{equation}
which we do as follows:
\begin{align*}
\bar{j_2} \circ j_1 
    &= \bar{j_1} \circ \bar{j_2} \circ j_1 &&\mbox{by R.1 and R.2,} \\
    &= HomP(a,j_1)(\bar{j_1} \circ \bar{j_2}) 
                              && \mbox{from definition of HomP,}\\
    &= HomP(a,j_2)(\bar{j_1} \circ \bar{j_2}) 
                              && \mbox{by (\ref{HomPIdentity}),}\\
    &= \bar{j_1} \circ \bar{j_2} \circ j_2 
                              && \mbox{from definition of HomP,}\\
    &= \bar{j_1} \circ j_2    &&\mbox{by R.1.}
\end{align*}
Finally from (\ref{faithstepone}),
(\ref{faithsteptwo}) and (\ref{faithstepthree}),
$j_1=j_2$ follows, as required.
\end{proof}

\subsubsection{$HomP$ and Ranges}

\begin{observation}
If \catcw and \catcpw are  categories with support 
and $F: \catc \morph \catcp$ is a functor that preserves the support operation,
 if morphism $f: a \morph b$ in \catcw
has a partial section $f_s: b \morph a$ then 
the morphism $F(f) : F(a) \morph F(b)$ in \catcpw has section
$F(f_s)$ and therefore $F$ preserves the pseudo range operator 
i.e. $F(\hat{f}) = \widehat{F(f)}$ 
\end{observation}

The converse is also true \commentary{This is now a bit of a textit{non sequitur}.}
\begin{lemma}
If \catcw is a range category and $HomP(a,-)$ is a range functor for all objects a of \catcw
then every morphism $f$ of \catcw has a partial section. 
\end{lemma}
\begin{proof} 
First consider that 
$\widehat{f}=id_b \circ \hat{f}=Hom(a,\hat{f})(id_b)$ 
and therefore that $\widehat{f}$ is in the image of $Hom(a,\hat{f})$.
Which is to say that $\widehat{f}$  is in the image of the function
$Hom(b,f)$ because $\reallywidehat{Hom(b,f)} = Hom(b,\hat{f})$ from the assumption that the range operator is preserved by $HomP(a,-)$. Therefore there is
at least one such $g: b \morph a$ such that $Hom(a,f)(g) = \hat{f}$ i.e. such
that
\begin{equation}
\label{gCondition}
g \circ f = \hat{f}.
\end{equation}

For any such $g$ we can show that the morphism $\hat{f} \circ g$ is a partial section of $f$
by showing  both conditions required of a partial section 
\begin{equation}
\label{sectionReqOne}
f \circ \hat{f} \circ g \circ f = f
\end{equation}  
and
\begin{equation}
\label{sectionReqTwo}
\hat{f} \circ g \circ f = \overline{\hat{f} \circ g}
\end{equation}
(\ref{sectionReqOne}) is the case because
\begin{align*}
f \circ \hat{f} \circ g \circ f &=  f \circ \hat{f} \circ \hat{f}     
                                               && \mbox{by (\ref{gCondition}),} \\
                                &=f       
                                               && \mbox{by RR.2.} 
\end{align*}
(\ref{sectionReqTwo}) is the case because
\begin{align*}
\hat{f} \circ g \circ f  &= \hat{f} \circ \hat{f}       
                                               && \mbox{by (\ref{gCondition}),} \\
                         &= \hat{f}     
                                               && \mbox{by lemma \ref{unirangelemma} (ii),} \\
                         &= \bar{\hat{f}}     
                                               && \mbox{by RR.1,} \\
                         &= \overline{f \circ g}    
                                               && \mbox{by (\ref{gCondition}),}\\
                         &= \overline{\hat{f} \circ g}     
                                               && \mbox{by RR.4.} \\
\end{align*}
\end{proof}

\subsubsection{Notes along the way}
\begin{worktt}
We might end up proceeding via
\begin{itemize}

\item showing that morphisms $j$ embedded from \catcw into the stickleback category  have the HomP suuport preservation propery
i.e have the property that for all morphisms $k$ 
\[
\overline{k \circ j} = \bar{k} \mbox{ iff } k \circ \overline{j} = k
\]
\item
 we will need to show if F and G are zigzags and G is of length 0 then 
the  condition holds,
\item we have already shown  that the supports of  morphisms satisfying this condition
are preserved by the HomP functor,
\item if one or more functors preserve supports of a morphism f then the various constructed functors also preserve the support of f.
\item Thereby showing that the constructed functors preserve supports of morphisms
satisfying the condition.
 \end{itemize} 
 \end{worktt}


\subsection{Quotients of \Par valued  Functors}
\subsubsection{Quotients and Supports}
\workt{Need review and improve}
Suppose \catcw is a restriction category and that $F: \catc \morph \Par$ is a restriction functor into the restriction category of sets and partial functions. Suppose that for every object $x$ of \catcw there is an equivalence relationship $\sim_x$ defined on the set $F(x)$. Suppose that this equivalence relationship has the following property(s):
For every morphism $j:x \morph y$ in \catcw, for all elements $k1,k2 \in F(x)$,
\begin{enumerate}
\item $F(j)(k_1)$ is defined iff $F(j)(k_2)$ is defined, 
\item $k_1 \sim_x k_2 \mbox{ and } F(j)(k_1) \mbox{ is defined } \implies F(j)(k_1) \sim_y F(j)(k_2).$
\end{enumerate}
Define functor $\Fquotient: \catcw \morph \Par$ by defining
\begin{align*}
&x             &&\mapsto F(x)/\sim_x 
                & \mbox{i.e. the set of equivalence classes of $F(x)$ wrt equivalence relation $\sim$}\\
&j: x \morph y &&\mapsto F(j)/\sim
\end{align*} 
where $F(j)/\sim$ is defined by
\begin{align*}
[k] \mapsto [F(j)(k)] && \mbox{provided $F(j)(k)$ is defined, is undefined otherwise.}
\end{align*}
which is well-defined because of the assumption made above.

It is easy to see that $\Fquotient$ is a functor (i.e. respects composition $\circ$ and identity morphisms.)

\begin{lemma}
The functor $\Fquotient$ is a restriction functor.
\end{lemma}
\begin{proof}
We need to show that  $\Fquotient$ respects the restriction operator
i.e. we need to show that for any morphism $j:x \morph y$ in \catcw, $\Fquotient(\bar{j})= \overline{\Fquotient(j)}$.

The LHS, $\Fquotient(\bar{j})$, is defined to  be the partial function
$$ [k] \mapsto [F(\bar{j})(k)] \mbox{ providing $F(\bar{j})(k)$ is defined, is undefined otherwise,}$$
in other words, since  $F$ assumed to be a restriction functor it can be defined as mapping
$$ [k] \mapsto [\overline{F(j)}(k)] \mbox{providing $\overline{F(j)}(k)$ is defined, is undefined otherwise,}$$
and this in turn, because of the definition of restriction in \Par, means that it maps
$$ [k] \mapsto [k] \mbox{providing $F(j)(k)$ is defined, is undefined otherwise.}$$

Meanwhile the RHS, $\overline{\Fquotient(j)}$  is defined as the partial function that 
maps $[k]$  to $[k]$  provided  $\Fquotient(j)(k)$ is defined, and to be undefined otherwise.

Because,
from the definition of $\Fquotient$, $\Fquotient(j)([k])$ is defined iff $F(j)(k)$ is defined then
we have shown that LHS and RHS are identical functions, as required.
\end{proof}

\subsubsection{Quotients and Ranges }
\begin{worktt}
Review and improve and put both this and previous quotient
section together. Will need put coproducts and HomP sections
together as well.
\end{worktt}
\begin{lemma}
\label{rangefunctorquotient}
\begin{worktt}
We need to give the context here. Read what I wrote in preparation paper to see if it helps.
\end{worktt}
The functor $\Fquotient$ is a range functor.
\end{lemma}
\begin{proof} \commentary{Proof reviewed on 12 Dec 2023.}
We need to show that  $\Fquotient$ respects the range operator
i.e. we need to show that for any morphism $j:x \morph y$ in \catcw, 
 $\Fquotient(\hat{j})= \widehat{\Fquotient(j)}$.

The LHS, $\Fquotient(\hat{j})$, is defined to  be 
the partial function which maps $[k']$, where $k' \in F(y)$, as follows
$$ [k'] \mapsto [F(\hat{j})(k')] \mbox{ providing $F(\hat{j})(k')$ is defined, and is undefined otherwise.}$$
In other words, since  $F$ assumed to be a range functor it can be defined as mapping
$$ [k'] \mapsto [\widehat{F(j)}(k')] \mbox{ providing $\widehat{F(j)}(k')$ is defined and being undefined otherwise,}$$
and this in turn, because of the definition of ranges in \Par, means that it maps
$$ [k'] \mapsto [k'] \mbox{providing there exists $k \in F(x)$ such that 
\oldt{$k'=F(j)(k)$}
\newt{$k'\sim_? F(j)(k)$} and being undefined otherwise.}$$

Meanwhile the RHS, $\widehat{\Fquotient(j)}$  is defined as the partial function that 
maps $[k']$  to $[k']$  provided  there exists $k \in F(x)$ such that
$\Fquotient(j)([k])=[k']$ to be undefined otherwise.

Because,
from the definition of $\Fquotient$ we have that $\Fquotient(j)([k])= [F(j)(k)]$   
we have  that LHS and RHS are identical functions, as required.
\end{proof}

\subsection{Coproducts of \Par valued  Functors}
\tbd 

\subsection{\Rel Valued Functors}
Let \Rel be the category of sets and binary relationships.

If $F: \catc \morph \Par$ is a functor then $F$ induces 
a functor $F_R : Stk(\catc) \morph \Par$.