

For the theory of data, 
the notion of \textit{range category} that we are interested in 
are define in terms of restriction categories.

\subsection{Restriction Categories}
These were introduced by 
Cockett \& Lack (\cite{COCKETT2002}) and 
constitute an algebra that nicely represents a system of partially defined terms or maps. Applied to the theory of data, they are relevant because data fields often only optionally have values and can be considered therefore to be partial, rather than total, functional relationships. 

Cockett \& Lack define restriction categories along the following lines:
\begin{definition}
A \textit{restriction category} is a category \catcw with an additional operator on morphisms (the restriction operator $\bar{\ }$), so that from a morphism
$f: a \morph b$ in \catcw there is a 
$$\bar{f}: a \morph a$$,
satisfying

R.1 For $f:a \morph b$ in \catcw $$\bar{f} \circ f =f$$.

R.2. If \fgsourcediag in \catcw then
$$\bar{g} \circ \bar{f}=\bar{f} \circ \bar{g}.$$

R.3. If \fgsourcediag in \catcw then
$$\overline{\bar{f} \circ g} = \bar{f} \circ \bar{g}$$.

R.4. If $\sequentialdiag{a}{b}{c}{f}{g}$ in \catcw then
$$f \circ \bar{g} = \overline{f \circ g} \circ f$$.
\end{definition}

We will be needing:

\begin{lemma}
\llabel{restrictioncatlemma}
In any restriction category \catc
\begin{enumerate} [(i)]
\item For $f:a \morph b$ in \catcw,
$$\bar{\bar{f}}=\bar{f}$$
\item w.R4.\footnote{This condition described as wR.4 in Cockett et al.} 
If $\sequentialdiag{a}{b}{c}{f}{g}$ in \catcw 
 then
$$\overline{f \circ g} = \overline{f \circ \bar{g}}.$$

\item \llabel{CHECKLemma}
If \catcw is a restriction category \catcw and if
$\sequentialdiag{a}{b}{c}{f}{g}$ in \catcw then 
$f \circ \bar{g} = f$ iff $\overline{f \circ g} = \overline{f}.$
\item \llabel{restrictioncatlemma:totalk} \commentary{UNUSED}
If $\sequentialdiag{a}{b}{c}{k}{j}$ in \catcw 
and $k$ is total i.e. $\bar{k}=id_a$
and if $k \circ \bar{j}$ is total i.e. $\overline{k \circ \bar{j}}=id_a$
then
$$k \circ \bar{j} = k.$$
\end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate} [(i)]
\item Use R3 with $g$ being $id_b$.
\item 
\begin{align*}
\overline{f \circ \bar{g}}
&= \overline{\overline{f \circ g} \circ f}
    && \text{by (R4),} \\
&= \overline{f \circ g} \circ \bar{f}
    && \text{by (R3),} \\
&= \bar{f} \circ \overline{f \circ g}
    && \text{by (R2),} \\
  &= \overline{\bar{f} \circ f \circ g}
    && \text{by (R3),} \\  
&= \overline{f \circ g}
    && \text{by (R1).}
\end{align*}
\item \commentary{CHECKLemma}
LEFT TO RIGHT
Assume $f \circ \bar{g} = f$
then
\begin{align*}
\overline{f \circ g} &= \overline{f \circ \bar{g}} && \mbox{by wR.4,}\\
                     &= \overline{f} && \mbox{from the assumption.}
\end{align*}
RIGHT TO LEFT
Assume $\overline{f \circ g} = \overline{f}$
then
\begin{align*}
f \circ \bar{g} &= \overline{f \circ g} \circ f 
                                  && \mbox{R.4,} \\
                &= \bar{f} \circ f 
                                  && \mbox{from the assumption,} \\
                &= f              && \mbox{by R.1.}
\end{align*}
\item 
\begin{align*}
k \circ \bar{j} &= k \circ \bar{\bar{j}}              && \mbox{from (i),}\\
                &= \overline{k \circ \bar{j}} \circ k && \mbox{by R.4,}   \\
                &= k                                  && \mbox{from assumption $k \circ \bar{j}$ is total.}
\end{align*}

\end{enumerate}
\end{proof}


\subsection{Categories with Support}
Though we are mainly restriction categories are the structures that we are interested 
one of the constructions we use relies on categories with a such an operator 
Cockett et al \cite{COCKETT2012} describe as categories with support 
such a operator \barItself for which R.4 does not hold, though R.1, R.2 and R.3 so, and for which  the following
weaker condition holds

wR.4  If $\sequentialdiag{a}{b}{c}{f}{g}$ in \catcw then
\[
\overline{f \circ g} = \overline{f \circ \bar{g}}.
\]
Such categories Cockett et al refer to as \textit{categories with support}. 

Categories with support may satisy an addition condition which is crucial in 
what follows and which we refer to as the additional weak R.4 rule:

awR.4  If $\sequentialdiag{a}{b}{c}{f}{g}$ in \catcw then
\[
\overline{f \circ g} = \bar{f} \mbox{ \quad implies \quad} f \circ \bar{g}=f.
\]
and if this holds in a category with support ww
will say that it has awR.4 support.  
The reverse implication holds in any category with support so that in
\commentary{might this hold in ANY category with support.}
a category with \workt{awR.4} support we have, for any composable $f$ and $g$:
\[
\overline{f \circ g} = \bar{f} \mbox{\quad iff \quad} f \circ \bar{g}=f.
\]



\begin{lemma}
\llabel{supportLemma}
In any category with support \catc, 
\begin{enumerate}[(i)]
\item
if $\sequentialdiag{a}{b}{c}{f}{g}$ in \catcw 
 then
 \[
 \overline{f} \circ \overline{f \circ g} = \overline{f \circ g}
 \]
 \item \llabel{RHomFunctorSubLemma}
if  $\triplesequentialdiag{a}{x_1}{x_2}{x_3}{k}{j_1}{j_2}$ 
in a \catcw  then 
if $\overline{k \circ j_1 \circ j_2} = \bar{k}$ 
then $\overline{k \circ j_1} = \bar{k}$.
\end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate}[(i)]
\item
\begin{align*}
\overline{f} \circ \overline{f \circ g}  
           &= \overline{f} \circ \overline{f \circ \overline{g}}
                               && \mbox{by wR.4} \\
           &= \overline{\overline{f} \circ f \circ \overline{g}}
                               && \mbox{by R.3}  \\
           &= \overline{f \circ \overline{g}}
                               && \mbox{by R.1}\\
           &= \overline{f \circ g}
                               && \mbox{ by wR.4, as required}
\end{align*}
\item

Suppose $\overline{k \circ j_1 \circ j_2} = \bar{k}$  then we have
\begin{align*}
\overline{k \circ j_1} &= \overline{\bar{k} \circ k \circ j_1}
                           &&\mbox{by R.1,} \\
                       &= \overline{\overline{k \circ j_1 \circ j_2} \circ k \circ j_1}
                           &&\mbox{from the initial assumption,} \\
                       &= \overline{k \circ j_1 \circ j_2} \circ \overline{k \circ j_1}
                           &&\mbox{by R.3,} \\
                       &= \overline{k \circ j_1 \circ j_2}
                           &&\mbox{by (i),} \\
                       &= \bar{k}
                           &&\mbox{from the initial assumption, as required.} 
\end{align*}
\end{enumerate}
\end{proof}


\subsection{Partial Ordering of $Hom(A,B)$}

In a category with support, and therefore in a restriction category, we can define a partial ordering on each hom set
Hom(a,b) by defining :
$$f \leq g \mbox{ iff } f = \bar{f} \circ g.$$

The relation $\leq$, so defined, is antisymmetric because if
\begin{equation}
\label{fleqg}
f = \bar{f} \circ g
\end{equation}
and
\begin{equation}
\label{gleqf}
g = \bar{g} \circ f
\end{equation}
then
\begin{align*}
f &= \bar{f} \circ g &&\mbox {this is (\ref{fleqg}),}        \\
  &= \bar{g} \circ \bar{f} \circ g &&\mbox{by R.1 and R.2,}\\
  &= \bar{g} \circ f               &&\mbox{using (\ref{fleqg}) again,} \\
  &= g                             &&\mbox{(by (\ref{gleqf}).} 
\end{align*}

When $f$ and $g$ are partial functions and
considered as morphisms in the category $\Par$ of sets and partial functions,  
then  $f \leq g$ iff  whenever $f$ is defined at an $x$,  then $g$ too is defined at x and $f(x) = g(x)$.

\begin{lemma}
\llabel{orderingsCompose}
In a restriction category \catcw, \commentary{category with support ??}
if 
$
\paralleldiag{a}{b}{f}{f'}
$
and
$
\paralleldiag{b}{c}{g}{g'}
$
in \catcw and if $f' \leq f$ and $g' \leq g$ 
then $f' \circ g' \leq f \circ g$.
\end{lemma}

\begin{lemma}
\llabel{pointwiseorderingsCompose}
In a restriction category \catcw, \commentary{category with support ??}
if 
$
\paralleldiag{a}{b}{f}{f'}
$
\nudgedown{0.6cm}
and
$
\paralleldiag{b}{c}{g}{g'}
$
and $h: a \morph c$ in \catcw 
and if $f' \leq f$ and $g' \leq g$ 
and if $h \leq f \circ g'$ and $h \leq f' \circ g$ 
then $h \leq f' \circ g'$.
\end{lemma}
\begin{proof}
Proof is handwritten on tablet.\highlight{CHECK.}
\end{proof}

\begin{lemma}
\llabel{backPropagationSubLemma}
\begin{enumerate}[(i)]
\item \item
if $\sequentialdiag{a}{b}{c}{f}{g}$ in a category with support
then $\overline{f \circ g} \leq \overline{f}$.
\item If \fgsourcediag in a category with support
if $\bar{f} \circ g = g$ 
then $\bar{g} \leq \bar{f}$.
\end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate}[(i)]
\item This is lemma \lref{supportLemma} rewritten to use $\leq$.
\item We have to show $\bar{g} \leq \bar{f}$ i.e. that $\bar{\bar{g}} \circ \bar{f} = \bar{g}$
i.e. that $\bar{g} \circ \bar{f} = \bar{g}$. This follows because
\begin{align*}
\bar{g} \circ \bar{f} &= \bar{f} \circ \bar{g} &&\mbox{by R.2,} \\
                      &= \overline{\bar{f} \circ g} && \mbox{by R.3,} \\
                      &= \bar{g} &&\mbox{from the initial assumption.}
\end{align*}
\end{enumerate}
\end{proof}
The following lemma is shaped for a very specific later use:
\begin{lemma}
\llabel{backPropagationTerminationLemma}
Suppose we have the following morphisms in a restriction category
\[
\begin{array} {c c c c c}
       &        &        &         &\kern-0.3cm\bOb{z}  \\[0.4cm]
\bOb{x}&        &\bOb{w}                     \\[0.4cm]
       & \bOb{y} 
\end{array}
\begin{arrows}
\ncarr{y}{x} \alabel{g}
\ncarr{y}{w} \blabel{h}
\ncarr{w}{z} \blabel{i}
\end{arrows}
\]
then if 
\begin{equation}
\label{ghTight}
\bar{g}=\bar{h}
 \end{equation}
 and 
 \begin{equation}
 \label{gPropagationNull}
 \overline{h \circ i} \circ g = g
 \end{equation}
 then
$\overline{h \circ i} = \overline{h}$. 
\end{lemma}
\begin{proof}
It suffices to show that both $\overline{h \circ i} \leq \bar{h}$
and $\bar{h} \leq \overline{h \circ i}$.

The first of these 
follows directly from lemma \ref{backPropagationSubLemma} (i).

The second follows from (\ref{gPropagationNull}), 
which by lemma \ref{backPropagationSubLemma} (ii)
gives us
$\bar{g} \leq \overline{h \circ i}$, and because $\bar{g} = \bar{h}$.
\end{proof}

