\documentclass[xcolor=pst,dvips]{beamer}   %, handout

\usepackage{mathptmx}
\usepackage{amsfonts}
\usepackage{wasysym}
\usepackage{url}
\usepackage{hyperref}

\newcommand{\sharedmacros}{../../SharedMacros}
%\input{\sharedmacros/general.macros}    %causes problems when used with bamer
\input{\sharedmacros/ccategories.macros}
\input{\sharedmacros/gats.macros}
\input{\sharedmacros/metagat.macros}
\input{\sharedmacros/ermacros}
\input{\sharedmacros/erdiagram}    %beamer aware version

\renewcommand{\erpictureFolder}[0]{../../SharedPictures}

\newcommand{\fundep}[3]{#2 \xrightarrow{#1} #3}  % copied from general.macros because 
                                                 % that causes numerous problems in the context of beamer

\usetheme{Szeged}
\usecolortheme{dolphin}

%\setbeamertemplate{navigation symbols}{}

\setcounter{equation}{0}

\title[John Cartmell]{Mathematical Theory of Data \\ -- Goodness Criteria}
%% Which is to say types as they are used in practice in software development and as represented in theory in categories and in syntactic type theories.
%% There is also a subplot concerning representation of context which certain types depend on -- again represented in practice and in theory. 
\author{John Cartmell}
\institute{Ad Otium}
\date{Jan 25, 2019}
\bibliographystyle{plainnat}
\usepackage{bibentry}
\nobibliography*

\begin{document}


\begin{frame}{xxx}
\titlepage
\end{frame}

\iffalse
\begin{frame}{Introduction}
\begin{itemize}
\item
What types of things are there and how are they related? 
\begin{itemize}
\pause \item Data specifications provide the answer to this question in the context of a software development. 
\pause \item Types theories provide the answer in the context of mathematics. 
\pause \item Category theory abstracts across both these domains.
\end{itemize}
\end{itemize}
\end{frame}
\fi

\begin{frame}{Data Specification}
A \textit{data specification} is a sketch $S$ for some kind of structured category \catc.
i.e. it is a directed graph plus some path equivalences (diagrams in other words) plus 
some additional structure.  

An \textit{instance} of a data specification is a mapping $D$ from the directed graph
to the category of sets and functions which preserves structure so as
to induce a structure preserving functor $D: \catc \morph \Fin$.
\end{frame}

\begin{frame}{The Requirement}
A \textit{requirement} for a data specification $S$ 
is a set of instances of the sketch $S$ or, equivalently, is a set $R_C$ of functors where for each
$D \in R_C$, $D: \catc \morph \Fin$, where \catcw is the category generated by the sketch $S$.
\end{frame}


\begin{frame}{Goodness Criteria of Type 1}
\begin{itemize}
\item Goodness Criteria of Type 1 : No redundancy. The sketch $S$ ought to be a minimum sketch for structured
category \catcw i.e. there should be no subsketch of $S$ which generates  \catc.
\pause{
\item Rationale: Follow software engineering principles:
\begin{itemize}
\item Ensure there is no redundancy in the  data that is stored and/or communicated.
\item Avoid duplicate program logic.
\end{itemize} 
}
\end{itemize}
\end{frame}

\begin{frame}{Goodness Criteria of Type 2}
The structured category  ought to be the tightest possible fit to the requirement $R_{\catc}$.
I will use the  \textit{maximally constrained}.
\begin{definition}
If $\catc$ is a structured category and $\reqtc$ is a requirement for $\catc$,  
if $I: \catc \morph \catcp$ is a structure preserving functor then 
say that $I$ is \textit{consistent with} requirement $\reqtc$ iff for all instances $D \in \reqtc$ 
there exists a structure preserving functor $D':\catcp \morph \Fin$ such that $I \circ D'=D$.
\end{definition}
\begin{definition}
Define \catcw to be \textit{maximally constrained} to the requirement $\reqtc$ iff 
for all structured categories $\catcp$ 
and for all structure preserving functors $I:\catc \morph \catcp$ that are consistent with $\reqtc$, 
for all functors $F: \catc \morph \Fin$  there exists an $F' : \catcp \morph \Fin$ such that $I \circ F'=F$.
\end{definition}
\end{frame}


\begin{frame}{Test}
Blob TWO
\end{frame}

\begin{frame}{Rationale}
\begin{itemize}
\item Good data modelling is good theorising and good theorising is having the tightest theory that fits the facts.
\item Software Engineering: 
\begin{itemize}
\item Document the requirement.
\item Validate data against the requirement. 
\item Restrict degrees of freedom in the data representation.
\item Use declarative style if possible.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Representation of Path Equivalences}
\begin{lemma}[Path Equivalence Representation Lemma]
If $\catc$ is a locally finite category and $\reqtc$ is a set of instances, if $\catc$ is 
\textit{maximally constrained} to the requirement $\reqtc$ then all path equivalences with respect
to $R_C$ are represented in \catcw
i.e. for all diagrams
$
\rule[-0.3cm]{0pt}{0.9cm} %to add vertical space of diagram -- based on lowering diagram 0.3cm and heght 0.9cm
                            % change thickness from 0pt to 1 pt to debug
\begin{array}{c p{0.5cm} c  }
 \Rnode{a}{a}            &&   \Rnode{b}{b}
\end{array} 
$
\ncarc[nodesep=2pt,arcangle=10,offset=2pt]{->}{a}{b}
\alabel{f}
\ncarc[nodesep=2pt,arcangle=-10,offset=-2pt]{->}{a}{b}
\blabel{g}
in $\catc$,  if in all instances $D \in \reqtc$, $D(f)=D(g)$, 
then $f=g$ in $\catc$.
\end{lemma}
\end{frame}


\begin{frame}
\begin{lemma}[Functional Dependency Representation Lemma]
If $\catc$ is a locally finite category and $\reqtc$ is a set of instances, if $\catc$ is 
\textit{maximally constrained} to the requirement $\reqtc$ then
all functional dependencies $\fundep{H}{f}{g}$  with respect to $\reqtc$ are represented in $\catc$.
\end{lemma}
\end{frame}

\begin{frame}
\begin{lemma}[Referential Inclusion Dependency Representation Lemma]
If $\catc$ is a locally finite category and $\reqtc$ is a set of instances, if $\catc$ is 
\textit{maximally constrained} to the requirement $\reqtc$ then
every referential inclusion dependency with respect to $\reqtc$ is represented in $\catc$.
\end{lemma}
\end{frame}

\iffalse
\begin{frame}{Bibliography}
\bibliography{../../SharedBibliography/temp/bibliography}
\end{frame}
\fi

%\fi

\end{document}