
\section{Entity Types}

\mynote The word `entity' as it is used in this book is used in its most general sense,  the sense in which it just means `thing'. The entities that are being modelled in entity modelling can be just any old 'things' at all --- and so we could just as well speak of `modelling things' as `entity modelling' but that the term entity modelling
has come to mean a particular way of modelling things in which things are described by describing their types and the types of relationships between them.

\mynote There is a proviso to this. We do not set out to model types whose instances are all known from the beginning (though these types are available and relevant). Rather we set out to model types whose instances (things) are open-ended and may vary from context to context. These are the subject entity types within an entity model. The whole entity model is expected to be realised in many contexts and the instances of any given subject type are expected to vary from context to context.
 
\mynote
Conversely, things that are known from the beginning and unchanging from one context to another, things that can be said to be `universal' are excluded from being subjects of our modelling. We do not model, therefore, whole nunbers, nor real nunbers, nor truth values (booleans) but take these as universal givens and, generally, we do not model language characters and language character sequences (strings) with all the richness that this would require and we take these too to be universal and given. 

\mynote
In philosophy the non-universal things that are our entitites are referred to as particulars and so we can say that subject entity types within an entity model represent types of things all of whose instances are particulars. 

\mynote Though entity modelling may be about just about any thing it is never about modelling everything.
An entity model serves to describe a domain of discourse. In the general case it describes the entities in this domain in relation not just to each other but to other given types which are types which serve to represent that which is universal. 

\section{Relationships}
\mynote Dictionary definitions of the term relationship include the following
\begin{itemize}
\item --- A relationship between two things is the way in which they are connected.
\item --- A relationship is a particular instance of being related.
\item --- In logic and mathematics relationship is another name for relation.
\item --- A relationship is a connection, association, or involvement between two things.
\end{itemize}

\mynote In two cases above the dictionary is describing relationships between two kinds of things and these are the relationships we are concerned with here and which, being between two things, are called binary relationships. How do we explain what these are? It is a lot clearer if we proceed by example.


\mynote
Well, friendship, acrimony, marriage, adjacency and succession can all be given examples of binary relationships. There are binary relationships too  between parent and child, between teacher and student and between a planet and the star that it orbits.   
The following are also binary relationships:
being employed by, in the sense that a person might be employed by a particular employee,
managing, in the sense that an employee might manage a project. There are binary relationships too between a sales order and the customer that it is ordered by and between a debit card and  the bank that issued it. A banking transaction has binary relationships both with the account debited and with the account credited. 

\mynote The most important feature of relationships as the term is used in entity modelling in this book is that the existence of a relationship instance between and two particular entities should be a matter of black and white.\footnote{we easily drawn in to speaking of binary relationships having instances and it is easy to talk about these instances as relationships which, though an overloading of the term relationship, doesn't present practical difficulties.
I think this overloading of the term relationship is already observable in the dictionary bullet points I gave above. Is this the same overloading as when we say that a pig is a type of animal but then speak of an instance of the type as a pig?} For example between any particular Jack and any particular Jill we must take it that there either is or is not a friendship between them. There can be no quality or relativisation of the friendship --- either it is or it isn't. If some binary association between things
(like friendship)  requires qualification in some way then it isn't appropriate to model it as a binary relationship in an entity model. This is a point of difference, by the way, between our terminology and that of Chen --- our relationships \textit{are} mathematical relations. 

\section{Attributes}
\mynote Though we do not set out to model types of universals and the relationships between them 
in an entity model, we do model relationships between subject entity types and types of universals. 

\mynote 
There are two kinds of relationship 
in an entity model those between entity types and those between 
entity types and  given types of universals. 
A relationship of the second kind is said to be an attribute. Such an attribute is a property of the subject entity type. 

\mynote Since relationships of the second kind are called attributes, the term relationship is reserved, in entity modelling terminology and subsequenly in these pages, for relationships of the first kind.  The principle components of an entity model are its entity types and their attributes and its relationships.\footnote{Some authors would say entity types, relationship types and attributes and though I have sympathy with this terminology I shall stick to relationship where others perhaps correctly have said relationship type}

\mynote
Note that the presence of attributes and therefore the involvement of given types of universals
is an optional feature of entity modelling. 

\section{On the naming of relationships}

\mynote Sometimes relationships are easy to name, sometimes not. 

\begin{center}
\begin{tabular} {| l | p{9cm} |}
\hline
parent--child & also described as the parenthood relationship\\
succedes--succeeded-by &  the succession relationship \\
married to             &  the matrimony relationship \\
owns--owned-by         & the ownership relationship\\
references--referenced-by & the referentiality relationship?\\
composed-of--part-of   &  the composition relationship\\
                       &  the aquaintance relationship\\
\hline
\end{tabular} 
\end{center}

\mynote We might model that a bicycle is related to its front wheel and that a bicycle is related to its back wheel but we will not then need to model,  and therefore ought not to model, that the front wheel of a bicycle is related to its back wheel because though there is such a relationship it is implied by the other two.  

\mynote 
We don't have a simple name for the relationship between a planet and the star it orbits. Instead we have  relative terms `orbits' and `orbited by' which describe each of the parties in the relationship from the point of view of the other. This dual naming of binary relationships is the way that relationships are named in an entity model. 
This is where a diagram comes in handy.

\section{Diagrams}
  
\mynote Mathematically speaking an entity model is not much more than a labelled directed graph and for this reason they are most usually presented as diagrams.

\mynote When entity models are presented in diagrams as directed graphs
they are called ER diagrams (E for entity, R for relationship) or sometimes
ERA diagrams (where the A is for attribute). 