
\begin{oldtt}
\commentary{can this go?}
\begin{lemma}
\llabel{forwardPropagationSubLemma}
In a range category
\begin{enumerate}
\item if $\sequentialdiag{a}{b}{c}{f}{g}$ 
then $\reallywidehat{f \circ g} \leq \hat{g}$,
\item if \fgsinkdiag and if $f\circ \hat{g} = f$ then
  $\hat{f} \leq \hat{g}$.
\end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate}
\item
We need show that 
$\overline{\reallywidehat{f \circ g}} \circ \hat{g} = \reallywidehat{f \circ g}$.
This follows because

\begin{align}
\overline{\reallywidehat{f \circ g}} \circ \hat{g} 
          &= \reallywidehat{f \circ g} \circ \hat{g}  &&\mbox{by RR.1,}//
          &= \reallywidehat{f \circ g} 
                &&\mbox{by lemma \ref{fglemma} (i) and (ii).}
\end{align}
\item
To prove $\hat{f} \leq \hat{g}$ we have to prove 
$\bar{\hat{f}} \circ \hat{g} = \hat{f}$ 
i.e. that $\hat{f} \circ \hat{g} = \hat{f}$. 
This follows because   
\begin{align*}
\hat{f} \circ \hat{g} 
     &= \reallywidehat{f \circ \hat{g}}
                  &&\mbox{by lemma \ref{fglemma} (ii),} \\
     &= \hat{f}   && \mbox{from the initial assumption.}
\end{align*}
\end{enumerate} 
\end{proof}
\end{oldtt}
\begin{oldtt}
\commentary{Is this used?}
The following lemma complements lemma \ref{backPropagationTerminationLemma} 
and the two will be used in tandem later.
\begin{lemma}
\llabel{forwardPropagationTerminationLemma}
Suppose we have the following morphisms in a range category
\commentary{I am thinking this shape might be \workt{wrong}.}
\commentary{Consider more carefully what reqd. in later proof.}
\[
\begin{array} {c c c c c}
       &\bOb{x}                            \\[0.4cm]
\bOb{w}        &        &\bOb{y}           \\[0.4cm]
       &       &        &         &\kern-0.3cm\bOb{z} 
\end{array}
\begin{arrows}
\ncarr{w}{x} \alabel{f}
\ncarr{y}{x} \blabel{g}
\ncarr{z}{y} \blabel{h}
\end{arrows}
\]
then if 
\begin{equation}
\label{forwardPropagationTight}
\widehat{f}=\widehat{g}
\end{equation}
and 
\begin{equation}
\label{forwardPropagationTwo}
f \circ \reallywidehat{h \circ g}=f
\end{equation}
then 
\begin{equation}
\label{forwardPropagationThree}
\reallywidehat{h \circ g} = \reallywidehat{g}.
\end{equation}
\end{lemma}
\begin{proof}
The result follows by showing that
$\reallywidehat{h \circ g} \leq \reallywidehat{g}$
and
$\reallywidehat{g} \leq \reallywidehat{h \circ g}$.

The first of these 
follows directly from lemma \ref{forwardPropagationSubLemma} (i).

The second follows from (\ref{forwardPropagationTwo}), 
which by lemma \ref{forwardPropagationSubLemma} (ii)
gives us
$\hat{f} \leq \reallywidehat{h \circ g}$,
and because $\hat{f} = \hat{g}$.
\end{proof}
\end{oldtt}

