 \documentclass[10pt,a4paper]{article}
\newcommand{\sharedmacros}{../../../SharedMacros}
\include{\sharedmacros/ccategories.macros}
\include{\sharedmacros/ermacros}
\include{\sharedmacros/erdiagram}
\include{\sharedmacros/theorems.macros}
\include{\sharedmacros/general.macros.v2}
\include{\sharedmacros/diagram.macros}
\include{\sharedmacros/rangeCategories.macros.tex}
\usepackage{xparse}
\usepackage{xstring} % for some diagram macros
\usepackage{fancyvrb} % used for documenting examples of macros 
                      % along with verbatim text of example.
\usepackage{anyfontsize}
\usepackage{mathtools}  % why have we got this?
\usepackage{mdframed}
\usepackage{alltt}    
\usepackage{cmll}
\usepackage{ulem}
\renewcommand{\ttdefault}{txtt}
\usepackage[left=1.5cm, right=4cm, marginparwidth=3cm, top=2cm, bottom=2.0cm]{geometry}
\usepackage{framed}  %used for box framing for example in asides
\usepackage[font=small]{caption}
\setlength{\captionmargin}{2cm}
\theoremstyle{remark}
\newtheorem*{lemma*}{Lemma}
\renewcommand{\term}[1]{\textit{#1}}  %SIMPLE UNINDEXED VERSION
\usepackage{imakeidx}
\makeindex[name=definitions, title=Index of Definitions]
\makeindex[name=lemmas, title=Index of Lemmas]

\usepackage{makeidx}
\makeindex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\srcFolder}{src}
%\includeonly{\srcFolder/MToD.rangeCategories.introduction.tex}
%\includeonly{\srcFolder/MToD.rangeCategories.restrictions_and_support.tex}
%\includeonly{\srcFolder/MToD.rangeCategories.ranges_sections_and_weak_range.tex} 
\includeonly{\srcFolder/MToD.rangeCategories.sticklebacks.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\begin{document}
\title{Range Categories and the Theory of Data\\[0.5em]
\large Foundations of the Mathematical Theory of Data}

% abstract here for ams
\author{John Cartmell}
\date{}
\maketitle
\begin{center}
Work in progress January 2026 Going on February, March, April.
Much text regarding failed hom functor idea hived off --- see
parallel whenEverythiingFellApart.
\end{center}
\begin{aside}
Much text regarding failed hom functor idea hived off --- see
parallel whenEverythingFellApart.
\end{aside}
\newcommand{\Fquotient}{F/\!\!\sim}  % \! is small negative space

\newcommand{\IfSforCwithRCwords}{If $S$ is a sketch for category \catcw and if $S$ is considered as a data specification with requirement $\reqtc$\ }

\newcommand{\fgsourcediag}{$\binarysourcediag{a}{b}{c}{f}{g}$}
\newcommand{\fgsinkdiag}{$\binarysinkdiag{a}{b}{c}{f}{g}$}
\newcommand{\FGsourcediag}{$\binarysourcediag{a}{b}{c}{F}{G}$}
\newcommand{\FGsinkdiag}{$\binarysinkdiag{a}{b}{c}{F}{G}$}


\newcommand{\cnc}{\!::\!}
\newcommand{\tighten}[1]{\mathcal{T}\bigl(#1\bigr)}
%\newcommand{\shorten}[3]{\mathcal{S}_{#3}(#1,#2)}
\newcommand{\shorten}[3]{\mathcal{S}^{#2}_{#3}\bigl(#1\bigr)}
\newcommand{\Tighten}[1]{\mathcal{T}\Bigl(#1\Bigr)}
%\newcommand{\shorten}[3]{\mathcal{S}_{#3}(#1,#2)}
\newcommand{\Shorten}[3]{\mathcal{S}^{#2}_{#3}\Bigl(#1\Bigr)}

\tableofcontents


\begin{aside}
I must user the term zigzag throughout and say that in an RR.5 range category every zigzag has a normal form. These normal forms are the 
short sticklebacks described by Cockett. 

It is potentially confusing though because these normal forms are tuples of morphisms 
of \catcw and these morphisms are generated from the sketch. The morphisms themselves expressed as paths of arrows from the sketch do not in general have normal forms.

$shorten(T(F))$ can be potentially written as $normal(F)$.
\end{aside}



\include{\srcFolder/MToD.rangeCategories.introduction.tex}
\include{\srcFolder/MToD.rangeCategories.restrictions_and_support.tex}
\include{\srcFolder/MToD.rangeCategories.ranges_sections_and_weak_range.tex} 
\include{\srcFolder/MToD.rangeCategories.Par_valued_functors.tex}
\include{\srcFolder/MToD.rangeCategories.sticklebacks.tex}

% the file should be renamed the data functor or something more semaantic
\include{\srcFolder/MToD.rangeCategories.carry.tex}
\include{\srcFolder/MToD.rangeCategories.Rel_category.tex}

\section{Goodness of Data Specifications}
\subsection{General Principles}

\include{\srcFolder/MToD.rangeCategories.goodness_principles.tex}

\subsection{Specific Criteria}

\include{\srcFolder/MToD.rangeCategories.goodness_criteria.tex}

\section{Goodness Lemmas}
\begin{aside}
Since all these lemmas use the functor $d_a$ then they all have the assumption that
\catcw is an RR.5 range category.
\end{aside}

\include{\srcFolder/MToD.rangeCategories.lemma2A.path_equivalence_representation.tex}

\include{\srcFolder/MToD.rangeCategories.lemma2B.fd_representation.tex}

\include{\srcFolder/MToD.rangeCategories.lemma2C.referential_dependency_representation.tex}

\bibliographystyle{alpha} 
\bibliography{../../../SharedBibliography/temp/bibliography}

\printindex[lemmas]


\end{document}
