\documentclass[10pt,a4paper]{article}

%\usepackage{geometry}
\usepackage{pstricks}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{stmaryrd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{enumerate}
\usepackage{calc}
\usepackage{float} % for H option for tables and figures
\usepackage{changepage} % used for adjustwidth
\usepackage{framed}
\usepackage{wasysym}
\usepackage{ulem}   %added because \sout was not striking out 
                    % supports use uline and uwave also dashuline dotuline
\usepackage{epsf}
\newcommand{\SharedMacros}{../../SharedMacros}
\newcommand{\SharedText}{../..SharedText}

%\usepackage{enumitem} % to be able to [resume] an enumeration but clashes with [(i)] !!
\input{\SharedMacros/ermacros}
\input{\SharedMacros/ccategories.macros}
\input{\SharedMacros/gats.macros}
\input{\SharedMacros/metagat.macros}
\input{\SharedMacros/erdiagram}
\input{\SharedMacros/theorems.macros}
\input{\SharedMacros/general.macros.v2}
\input{\SharedMacros/alternate.nonbeamer.macros}
\input{\SharedMacros/indexedsets.macros.tex}
\input{\SharedMacros/barkerEllis.macros.tex}

\usepackage{hyperref}
\setcounter{equation}{0}

\newcommand{\ImagesFolder}{../images}
\renewcommand{\erpictureFolder}[0]{../images}
\newcommand{\handCraftedImagesFolder}{../handCraftedImages}
\newcommand{\photosFolder}{../photos}

\input{../localMacros}

\begin{document}
Travelling to the airport, I received a text message about my flight. The message read:
\begin{equation}
\label{LH2502PhoneMessage}
\text{\parbox{9cm}{\textit{
Your flight LH2502 from Munich to Manchester on 14 August 2024 at 15:55 will depart from gate L06.}}}
\end{equation}
Consider this message. It contains multiple referentials and individually or combined these make reference to
multiple entities:
\begin{itemize}
    \item the flight number, LH2502, makes reference to an \textit{airline service},\footnote{which, from one point of view at least, is a little odd because the name flight number purports to reference a flight. It isn't so odd though because on any given day flight numbers do reference flights.}
    \item each airport name, Munich, respectively, Manchester, makes reference to an \textit{airport},
    \item the combination of flight number, LH2502, and date, 14 August 2024, make reference to a \textit{flight},
    \item the combination of the airport the flight is identified as being from, Munich, and the gate number, 
    L06, identifies a \textit{gate}.
\end{itemize}


These referentials, the names (Munich, Manchester), numbers (LH2502, L06) and the date (14 August 2024)
I can understand as the values of attributes of the various referenced entitites. 
The types of these entities (airline service, airport, flight and gate), the relationships between them and the attributes employed
(flight number, date of departure, airport name and gate number) I can arrange on a diagram like this:
\begin{equation}
\label{boardingGate1}
\raisebox{-1.5cm}{\input{\ImagesFolder/boardingGate1.tex}}
\end{equation}

When we analyse in detail how the message comes to references a specific gate 
 then we see that 
the referencing only works because of what we know, \textit{a priori}, about this arrangement of relationships. In particular, we know  that for every flight $x$, 
the gate that the flight leaves from (once allocated) is at the airport that is  the departure airport for the
service operated by the flight. That is 
\begin{equation}
d2(r3(x))=r1(d1(x)), \mbox{ provided $r3(x)$ is defined.}
\end{equation}
which in variable-free notation can be written as
\begin{equation}
r3 \circ d2 \leq d1 \circ r1
\end{equation}

\begin{erboxedFigure}{H}{airlineFlightLeavingFromScope}{
In this diagram of relationships there are two paths of relationships leading from 
type \textit{flight} to type \textit{airport} --- one mediated by \textit{airline service} and one mediated by \textit{gate}. Both these paths are equivalent once a departure gate has been allocated. }
\input{\ImagesFolder/airlineFlightLeavingFromScope.tex}
\end{erboxedFigure}

There is knowledge that we have that enables us to interpret the message regarding the departure gate. 
It is best expressed as an algebraic inequality. 
Algebra like this is common but unrecognised in everyday speech, as in this text message, but also in internal messaging and data representation within computer systems. 

\end{document}
