 \documentclass[12pt,a4paper]{article}

\newcommand{\sharedmacros}{../../SharedMacros}
\input{\sharedmacros/ccategories.macros}
\input{\sharedmacros/ermacros}
\input{\sharedmacros/erdiagram}
\input{\sharedmacros/theorems.macros}
%\input{\sharedmacros/general.macros}
\usepackage{mathptmx}  % This changes font to roman
\usepackage{anyfontsize}
\usepackage{mathtools}  % why have we got this?
\usepackage{alltt}    
\usepackage{mnsymbol} %used for rightpitchfork
\usepackage{cmll}
\usepackage{ulem}
\renewcommand{\ttdefault}{txtt}
\usepackage[left=1.5cm, right=4cm, marginparwidth=3cm, top=2cm, bottom=2.0cm]{geometry}
\usepackage{framed}
\usepackage[font=small]{caption}
\setlength{\captionmargin}{2cm}
\theoremstyle{remark}
\newtheorem*{lemma*}{Lemma}

\usepackage{arydshln} % vertical dashed lines between columns of an array

\input{../macros/generalmacros.tex}
\input{../macros/othermacros.tex}
\input{../macros/beamerclashmacros.tex}
%ENDCOPY

\begin{document}


\section{Preamble}
\note
Title: Mathematical Theory of Data-- Generic Goodness Criteria for Data Specifications
\note It says in my abstract that \textit{I will explore the relationship between sketches of categories with additional structure and data specifications}. 
Actually what I really want to do is to explore the relationship between sketches of categories as data specifications and the relational model of data
because I would like to establish sketches as a generic theoretical framework for data specification.
To that end I want to propose and elaborate  generic goodness criteria for sketches of categories as data specifications and  show how  the classic relational database normal form criteria follow as special cases of the generic criteria. 
\note
Data specifications ... What are they? How are they used? What do they look like? Well they are used to describe the data that underlies systems and the data that is communicated between s/w agents. One that I recently developed look like this.
This  data specification describes data processed within and stored by a technical scientific application. 
The application had and has  an architecture which dictated use of  four different coding languages -- IDL, XML, Java and Typescipt.  
This data specification was  written in a language called ER Script -- E for entity, R for relationship -- 
and code generators were used to generate code  in each of the four languages directly from the data specification.
\note
In this specification there are many navigation paths through the data. Essentially they can be seen in this diagram here. 
Also, in many cases there are pairs of distinct navigation paths that are equivalent. In other words there are many diagrams that commute.
\note
Amazingly in this diagram there are two pullback cubes one stacked upon the other -- so 11 pullback squares in one part of this diagram.
\note 
With the experience of this data and other highly structured data and overtime as I have applied categorical concepts in my programming 
it has been natural for me to take a category-theory inspired view of data
and to wonder what relationship could be established between this categorical view and the well-established relational view of data.
\note 
There are a couple of papers that describe such a relationship for example a paper by
Michael Johnson and Robert Rosebrugh from 2001. However their paper does not go very far and 
 in my opinion  stops short of some quite low hanging fruit.
\note 
Now part of the appeal of the relational model early on on its lifetime was that there was   a body of theory developed about the model. 
... but this body of theory is specific to the relational model i.e. it is not generic it doesn't apply to specifications of data that is stored and/or communicated non-relationally ... as for example the data described here ... in this diagram.
\note
I believe that a MToD can be developed which is generic and that such a theory can support software technology and engineering more effectively than current theory does. To persue this I have formulated generic goodness criteria that apply to data specifications in general rather than to specific models or technologies and I have tried to show that the classic relational database normal form criteria are implied by these generic criteria. 
I \textbf{have} made progress in this....and  this is what I am going  to go on and describe to you. Gratitude.

\end{document}

