

\note \commentary{added 24 April 2021} I am going to say that two theories are fungible\footnote{I am using this word (which I learnt from Ian Stewart) so as to avoid use of the word `equivalent' at this point.} when one is just as good as the other for a purpose at hand.
\note 
Are the theories of meta-gat algebras, contextual categories and contextual categories with families fungible? Is one of these better than the the others? I have in the past been of the opinion that all three are fungible and but also that there is a primacy of the  first over the second over the third and this because the second and third fall foul of occam's razor and multiply entitites needlessly.  Recently I have become unsure that these three theories are fungible -- it depends, of course, on what the purpose at hand is.
\note
To investigate what we might mean by fungible conside generalised algebraic theories $U_1$ :
\begin{equation*}
\label{tmorphism}
\begin{array}{c}
\isT{A} \\
\isT{B} \\
\ofT{y}{B} \tstyle \ofT{f(y)}{A}
\end{array}
\end{equation*}
and $U_2$ :
\begin{equation*}
\label{tfamily}
\begin{array}{c}
\isT{A} \\
\ofT{x}{A} \tstyle \isT{B(x)} \\
\end{array}
\end{equation*}

Now to every set $I$ there is an isomorphism between morphism with codomain some set $I$ and families of $I$-indexed sets
and so the categories $U_1$-alg and $U_2$-alg of instances of these theories in the contextual category $\Fam$ are equivalent 
 With respect to a contextual category \catcw, in general, the category of internal $U_1$ objects and the category of internal  $U_2$ objects will not necesarily be equivalent however if category 
\catcw has a $\Sigma$-structure and has strong identity types, as has $\Fam$, then these two categories are equivalent.  

Likewise the two theories are not isomorphic in the category of generalised algebraic theories (nor are they equivalent 
in the 2-category of generalised algebraic theories)  however in generalised algebraic theories enriched with $\Sigma$
and strong identity types these theories are isomorphic/equivalent\commentary{delete as approriate} in the category
of such enriched theories.

\note I have been guilty of double think in the past. I have always instinctively felt that $U_1$ and
$U_2$ were not fungible but that theories of contextual categories and meta-gat algebras were fungible.
Both cannot be true.

\note Another example concerns categories with finite products and cartesian multi-categories. These are fungible if the purpose at hand is to look at such structures in (internal to) $\Fam$ but
not so if the purpose is to look at such structures internal to  more general contextual categories. 

\note The conclusion of this afterthought is that it would be improper for foundations to start with contextual categories
(or C-systems, or contextual categories with families). If a foundation is to start with any such thing it should start with 
meta-gat algebras. These are the same things as B-systems except, that is,  for the concern  for foundations and the resultant sensitivity to what we mean by `same as'.   


