
\usepackage{mathptmx}
\usepackage{amsfonts}
\usepackage{wasysym}
\usepackage{url}
\usepackage{hyperref}

\newcommand{\sharedmacros}{../../../SharedMacros}
%\input{\sharedmacros/general.macros}    %causes problems when used with bamer
\input{\sharedmacros/ccategories.macros}
\input{\sharedmacros/gats.macros}
\input{\sharedmacros/metagat.macros}
\input{\sharedmacros/ermacros}
\input{\sharedmacros/erdiagram}    %beamer aware version
\input{\sharedmacros/general.macros.v2} %try this instead
\input{\sharedmacros/diagram.macros}
\input{../macros/beamermacros}
%\input{../macros/generalmacros} %get rid of this
\input{../../macros/othermacros}
\input{../macros/sketchmacros}
\input{../macros/directedgraphsmacros}

\renewcommand{\erpictureFolder}[0]{../../SharedPictures}
\setcounter{equation}{0}


\title[John Cartmell]{Elements of a Mathematical Theory of Data}
%% Which is to say types as they are used in practice in software development and as represented in theory in categories and in syntactic type theories.
%% There is also a subplot concerning representation of context which certain types depend on -- again represented in practice and in theory. 

\author{John Cartmell}
\institute{\\}
\date{Dec 12, 2023}
\bibliographystyle{plainnat}
\usepackage{framed}
\usepackage{bibentry}
\usepackage{colortbl}
\usepackage{ulem}   % for \dashuline{dashing} for seconday key
\usepackage{listings}
\lstset{%
  escapeinside={(*}{*)},%
}
\usepackage{arydshln} % vertical dashed lines between columns of an array
\usepackage{pst-arrow} %for \bigarrow


%Redefine the \Fin macro to be category of sets
%\renewcommand{\Fin}{\Set}


\newcommand{\slidecontext}{Introduction} % to be redefined before use

\newcommand{\comingnext}[1]{
\begin{frame}{COMING NEXT}
\begin{center}
\Large #1
\end{center}
\end{frame}
}


\begin{document}
\begin{frame}
\titlepage
\nobibliography{../../SharedBibliography/temp/bibliography}
\end{frame}

\begin{frame}{Background}
\begin{itemize}
\item Concept Instance Algebras
\item Preparation for a Mathematical Theory Of Data
\item Instances of Generalised Algebraic Theories
\end{itemize}
\vspace{3cm}
\url{www.researchgate.net/profile/John-Cartmell}
\end{frame}


\begin{frame}{Terminology}
\begin{itemize}
\item \textit{modelling} for me is \textit{theorizing}
\item \textit{instances of theories}  rather then \textit{models of theories}
\item \textit{data specifications} except when I forget and I call them \textit{data models}
\item the act of constructing data specifications is \textit{data modelling}
\item a \textit{model of data} is a meta-theory (a meta-model) describing what constitutes a data specification 
\item most significantly there are \textit{relational} and \textit{nested relational} models of data
\item the mathematical theory of data is a new proposed meta-theory of data with a claim to universiality.
\end{itemize}
\end{frame}


\section{Mathematical Theory of Data LITE}

\begin{frame}{The Thesis}
\begin{itemize}[<+->]
  \item A key activity in s/w systems development and maintenance is the creation and maintenance of specifications of the data stored in databases and variously intra-communicated between components of a system and inter-communicated between systems in messages. 
  \item There are gross inefficiencies in methodologies and working practices used in carrying out this key activity and in the data specifications themselves. These are reflected in undue complexity in program code to access and manage the data.
  \item These inefficiences have been established and endorsed by theory and by those experts who promote this theory and associated methodologies.
  \item A new theory is required to both expose and remedy the shortcomings.
  \item This should be the role of a new \textit{Mathematical Theory of Data}.
  \item The dual challenge is to develop the theory and to positively impact best practice.
  \item Here, today, I am talking about some elements of this meta-theory.
\end{itemize}
\end{frame}

\begin{frame}{Mathematical Theory of Data}
\begin{itemize}
\item is a meta-theory,
\item covers \textit{principles} and \textit{criteria} for goodness of data specifications,
\item reveals the significance of commutative diagrams and therefore category theory,
\item slogan \textit{--- good data modelling is good theorising}. 
\end{itemize}
\end{frame}

\input{../slides/Shlaer&Lang/relationalDataSlide}  % this slide defines inclusion dependencies
\input{../slides/Shlaer&Lang/relationalDataSchematicSlide}
\input{../slides/Shlaer&Lang/sketchSlide}
\input{../slides/Shlaer&Lang/nestedRelationalDataSlide}
\input{../slides/Shlaer&Lang/nestedRelationalDataSchematicSlide}
\input{../slides/Shlaer&Lang/contextualSketchSlide}

\iffalse
\subsection{Example}
\begin{frame}{Example -- LCMSMS Data}
\scalebox{0.2}{
\input{\erpictureFolder/chromatogramanalysisrecord}
}
\pause
\vspace{-5cm}
\begin{block}{}
This diagram has representations of
\begin{itemize}
  \item a terminal object,
  \item 35 structurally implemented relationships, 
  \item 22 reference relationships implemented by foreign key,
  \item 5 coproducts,
  \item approx 15 commutative diagrams,
  \item xxx pullback diagrams.
\end{itemize}
Generated into code in XML, ECMA Javascript, Python.
\end{block}
\end{frame}
\fi

\begin{frame}{Nested data structures}
draw single nested relational data model
less columns now and that is because some structure is represented niot by foreign keys but by the nesting of struture. Thats why this example is representative of how data about departments, professors and students might be represented in other messaging systems such as XML ad IDL and also in hierarchical databases. 

decisions about how which data relationships to represent by nesting are recorded by distinguishing certain morphisms of the sketch as being represented `structurally'
(in IDL for eaxample the keyword would be struct).

These distinguished morphisms determine (have you guessed this yet) a contextual category 
\end{frame}

\begin{frame}
same again diagram with dependency morphisms replacing d1 and d2
\end{frame}

\begin{frame}{Missing Data and Range Categories}
\end{frame}

\begin{frame}{The problem caused by absence.}
\begin{itemize}
\item relational schemas (SQL) can be generated from this diagram. So to data structure definitions in programming language and in messaging technologies such as XML and IDL. 
\item the commutivity of the diagrams, $s \circ d2 = d_1$ and $h circ d2 =id$, is reflected in the relational structure and the other representations such as in XML and IDL.  
\item that the triangles commutes has an effect on the relational structure but equally on xml or idl or other such message structures. 
\end{itemize}
\end{frame}

\begin{frame}{Generalities at this point}

go off into generalities at this point. data specificatioms as theories of what is
\end{frame}


\iffalse
\subsection{Preamble}
\input{../slides/normalformliterature.slides}

\subsection{Introduction}
\input{../slides/introductory.slides}
\fi

\iffalse
\subsection{Example Data}
\input{../slides/data.slides}

\subsection{Directed Graphs}
\input{../slides/directedgraph.slides}

\subsection{Categories}
\input{../slides/goodnesscriteria.category.slides}


\subsection{Categories with Designated Monomorphisms and Epimorphims}
\input{../slides/goodnesscriteria.epimono.slides}

\subsection{Categories with Finite Products and Designated Monomorphisms and Epimorphims}
\input{../slides/goodnesscriteria.products.slides}

\begin{frame}{Summary}
\begin{itemize}
\item If a sketch of a category with designated monos and epis, finite products and epi-mono splits is
 considered as a data specification with a given requirement then there are  two generic criteria which when met
 ensure that it meets an abstract BCNF criteria.
\item Similar results can be obtained for 2NF and where appropraite 3NF criteria as well as for 4NF and 5NF but in this case a category with limits is required.
\end{itemize}
\end{frame}
\fi



\iffalse
\subsection{Entity-Relationship Notation}
\input{../slides/ERA.slides}
\fi

\iffalse
\subsection{Unused}
\input{../slides/unused.slides}
\fi

\end{document}