
\begin{frame}{E.F. Codd 1970}
\begin{itemize}
	\item In 1970 E.F.Codd introduces the relational model of data and introduces the idea of normal form.
\end{itemize}
	\displaybibentry{Codd1970}
\end{frame}

\begin{frame}{E.F. Codd 1971}
In 1971, the terms `functional dependency' and  `third normal form' are introduced 
in an IBM techical report published in a now out of print book (also unavailable on Amazon).

\displaybibentry{Codd1971} 
\end{frame}

\begin{frame}{Fagin 1977,1979}
\begin{itemize}
	\item In 1977 Fagin introduces `fourth normal form' (4NF) and `multivalued dependencies'.
	\item In 1979, Fagin describes `projection-join normal form' also known as fifth normal form (5NF).
\end{itemize}
\displaybibentry{Fagin1977} 

\displaybibentry{Fagin1979} 
\end{frame}

\begin{frame}{Normal Forms}
Ling and Goh 1992 
\begin{quote}
Since
classical normal forms (including the Improved 3NF)
have failed to consider the effects of INDs on the structure
of a database, they are inadequate in characterizing a
database scheme which is truly devoid of redundancies.
In consideration of the above, we propose a new normal
form, called Inclusion Normal Form (IN-NF)...
\end{quote}
\end{frame}

\begin{frame}{E.F. Codd and the Relational Model of Data}
The following dedication is writ large in Codd's 1990 book:
\begin{quote}
To fellow pilots and aircrew
in the Royal Air Force
during World War II
and the dons at Oxford.
These people were the source of my determination to
fight for what I believed was right during the ten or
more years in which government, industry, and
commerce were strongly opposed to the relational
approach to database management.
\end{quote} 

By 2020 Oracle Corporation, founded in 1977, were the world's second largest software company with a 42\% share of an 
estimated \$30billion market for relational database technology. 
\end{frame}

\begin{frame}{E.F. Codd and the Relational Model of Data}
\displaybibentry{CoddBook1990}
\begin{quote}
The relational model is solidly based on two parts of mathematics: first-
order predicate logic and the theory of relations.
\end{quote} 

On the downside the relational model doesn't do anything to encourage  the programmer as navigator nor does it offer conceptualisation of navigation path equivalence i.e. commutative diagrams.
\end{frame}

